### 2021-11-03 paper/proposal/main.tex
adding project proposal

\documentclass[twocolumn, 12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum}
\usepackage{titling}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }

\urlstyle{same}

\title{When machine learning does you dirty: Robustness to measurement error in environmental economics}
\author{Aaron Watt}
\date{October 2021}
\setlength{\droptitle}{-10em}

% READ THE DISSERTATION THAT MEREDITH SENT CHU YU
% READ SOL'S GRAD STUDENTS' JOHNATHAN PROCTOR MEASUREMENT ERROR ML WORK
%     https://www.nber.org/papers/w28045
% GET PRUPLE AIR HISTORICAL DATA AND KEEP UPDATED VIA AWS FREE INSTANCE UPATING S3 BUCKET

\begin{document}
\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle

    \begin{abstract}
    % Research Question
    Spatially-dense machine-learning-predicted data are becoming more available and social scientists and regulators are availing themselves of it. However, these data are being treated as highly accurate and estimates of the measurement error in these predicted data are often excluded from analyses. Are pollution regulation policies and policy analyses sensitive to the inclusion of measurement error?

    % Proposed Model

    % Dataset
    Previous work by \href{https://www.aeaweb.org/articles?id=10.1257/pandp.20191064}{Fowlie, Rubin, and Walker (2019)} on the issue of measurement error in satellite data has explored two spatially dense ($<$1km$^2$ resolution) satellite-derived datasets that have been produced to estimate ground-level PM2.5 pollution. These data offer a feasible option for the EPA to regulate pollution without installing more expensive pollution monitors. Ground-level PM2.5 pollution concentration is estimated using aerosol optical depth (AOD) satellite data and various predictive models -- PM2.5 is not directly measurable (yet) using current satellite instruments.

    To help estimate measurement errors of these datasets, we can download PurpleAir data using their API.

    % Estimation and testing plan
    Adoption of PurpleAir monitors is non-random so I need to develop and estimate a selection model into PurpleAir adoption. PurpleAir monitors are also known to be less accurate and might be systematically biased, so estimating PM2.5 using PurpleAir measurements and the sparse (but highly accurate) EPA monitors will be important.

    \vspace{2em}
    \end{abstract}
  \end{@twocolumnfalse}
  ]

%Succinct description of research question; blurb describing a proposed model, dataset, and plan for estimation/testing.
\section{Research Question}
How much do prediction errors matter in pollution regulation? Does incorporating prediction errors of machine-learning-produced pollution data affect the policy categorization of areas without a pollution monitor?

\newpage
\section{Draft Project Outline}
\begin{enumerate}
    \item write the code to scrape the purple air data
    \item apply a selection model to the purple air data (since monitor adoption is non-random)
    \item write model of prediction errors of dense prediction of a spatially sparse, temporally dense variable using stats literature
    \item generate descriptive stats for purple air data and estimates of measurement error using EPA monitors as ground truth
    \item explore machine learning prediction of prediction errors using purple air data and using a leave-out group for validation testing
    \item produce draft tables of changes in attainment status
\end{enumerate}

% \section{Introduction}

% % Something about measurement error in environmental economics... robustness-to-measurement-error replication study, focusing on omitted measurement error from scientifically measured data (like temperature, elevation, pollution concentration, satellite approximated surface characteristics, etc).

% \section{Literature Review}

% \section{Methods/Data}

% \section{Results}

% \section{Discussion}

% \section{Conclusion}
\end{document}

### 2021-11-03 paper/report/summary_stats.Rmd
initial commit of summary stats R files

---
title: "Second Year Paper Summary Statistics"
author: "Aaron Watt"
email: aaron@acwatt.net
date: "2021-11-08"
output:
  pdf_document:
    toc: true
    toc_depth: 2
header-includes:
   - \usepackage{dcolumn}
   - \usepackage{amsmath}
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'ARE_213_PS2a.pdf')) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<!--
R version 3.6.3 (2020-02-29)
Purpose of script: Second Year Paper summary statistics

Notes: Need to ps2a.R from ps2a folder (without Rstudio being started first)
       if Rstudio is already started, the working directory will not be set
       to the ps2a/ folder

       \usepackage{dcolumn}: dcolumn is needed in latex compilation 
-->

```{r Settings, echo=FALSE}
# stargazer table type (html, latex, or text)
# Change to latex when outputting to PDF, html when outputting to html
table_type = "text"
```

```{r packages, results='hide', message=FALSE, echo=T}
library(tidyverse)
library(foreign)
library(stargazer)
```

### 2021-11-03 acwatt_syp_code/build/purpleair_download.py
Adding initial purple air download script

@@ -4,7 +4,7 @@
 
 # Built-in Imports
 # Third-party Imports
-from purpleair.network import SensorList
+from purpleair.network import SensorList, Sensor
 # Local Imports
 
 
@@ -28,10 +28,17 @@ Conditions of use:
 def list_sensors():
     """Return list of sensors that have valid data."""
     p = SensorList()  # Initialized 22,877 sensors!
-    # Other sensor filters include 'outside', 'all', 'family', and 'no_child'
-    df = p.to_dataframe(sensor_filter='useful',
+    # Other sensor filters include 'outside', 'all', 'useful', 'family', and 'no_child'
+    df = p.to_dataframe(sensor_filter='all',
                         channel='parent')
-    print(len(df))  # 16902
+    return df
+
+
+def filter_data(df):
+    df = (df.query('location_type == "outside"')
+          .query('downgraded == False')
+          .query('flagged == False')
+          )
     return df
 
 
@@ -41,6 +48,13 @@ def download_pm25(sensor_list):
         pass
 
 
+def get_data_from_sensor(sensor_id):
+    sensor = Sensor(sensor_id)
+    df = sensor.parent.get_historical(weeks_to_get=1,
+                                  thingspeak_field='secondary')
+    return df
+
 
 if __name__ == "__main__":
-    sensor_list = list_sensors()
+    sensor_df = list_sensors()
+    sensor_df = filter_data(sensor_df)

### 2021-11-03 acwatt_syp_code/analyze/maps.py
Adding maps

@@ -1,0 +1,67 @@
+#!/usr/bin/env python
+
+"""Functions for mapping data"""
+
+# Built-in Imports
+import pandas as pd
+import matplotlib.pyplot as plt
+import descartes
+import geopandas as gpd
+from geopandas.tools import sjoin
+from pathlib import Path
+from shapely.geometry import Point, Polygon
+# Third-party Imports
+# Local Imports
+
+save_dir = Path('output/summary_stats/maps')
+world = gpd.read_file(gpd.datasets.get_path('naturalearth_lowres'))
+states_shp = Path('data/cb_2018_us_state_5m/cb_2018_us_state_5m.shp')
+all_states = gpd.read_file(states_shp)
+
+lower48 = (all_states
+    .loc[all_states.STATEFP.astype(int) < 60]
+    .loc[~all_states.NAME.isin(['Alaska', 'Hawaii'])])
+california = all_states.loc[all_states.NAME.isin(['California'])]
+sensor_points = [Point(xy) for xy in zip(sensor_df['lon'], sensor_df['lat'])]
+sensors = gpd.GeoDataFrame(sensor_df, crs=all_states.crs, geometry=sensor_points)
+sensors_lower48 = sjoin(sensors, lower48, how='inner')
+sensors_california = sjoin(sensors, california, how='inner')
+
+polygon_kwgs = {'facecolor': "grey",
+                'edgecolor': 'blue',
+                'lw': 0.7,
+                'alpha': 0.5}
+marker_kwgs = {'facecolor': (1, 0, 0, 0.2),
+               'markersize': 20,
+               'marker': 'o'}
+
+plt.style.use('ggplot')
+fix, ax = plt.subplots(figsize=(15, 8))
+world.plot(ax=ax, **polygon_kwgs)
+sensors.plot(ax=ax, **marker_kwgs)
+plt.title("Valid Purple Air Monitor Locations, Global")
+plt.xlabel('Longitude')
+plt.ylabel('Latitude')
+plt.tight_layout()
+plt.savefig(save_dir / 'purple_air_sensor_map_global.png')
+
+
+fix, ax = plt.subplots(figsize=(15, 8))
+lower48.plot(ax=ax, **polygon_kwgs)
+sensors_lower48.plot(ax=ax, **marker_kwgs)
+plt.title("Valid Purple Air Monitor Locations, Contiguous United States")
+plt.xlabel('Longitude')
[…]

### 2021-11-03 acwatt_syp_code/analyze/maps.py
updating maps

@@ -36,7 +36,7 @@ marker_kwgs = {'facecolor': (1, 0, 0, 0.2),
                'marker': 'o'}
 
 plt.style.use('ggplot')
-fix, ax = plt.subplots(figsize=(15, 8))
+fix, ax = plt.subplots(figsize=(7.5, 4))
 world.plot(ax=ax, **polygon_kwgs)
 sensors.plot(ax=ax, **marker_kwgs)
 plt.title("Valid Purple Air Monitor Locations, Global")
@@ -46,7 +46,7 @@ plt.tight_layout()
 plt.savefig(save_dir / 'purple_air_sensor_map_global.png')
 
 
-fix, ax = plt.subplots(figsize=(15, 8))
+fix, ax = plt.subplots(figsize=(7, 4))
 lower48.plot(ax=ax, **polygon_kwgs)
 sensors_lower48.plot(ax=ax, **marker_kwgs)
 plt.title("Valid Purple Air Monitor Locations, Contiguous United States")
@@ -55,7 +55,7 @@ plt.ylabel('Latitude')
 plt.tight_layout()
 plt.savefig(save_dir / 'purple_air_sensor_map_us.png')
 
-fix, ax = plt.subplots(figsize=(8, 15))
+fix, ax = plt.subplots(figsize=(8, 8))
 california.plot(ax=ax, **polygon_kwgs)
 sensors_california.plot(ax=ax, **marker_kwgs)
 plt.title("Valid Purple Air Monitor Locations, California")

### 2021-11-03 paper/report/summary_stats.Rmd
Updating map colors to look coooool

@@
    - \usepackage{amsmath}
-knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'ARE_213_PS2a.pdf')) })
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'ARE219_SYP_summary-stats_acWatt.pdf')) })
 ---
@@
 
-```{r packages, results='hide', message=FALSE, echo=T}
+```{r packages, results='hide', message=FALSE, echo=F}
 library(tidyverse)
@@
 ```
+![Valid Purple Air Monitor Locations, Global](../../output/summary_stats/maps/purple_air_sensor_map_global.png)
+![Valid Purple Air Monitor Locations, Contiguous United States](../../output/summary_stats/maps/purple_air_sensor_map_us.png)
+![Valid Purple Air Monitor Locations, Global](../../output/summary_stats/maps/purple_air_sensor_map_california.png)

### 2021-11-03 acwatt_syp_code/analyze/maps.py

@@ -27,11 +27,10 @@ sensors = gpd.GeoDataFrame(sensor_df, crs=all_states.crs, geometry=sensor_points
 sensors_lower48 = sjoin(sensors, lower48, how='inner')
 sensors_california = sjoin(sensors, california, how='inner')
 
-polygon_kwgs = {'facecolor': "grey",
-                'edgecolor': 'blue',
-                'lw': 0.7,
-                'alpha': 0.5}
-marker_kwgs = {'facecolor': (1, 0, 0, 0.2),
+polygon_kwgs = {'facecolor': (100/255, 100/255, 100/255, 1),
+                'edgecolor': 'white',
+                'lw': 0.2}
+marker_kwgs = {'facecolor': (0, 204/255, 204/255, 0.2),
                'markersize': 20,
                'marker': 'o'}
 

### 2021-11-03 acwatt_syp_code/analyze/maps.py
Adding main run.py script and consolidating/modularizing the maps code.

@@ -10,57 +10,79 @@ import geopandas as gpd
 from geopandas.tools import sjoin
 from pathlib import Path
 from shapely.geometry import Point, Polygon
+
 # Third-party Imports
 # Local Imports
 
-save_dir = Path('output/summary_stats/maps')
-world = gpd.read_file(gpd.datasets.get_path('naturalearth_lowres'))
-states_shp = Path('data/cb_2018_us_state_5m/cb_2018_us_state_5m.shp')
-all_states = gpd.read_file(states_shp)
-
-lower48 = (all_states
-    .loc[all_states.STATEFP.astype(int) < 60]
-    .loc[~all_states.NAME.isin(['Alaska', 'Hawaii'])])
-california = all_states.loc[all_states.NAME.isin(['California'])]
-sensor_points = [Point(xy) for xy in zip(sensor_df['lon'], sensor_df['lat'])]
-sensors = gpd.GeoDataFrame(sensor_df, crs=all_states.crs, geometry=sensor_points)
-sensors_lower48 = sjoin(sensors, lower48, how='inner')
-sensors_california = sjoin(sensors, california, how='inner')
-
-polygon_kwgs = {'facecolor': (100/255, 100/255, 100/255, 1),
-                'edgecolor': 'white',
-                'lw': 0.2}
-marker_kwgs = {'facecolor': (0, 204/255, 204/255, 0.2),
-               'markersize': 20,
-               'marker': 'o'}
-
 plt.style.use('ggplot')
-fix, ax = plt.subplots(figsize=(7.5, 4))
-world.plot(ax=ax, **polygon_kwgs)
-sensors.plot(ax=ax, **marker_kwgs)
-plt.title("Valid Purple Air Monitor Locations, Global")
-plt.xlabel('Longitude')
-plt.ylabel('Latitude')
-plt.tight_layout()
-plt.savefig(save_dir / 'purple_air_sensor_map_global.png')
 
 
-fix, ax = plt.subplots(figsize=(7, 4))
-lower48.plot(ax=ax, **polygon_kwgs)
-sensors_lower48.plot(ax=ax, **marker_kwgs)
-plt.title("Valid Purple Air Monitor Locations, Contiguous United States")
-plt.xlabel('Longitude')
-plt.ylabel('Latitude')
-plt.tight_layout()
-plt.savefig(save_dir / 'purple_air_sensor_map_us.png')
-
[…]

### 2021-11-03 acwatt_syp_code/build/purpleair_download.py

@@ -55,6 +55,7 @@ def get_data_from_sensor(sensor_id):
     return df
 
 
-if __name__ == "__main__":
+def load_sensor_data():
     sensor_df = list_sensors()
     sensor_df = filter_data(sensor_df)
+    return sensor_df

### 2021-11-03 acwatt_syp_code/run.py

@@ -1,0 +1,14 @@
+#!/usr/bin/env python
+
+"""Module Docstring"""
+
+# Built-in Imports
+# Third-party Imports
+# Local Imports
+import analyze.maps as am
+import build.purpleair_download as pad
+
+if __name__ == "__main__":
+    sensor_df = pad.load_sensor_data()
+    am.make_all_sensor_maps(sensor_df)
+    print()

### 2021-11-03 paper/report/summary_stats.Rmd
Adding main run.py script and consolidating/modularizing the maps code.

@@
 ```
-![Valid Purple Air Monitor Locations, Global](../../output/summary_stats/maps/purple_air_sensor_map_global.png)
+![Valid Purple Air Monitor Locations, Global](../../output/summary_stats/maps/purple_air_sensor_map_world.png)
 ![Valid Purple Air Monitor Locations, Contiguous United States](../../output/summary_stats/maps/purple_air_sensor_map_us.png)

### 2021-11-03 acwatt_syp_code/analyze/maps.py
Increasing DPI of maps

@@ -51,7 +51,7 @@ def sensor_df_to_geo(df, area):
     return sensors_gdf, base_gdf
 
 
-def sensor_plot(sensor_df, area='world', rootname='are219'):
+def sensor_plot(sensor_df, area='world', rootname='are219', dpi=300):
     """Save png of plot of sensors in given geographical area.
 
     sensor_df = dataframe of sensor data, inlcuding columns = lat, lon
@@ -80,7 +80,7 @@ def sensor_plot(sensor_df, area='world', rootname='are219'):
     plt.xlabel('Longitude')
     plt.ylabel('Latitude')
     plt.tight_layout()
-    plt.savefig(save_dir / f'purple_air_sensor_map_{area}.png')
+    plt.savefig(save_dir / f'purple_air_sensor_map_{area}.png', dpi=dpi)
 
 
 def make_all_sensor_maps(sensor_df):

### 2021-11-03 acwatt_syp_code/run.py

@@ -11,4 +11,3 @@ import build.purpleair_download as pad
 if __name__ == "__main__":
     sensor_df = pad.load_sensor_data()
     am.make_all_sensor_maps(sensor_df)
-    print()

### 2021-11-03 acwatt_syp_code/build/purpleair_download.py
testing purple air data fields

@@ -58,4 +58,21 @@ def get_data_from_sensor(sensor_id):
 def load_sensor_data():
     sensor_df = list_sensors()
     sensor_df = filter_data(sensor_df)
+    # Single sensor
+
+    se = Sensor(int(sensor_df.index[0]))
+    print(se)
+    print(se.parent)
+    print(se.child)
+    print(se.parent.as_flat_dict())
+    se.get_field(3)
+    se.get_field(4)
+    print(se.thingspeak_data.keys())
+    df = se.parent.get_historical(weeks_to_get=1,
+                                  thingspeak_field='secondary')
+
+    print(df.head())
+    print(df.tail())
     return sensor_df
+
+

### 2021-11-04 acwatt_syp_code/utils/config.py
Setting up config file

@@ -1,0 +1,78 @@
+#!/usr/bin/env python
+# Python 3.7
+# File name: config.py
+# Authors: Aaron Watt
+# Date: 2021-11-03
+"""Module to be imported for project settings and paths."""
+
+# Built-in Imports
+from pathlib import Path
+
+
+# CLASSES --------------------------
+class Paths:
+    """Inner paths class to store project paths commonly used.
+
+    This will search the current working directory path for the name of the
+    repo (beecensus). Since this code is only called from main.py, and main.py
+    is inside the repo, it should be able to find the beecensus path.
+    This also means the name of the repo cannot be changed.
+    Since this is an inner class, paths will be accessible in the following way:
+    Project = ProjectSettings()  # instance of the outer class
+    Project.paths.root  # this will be the pathlib path to the github repo root
+    """
+    def __init__(self):
+        # add root path of the project / git repo
+        self.root = Path(*Path.cwd().parts[:Path.cwd().parts.index('are219') + 1])
+        # Top-level paths
+        self.code = self.root / 'acwatt_syp_code'
+        self.docs = self.root / 'docs'
+        self.output = self.root / 'output'
+        self.paper = self.root / 'paper'
+        # Data directories
+        self.data = Data(self.root / 'data')
+
+
+class Data:
+    """Inner inner paths class to store data file paths."""
+    def __init__(self, data_dir):
+        self.root = data_dir
+        self.checkpoints = self.root / 'checkpoints'
+        self.configs = self.root / 'configs'
+        self.gis = self.root / 'gis'
+        self.gis_county = self.gis / 'cb_2018_us_county_500k' / 'cb_2018_us_county_500k.shp'
+        self.gis_state = self.gis / 'cb_2018_us_state_5m' / 'cb_2018_us_state_5m.shp'
+        self.purpleair = self.root / 'purpleair'
+        self.tables = self.root / 'tables'
+        self.temp = self.root / 'temp'
+        self.test_data = self.root / 'test_data'
+        # Lookup tables
+        self.lookup_jpg = self.tables / 'tbl_jpg_lookup.csv'
+        self.lookup_fips = self.tables / 'tbl_fips_lookup.csv'
+
+
+class GISSettings:
+    """Class to hold settings for gis portion of project.
+
+    Possible geographical names:
+        - Philadelphia County, PA
+    """
+    def __init__(self):
+        # Name of geographical area to be used in the search for bees.
[…]

### 2021-11-04 paper/report/summary_stats.Rmd
Draft of summary stats -- cumulative sensor adoption over time vs daily AQI in CA

--- 2021-11-03 paper/report/summary_stats.Rmd
@@
    - \usepackage{amsmath}
-knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'ARE_213_PS2a.pdf')) })
+   - \usepackage{placeins}
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'ARE219_SYP_summary-stats_acWatt.pdf')) })
 ---
@@
 ```
+
+# Geographical Distribution of Purple Air Sensors
+
+We can see from the figure below that the majority of Purple Air sensors are currently concentrated in the United States and Europe.
+
+![Valid Purple Air Monitor Locations, Global](../../output/summary_stats/maps/purple_air_sensor_map_world.png)
+*Note*: These location data were collected in 2021, and I am unsure if there is a way to see the historical location of existing sensors or information on sensors that dropped out of the network.
+
+\FloatBarrier
+\newpage
+In the US, the sensors are concentrated on the coastlines, with the majority in California.
+
+![Valid Purple Air Monitor Locations, Contiguous United States](../../output/summary_stats/maps/purple_air_sensor_map_us.png)
+
+\FloatBarrier
+\newpage
+Within California, we can see the sensors are clustered around the Bay area and LA.
+
+![Valid Purple Air Monitor Locations, California](../../output/summary_stats/maps/purple_air_sensor_map_california.png)
+
+\FloatBarrier
+\newpage
+# Distribution of adoption over time
+
+For California, we can see from the below graph that adoption of Purple Air pollution monitors (in blue) increased steadily until summer of 2020, when adoption went up dramatically.
+There were particularly bad smoke days in populated areas of California that summer (e.g., "Mars Day").
+However, these data are biased toward recent increases in adoption because I only have access to the sensors that are currently in the network -- I have no data on the sensors that dropped out, therefore the increases pre-2020 are likely understated.
+
+A more interesting red graph would probably be population-distance-weighted fires per-month or a population-weighted measure of visibility.
+
+![Purple Air Adoption over time vs. Air Quality in California](../../output/summary_stats/graphs/purple_air_sensor_cum_time_california.png)
+
+\FloatBarrier
+\newpage
+# Data plan
+
+I am planning to limit my second year project to California and focus on comparing Purple Air PM2.5 data to EPA monitor data.
+I have not yet finished collecting PM2.5 data from Purple Air monitors but plan to collect years 2018-2021 for all California sensors via the Purple Air API.
+I am also in the process of collecting EPA PM2.5 data via the EPA API.
+
+In order to calibrate a selection model into Purple Air adoption, I plan to collect Census Block Group level demographic data and calibrate the model on the census block group level over 2018-2020 (inclusive).
+I also plan to collect weather data from PRISM.
+
+To model measurement error of sensors (EPA sensors vs.
+Purple Air sensors), I plan to use pollution distribution data generated for IPCC climate models.
+I've been told these data are available to download and are probably the best measure of "ground truth" that I have access to.

### 2021-11-04 acwatt_syp_code/analyze/maps.py

@@ -3,37 +3,30 @@
 """Functions for mapping data"""
 
 # Built-in Imports
+import json
 import pandas as pd
 import matplotlib.pyplot as plt
-import descartes
 import geopandas as gpd
-from geopandas.tools import sjoin
+import seaborn as sns
 from pathlib import Path
-from shapely.geometry import Point, Polygon
-
+from shapely.geometry import Point
 # Third-party Imports
 # Local Imports
+from ..utils.config import PATHS
 
 plt.style.use('ggplot')
-
-
-def root_path(rootname='are219'):
-    """Return pathlib path to root of git project."""
-    parts = Path.cwd().parts
-    i = parts.index(rootname)
-    return Path(*parts[:i + 1])
 
 
 def sensor_df_to_geo(df, area):
     """Return geopandas dataframes: one with sensor points, one with geographic area polygons.
 
-    sensor_df = dataframe of sensor data, inlcuding columns = lat, lon
+    df = dataframe of sensor data, inlcuding columns = lat, lon
     area = string indicating area = ['world', 'us', 'california']
     """
     if area == 'world':
         base_gdf = gpd.read_file(gpd.datasets.get_path('naturalearth_lowres'))
     elif area in ['us', 'california']:
-        path_ = root_path() / 'data' / 'cb_2018_us_state_5m' / 'cb_2018_us_state_5m.shp'
+        path_ = PATHS.data.gis / 'cb_2018_us_state_5m' / 'cb_2018_us_state_5m.shp'
         base_gdf = gpd.read_file(path_)
         if area == 'us':
             base_gdf = (base_gdf
@@ -41,34 +34,38 @@ def sensor_df_to_geo(df, area):
                 .loc[~base_gdf.NAME.isin(['Alaska', 'Hawaii'])])
         elif area == 'california':
             base_gdf = base_gdf.loc[base_gdf.NAME.isin(['California'])]
+    elif area == 'benton county':
+        path_ = PATHS.data.gis / 'cb_2018_us_county_500k' / 'cb_2018_us_county_500k.shp'
+        base_gdf = gpd.read_file(path_)
+        base_gdf = (base_gdf
+            .loc[base_gdf.STATEFP.astype(int) == 41]
+            .loc[base_gdf.NAME == 'Benton'])
     else:
         print(f'area = {area}: Not a supported area parameter.')
         raise
 
[…]

### 2021-11-04 acwatt_syp_code/build/purpleair_download.py

@@ -3,13 +3,15 @@
 """Functions to download purple air data via their API"""
 
 # Built-in Imports
+from datetime import datetime, timedelta
+import json
+import pandas as pd
 # Third-party Imports
+import purpleair.network as pan
 from purpleair.network import SensorList, Sensor
-# Local Imports
+# Local imports
+from ..utils.config import PATHS
 
-
-READ_KEY = '5498FF4F-1642-11EC-BAD6-42010A800017'
-WRITE_KEY = '5499CF6C-1642-11EC-BAD6-42010A800017'
 
 # If retrieving data from multiple sensors at once, please send a single request
 # rather than individual requests in succession.
@@ -25,12 +27,11 @@ Conditions of use:
 """
 
 
-def list_sensors():
+def df_sensors():
     """Return list of sensors that have valid data."""
-    p = SensorList()  # Initialized 22,877 sensors!
+    p = pan.SensorList()  # Initialized 22,877 sensors!
     # Other sensor filters include 'outside', 'all', 'useful', 'family', and 'no_child'
-    df = p.to_dataframe(sensor_filter='all',
-                        channel='parent')
+    df = p.to_dataframe(sensor_filter='useful', channel='parent')
     return df
 
 
@@ -42,24 +43,111 @@ def filter_data(df):
     return df
 
 
-def download_pm25(sensor_list):
+def dl_pm25(sensor_list):
     """Save dataframe of PM 2.5 data from all valid sensors."""
     for s in sensor_list:
         pass
 
 
-def get_data_from_sensor(sensor_id):
-    sensor = Sensor(sensor_id)
-    df = sensor.parent.get_historical(weeks_to_get=1,
-                                  thingspeak_field='secondary')
+def dl_from_sensor(sensor_id, start_date=datetime.now()):
+    df = (pan.Sensor(sensor_id)
+          .parent.get_historical(weeks_to_get=1,
+                                 thingspeak_field='secondary',
+                                 start_date=start_date))
     return df
 
 
-def load_sensor_data():
-    sensor_df = list_sensors()
-    sensor_df = filter_data(sensor_df)
-    # Single sensor
+def dl_in_geo_area(gdf, start_date, num_weeks=10):
+    """Download data from sensors in gdf between start_date and end_date.
 
+    gdf = geopandas dataframe that has already been filtered for sensors within
+        a geographic area using analyze.maps.sensor_df_to_geo()
[…]

### 2021-11-04 acwatt_syp_code/run.py

@@ -1,13 +1,0 @@
-#!/usr/bin/env python
-
-"""Module Docstring"""
-
-# Built-in Imports
-# Third-party Imports
-# Local Imports
-import analyze.maps as am
-import build.purpleair_download as pad
-
-if __name__ == "__main__":
-    sensor_df = pad.load_sensor_data()
-    am.make_all_sensor_maps(sensor_df)

### 2021-11-04 acwatt_syp_code/utils/__init__.py

@@ -1,0 +1,7 @@
+#!/usr/bin/env python
+
+"""Module Docstring"""
+
+# Built-in Imports
+# Third-party Imports
+# Local Imports

### 2021-11-04 acwatt_syp_code/utils/config.py

@@ -39,6 +39,7 @@ class Data:
         self.root = data_dir
         self.checkpoints = self.root / 'checkpoints'
         self.configs = self.root / 'configs'
+        self.epa = self.root / 'epa'
         self.gis = self.root / 'gis'
         self.gis_county = self.gis / 'cb_2018_us_county_500k' / 'cb_2018_us_county_500k.shp'
         self.gis_state = self.gis / 'cb_2018_us_state_5m' / 'cb_2018_us_state_5m.shp'
@@ -47,7 +48,7 @@ class Data:
         self.temp = self.root / 'temp'
         self.test_data = self.root / 'test_data'
         # Lookup tables
-        self.lookup_jpg = self.tables / 'tbl_jpg_lookup.csv'
+        self.lookup_location = self.tables / 'tbl_location_lookup.csv'
         self.lookup_fips = self.tables / 'tbl_fips_lookup.csv'
 
 
@@ -65,7 +66,8 @@ class GISSettings:
 class PurpleAirSettings:
     """Class to hold settings for Purple Air API when downloading data."""
     def __init__(self):
-        pass
+        self.read_key = '5498FF4F-1642-11EC-BAD6-42010A800017'
+        self.write_key = '5499CF6C-1642-11EC-BAD6-42010A800017'
 
 
 # FUNCTIONS --------------------------

### 2021-11-04 run.py

@@ -1,0 +1,17 @@
+#!/usr/bin/env python
+
+"""Module Docstring"""
+
+# Built-in Imports
+# Third-party Imports
+# Local Imports
+import acwatt_syp_code.analyze.maps as am
+import acwatt_syp_code.build.purpleair_download as pad
+from acwatt_syp_code.utils.config import PATHS
+
+if __name__ == "__main__":
+    sensor_df = pad.load_current_sensor_data()
+    # sensor_gdf, base_gdf = am.sensor_df_to_geo(sensor_df, 'california')
+
+    # pad.download_in_geo_area(sensor_gdf, )
+    am.make_all_sensor_maps(sensor_df)

### 2021-11-09 paper/report/summary_stats.Rmd
updating summary stats

@@
 library(stargazer)
+library(ggplot2)
+library(lubridate)
+library(kableExtra)
 ```
@@
 
+```{r summary stats, echo=F, warning=F}
+pa = read.csv('../../data/purpleair/date_created_california_manual_request_2021-11-09_metadata.csv') %>%
+    mutate(date_start = as.Date(date_start)) %>%
+    arrange(date_start) %>% mutate(order=(1:nrow(.))) %>%
+    select(-date_created)
+pa_counts = pa %>% count(COUNTYFP, year(date_start), sort = TRUE)
+
+year_stats0 = pa %>%
+    count(year(date_start))
+
+year_stats1 = pa_counts %>% 
+    group_by(`year(date_start)`) %>% 
+    summarize(max_n = max(n))
+pa_county = read.csv('../../data/purpleair/current_pa_sensors_metadata.csv')
+pop = read.csv('../../data/gis/co-est2019-annres-06.xlsx')
+aqi = read.csv('../../data/epa/aqi/daily_aqi_by_county_2016-2021.csv')
+# stargazer(pa, type='text')
+
+year_stats = pa %>%
+    group_by(year(date_start)) %>%
+    summarise(pm2.5_ym = mean(pm2.5, na.rm=T)) %>%
+    merge(., year_stats0) %>%
+    merge(., year_stats1)
+
+year_stats %>%
+    rename("Year Sensor Joined the Network"="year(date_start)",
+           "Mean PM2.5"="pm2.5_ym",
+           "Sensors joined in CA"=n,
+           "Max # of Sensors joining in any CA county"=max_n) %>%
+    kbl(caption = "Statistics of Purple Air Pollution Sensors Joining the Network", 
+        align = 'ccc', digits = 3) %>%
+    kable_minimal()
+
+pa2 = merge(pa, pa_county, by.x = "sensor_index", by.y = "id")
+
+```
+
+\FloatBarrier
+
+In the Figure 3 and Table 1, we see that many Purple Air pollution sensors joined
+the network in years of high pollution, or just after years of high pollution.
+One research question I hope to address is what might explain the spatial and
+temporal missingness of pollution data from county-level sensors that the EPA
+requires non-attainment counties to keep. Because Purple Air monitors create a 
+network spread over the state, I can estimate pollution in areas that don't have
+EPA monitors or in places that have EPA monitors but at times when they are shut
+off. However, from Figure 2, we can see the distribution of Purple Air monitors
+is not distributed evenly or randomly.
+
 \FloatBarrier
@@
 
-I am planning to limit my second year project to California and focus on comparing Purple Air PM2.5 data to EPA monitor data.
-I have not yet finished collecting PM2.5 data from Purple Air monitors but plan to collect years 2018-2021 for all California sensors via the Purple Air API.
-I am also in the process of collecting EPA PM2.5 data via the EPA API.
+<!-- This could become a section of your final paper.
+Describe the dataset.
+Provide summary statistics for relevant variables.
+Discuss issues that may affect your estimates.
+For example: measurement error, reverse causality, omitted variables bias, weak instruments, failure to converge.
+-->
 
-In order to calibrate a selection model into Purple Air adoption, I plan to collect Census Block Group level demographic data and calibrate the model on the census block group level over 2018-2020 (inclusive).
-I also plan to collect weather data from PRISM.
+<!-- Description: Several page writeup, suitable for inclusion in your final paper. Your advisor must sign off on this output. -->
 
-To model measurement error of sensors (EPA sensors vs.
-Purple Air sensors), I plan to use pollution distribution data generated for IPCC climate models.
-I've been told these data are available to download and are probably the best measure of "ground truth" that I have access to.
+I am planning to limit my second year project to California and focus on
+comparing Purple Air PM2.5 data to EPA monitor data. I have not yet finished
+collecting PM2.5 data from Purple Air monitors but plan to collect years
+2016-2021 for all California sensors via the Purple Air API. I am also in the
+process of collecting EPA PM2.5 data via the EPA API and have a small sample
+from Meredith.
+
+In order to calibrate a selection model into Purple Air adoption, I plan to
+collect Census Block Group level demographic data and calibrate the model on the
+census block group level over 2018-2020 (inclusive). I also plan to collect
+weather data from PRISM.
+
+To model measurement error of sensors (EPA sensors vs. Purple Air sensors), I
+plan to use pollution distribution data generated for IPCC climate models. I've
+been told these data are available to download and are probably the best measure
+of "ground truth" that I have access to.
+
+# Issues that will affect my estimates of pollution
+
+Measurement error from the Purple Air monitors will need to be addressed. I plan
+to explore a combination of averaging over multiple sensors and regression-based
+correction using EPA monitors that are near some Purple Air monitors as ground 
+truth pollution for that location.
+
+I also plan to estimate a selection model into Purple Air adoption so omitted
+variable bias will be a large concern. I will have census data at the census
+block group level that I can match to the location of the PA monitors. But that
+is not unit-level data so also suffers from measurement error if I am estimating
+an adoption model at the unit level. I am also interested in modeling this as
+a binary discrete choice model with block- or county-level adoption shares (percentage
+of the PA monitors adopted in that month, in that county). With either of these
+models of adoption, I will need to instrument PA sensor prices, but since it is
+a homogeneous good sold by only one firm and the prices have remained nearly the
+same for 4 years, I would predict I will have a weak instrument problem.

### 2021-11-09 acwatt_syp_code/analyze/maps.py

@@ -17,7 +17,7 @@ from ..utils.config import PATHS
 plt.style.use('ggplot')
 
 
-def sensor_df_to_geo(df, area):
+def sensor_df_to_geo(df, area, by=None):
     """Return geopandas dataframes: one with sensor points, one with geographic area polygons.
 
     df = dataframe of sensor data, inlcuding columns = lat, lon
@@ -33,7 +33,12 @@ def sensor_df_to_geo(df, area):
                 .loc[base_gdf.STATEFP.astype(int) < 60]
                 .loc[~base_gdf.NAME.isin(['Alaska', 'Hawaii'])])
         elif area == 'california':
-            base_gdf = base_gdf.loc[base_gdf.NAME.isin(['California'])]
+            if by is None:
+                base_gdf = base_gdf.loc[base_gdf.NAME.isin(['California'])]
+            elif by == 'county':
+                path_ = PATHS.data.gis / 'cb_2018_us_county_500k' / 'cb_2018_us_county_500k.shp'
+                base_gdf = gpd.read_file(path_)
+                base_gdf = base_gdf.loc[base_gdf.STATEFP.astype(int) == 6]
     elif area == 'benton county':
         path_ = PATHS.data.gis / 'cb_2018_us_county_500k' / 'cb_2018_us_county_500k.shp'
         base_gdf = gpd.read_file(path_)
@@ -98,8 +103,8 @@ def weighted_average(df, data_col, weight_col, by_col):
 
 
 def culumative_time_plot(dpi=300):
-    # Purple Air Data (date created for all current california sensors)
-    date_file = PATHS.data.test_data / 'date_created_california_manual_request.txt'
+    # Purple Air Data (date_created for all current california sensors)
+    date_file = PATHS.data.test_data / 'date_created_california_manual_request_2021-11-09.txt'
     dict_ = json.load(open(date_file,))
     df1 = (pd.DataFrame(dict_['data'], columns=dict_['fields'])
           .assign(date_start=lambda x: pd.to_datetime(x['date_created'], unit='s')))
@@ -143,14 +148,31 @@ def culumative_time_plot(dpi=300):
     save_dir = PATHS.output / 'summary_stats' / 'graphs'
     plt.savefig(save_dir / f'purple_air_sensor_cum_time_california.png', dpi=dpi)
 
[…]

### 2021-11-09 acwatt_syp_code/build/purpleair_download.py

@@ -29,9 +29,11 @@ Conditions of use:
 
 def df_sensors():
     """Return list of sensors that have valid data."""
-    p = pan.SensorList()  # Initialized 22,877 sensors!
+    p = pan.SensorList()  # Initialized 22,877 sensors!22,836
     # Other sensor filters include 'outside', 'all', 'useful', 'family', and 'no_child'
     df = p.to_dataframe(sensor_filter='useful', channel='parent')
+    # TODO: Cache and check if this has already been downloaded in past 7 days
+    #   if not, redownload and update the cache
     return df
 
 

### 2021-11-29 paper/5min-presentation/5min-presentation.Rmd
Added 5 min presentation

---
title: "Measuring bias in (strategically) missing EPA pollution data"
author: Aaron C Watt
date: December 2, 2021
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex
classoption: "aspectratio=169"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
# Had to run tinytex::install_tinytex()
```

# Motivation
## Clean Air
- The Clean Air Act (1970) established National Ambient Air Quality Standards
(NAAQS) for US counties.

- Binary standards: either "attainment" or "non-attainment"

<!-- - Annual and Daily standards that both must be met. -->

<!-- - Regulatory penalties levied on non-attainment areas can be large. -->

- Air quality can vary significantly throughout a day, week, quarter.

<!-- add space here -->
- NAAQS only requires each monitor to report 75\% of their readings, per quarter.

- EPA pollution monitors can be shut off for unreported reasons.

## Research Questions

- How biased is *missing* air pollution data from self-reporting US EPA monitors?

- Does this bias significantly change NAAQS attainment status?

# Project Overview
<!-- :
     latex_engine: xelatex
header-includes:
   - \usetheme[style=dark]{Nord}
   - \usepackage{YGbeamer}
(i) an interesting, well-posed question; 
(ii) an adequate model, such that the question makes sense within the context of
the model; 
(iii) an identification strategy, such that (again within the context of the 
model) the question can be answered; 
“What is your identification strategy?” = “What research design (and assumptions) are you using?”
(iv) a set of proposed hypothesis tests or statistics which map your
data into falsifiable statements that can provide answers to your question. You
should /not/ provide any results; this presentation is about empirical strategy,
not about data-dependent outcomes.
-->

## Previous Works
- Spotting Strategic Behavior in EPA Monitor Shutoffs:
Mu et al. 2021
<!-- [ put more here on strategic behavior evidence that motivates the interest in measuring -->
<!-- the size / distribution of bias in missing data/] -->

- Strategic decisions of monitor locations: Grainger et al. 2017

- Using satellite data to fill gaps in air pollution monitoring: Sullivan, Krupnick 2018

- Prediction error is significant in satellite-based estimates of air pollution: Fowlie, Rubin, Walker 2019

## This project
- Focus on EPA pollution data that is missing *in time*; limited to California.
- Using new consumer-based pollution monitors to understand the bias in EPA data.
- Avoids using satellite estimates (has been shown to have significant error).
<!-- - Understanding the distribution of bias in missing California air pollution data using consumer-based pollution -->
<!-- monitors that have low error near their location. -->

# Purple Air Monitors
[insert maps of California EPA and PA monitors, timelaps GIF? Timeline of adoption]

[insert pictures of PA outdoor monitors]

# Models

1. Predict EPA pollution at missing times

2. Estimate bias between predicted pollution at missing times and reported pollution at nonmissing times.

3. Estimate California counties' counterfactual attainment status using included predicted missing pollution data.

# Models: Predictive model of each EPA monitor PM2.5 pollution

\[
EPA_{i,t} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{j,k} PA_{j,t} + u_{i,t}
\]

- Analysis done at the month and quarter level; suppressing that subscript.
- $t$ is a unique hour within a given month or quarter.
- EPA monitor $i$ at time $t$ reads PM2.5 pollution $EPA_{i,t}$.
- For each EPA monitor $i$, there are $J_i$ Purple Air monitors within a 10-mile radius.
[insert diagram of two EPA monitors with PA monitors surrounding them]
- Purple Air monitor $j\in J_i$ at time $t$ reads PM2.5 pollution $PA_{j,t}$.
[insert diagram of one EPA monitor and surrounding PA monitors, with wind directions]

# Models: Hour-by-Day-of-week Bias of Missing EPA Monitor Pollution Data
\def\mm{\mathcal{M}}
\def\nn{\mathcal{N}}
\vspace{-1em}
\begin{align*}
\intertext{\textbf{Missingness Bias:}}
Bias_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t} 
- \frac{1}{|\mm_{h,d}|}\sum\limits_{t\in \mm_{h,d}}\widehat{EPA}_{i,t}\\
\text{where } \mm_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Missing}\};\\
  \nn_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Non-missing}\} \\
\uncover<2->{\intertext{We can also define the \textbf{algorithm bias}     as the Hour-by-Day-of-week prediction error}
\widetilde{Bias}_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t} 
- \frac{1}{|\mm_{h,d}|}\sum\limits_{t\in \mm_{h,d}}\widehat{EPA}_{i,t} \\}
\uncover<3->{\intertext{We can also define $Bias_{j,h,d}$ and $\widetilde{Bias}_{j,h,d}$ for PA monitor $j$ (we'll come back to this).}}
\end{align*}

<!-- \begin{itemize}[<+->] -->
<!-- \item \begin{align*} -->
<!-- Bias_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t}  -->
<!-- - \frac{1}{|\mm_{h,d}|}\sum\limits_{t\in \mm_{h,d}}\widehat{EPA}_{i,t}\\ -->
<!-- \text{where } \mm_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Missing}\};\\ -->
<!--   \nn_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Non-missing}\} -->
<!-- \end{align*} -->

<!-- \item We can also define the algorithm bias as the Hour-by-Day-of-week prediction error -->
<!-- \[ -->
<!-- \widetilde{Bias}_{i,h,d} = \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t}  -->
<!-- - \frac{1}{|\mm_{h,d}|}\sum\limits_{t\in \mm_{h,d}}\widehat{EPA}_{i,t} -->
<!-- \] -->

<!-- \item We can also define $Bias_{j,h,d}$ and $\widetilde{Bias}_{j,h,d}$ for PA monitor $j$ (we'll come back to this). -->
<!-- \end{itemize} -->

# Models: County Attainment Status

\begin{align*}
Attain^{annual}_c &= 1 \text{ if \textbf{reported} annual average PM2.5 below threshold}^*\\
  &= 1[\text{equation here}]\\[1em]
Attain^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{reported} daily average PM2.5 below threshold}^*\\
  &= 1[\text{equation here}]\\[1em]
\widehat{Attain}^{annual}_c &= 1 \text{ if \textbf{predicted} annual average PM2.5 below threshold}^*\\[1em]
\widehat{Attain}^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{predicted} daily average PM2.5 below threshold}^*
\end{align*}
$^*$averaged over 3 years in NAAQS standard.
[fill in equations and thresholds]

# Identification Strategy

- Direct causal link between reported pollution levels, attainment status, and
regulatory penalties / attainment requirements

- Assumption: nearby PurpleAir monitors that are good predictors for EPA monitors
during non-missing times will also be good predictors during missing times.
  - Specifically, reasons for EPA data missingness are not correlated with missingness
  or measurement error in PurpleAir data

# Proposed Statistical Test

- The $J_i$ group of PurpleAir sensors is (in a sense) a synthetic control for the EPA sensor $i$.

- **The question of bias can be stated**: are the data observed during the times when the EPA monitor is turned off significantly different from the data observed when the monitor is turned on? Is it more different than by random chance?

- Implies an Abadie et al. 2011 style permutation inference test for each EPA monitor $i$.

# Proposed Statistical Test (Is the bias larger than by random chance?)

- Calculate missingness bias and algorithm bias for the EPA monitor

- Pick PA monitor $j\in J_i$, temporarily remove data for original hours missing from EPA monitor ($\mm_{i,h,d}$).

- Construct similar placebo synthetic control for PA monitor $j$ (predict $\widehat{PA}_{j,t}$).

- Calculate missingness bias and algorithm bias for the PA monitor: $Bias_{j,h,d}$

- Repeat for all PA sensors.

- **Graphical test**: For EPA sensor $i$, compare graph of $Bias_{i,h,d}$ to placebo $Bias_{j,h,d}$ for $j\in J_i$.

- **Permutation inference p-value**:

  - Calculate sum of squared missingness bias and sum of squared algorithm bias for EPA sensor $i$ and PA sensors $j\in J_i$.
  \vspace{0.5em}
  - Ratio$_k$ = sum of squared missingness bias / sum of squared algorithm bias
  \vspace{0.5em}
  - p-value = $\frac{\text{\# of PA sensors in $i$'s radius with Ratio$_j$ larger than Ratio$_i$}}{\text{\# of PA sensors in $i$'s radius}}$

# Extensions

- Welfare analysis based on attainment status changes and required reductions in pollution.

- Comparing county population-weighted PM2.5 pollution to EPA sensors to estimate location-based bias.

# Appendix A: PurpleAir monitor correction factor

![PurpleAir correction equation for EPA monitor PM2.5 (RH = relative humidity, also measured by PA monitor)](PA_correction_equation.png){width=50%}

\beamerdefaultoverlayspecification{}
# Appendix B: Data Plan
### Datasets

- Hourly PM2.5 Pollution data from California EPA pollution monitors (2015-2020)

- 2-minute PM2.5 Pollution data from California PurpleAir sensors, hourly averages taken

### Deadlines
- Dec. 5: PurpleAir is downloading/averaging on 4 AWS tiny linux instances, sending CSVs to S3 bucket
- Dec. 12: Proof of concept for 2 EPA sensors (Fresno, and [need to pick another low on Mu's list])
- Dec. 19: Data warehouse setup and transfer of existing Purple Air data

# Appendix B: Data Plan

### Data Wearhouse
- AWS Linux Cassandra database (noSQL columnar, designed for large queries of columns)
- Python pushes and pulls

### Storage costs
- ~ 30,000 sensors, 50 variables, 2 minute intervals, 5 years of data = 107.55 Terabytes
- Depending on method of storage: \$2,900 - \$13,300 per month
- Only storing hourly means and SD: \$4 - \$15 per month

<!-- Extra
# Example
### Two EPA monitors:
- Fresno (Mu et al. identified as "interesting")
- other county low on the interesting list

# Data Plan
### Datasets

### Calendar

### Data Wearhouse
- AWS Linux Cassandra database (noSQL columnar, designed for large queries of columns)

### storage costs
- ~ 30,000 sensors, 50 variables, 2 minute intervals, 5 years of data = 107.55 Terabytes
- Depending on method of storage: \$2,900 - \$13,300 per month
- Only storing hourly means and std. deviations: \$4 - \$15 per month

-->

### 2021-11-29 paper/5min-presentation/5min-presentation.Rmd
updated version of 5 min pres

@@
 classoption: "aspectratio=169"
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'acWatt_purpleair_5min-pres_2021-12-02.pdf')) })
 ---
@@
 
+Source: \url{https://www.epa.gov/air-sensor-toolbox/technical-approaches-sensor-data-airnow-fire-and-smoke-map}
+
 \beamerdefaultoverlayspecification{}

### 2021-11-29 paper/5min-presentation/5min-presentation.Rmd
commenting

--- 2021-11-29 paper/5min-presentation/5min-presentation.Rmd
@@
 library(tinytex)
-# Had to run tinytex::install_tinytex()
+# Had to run tinytex::install_tinytex() to use XeLaTeX if using YGbeamer.sty
 ```

### 2021-11-30 acwatt_syp_code/analyze/maps.py
modified to get all the sensors available in the US.

@@ -32,7 +32,8 @@ def sensor_df_to_geo(df, area, by=None):
         if area == 'us':
             base_gdf = (base_gdf
                 .loc[base_gdf.STATEFP.astype(int) < 60]
-                .loc[~base_gdf.NAME.isin(['Alaska', 'Hawaii'])])
+                # .loc[~base_gdf.NAME.isin(['Alaska', 'Hawaii'])]
+            )
         elif area == 'california':
             if by is None:
                 base_gdf = base_gdf.loc[base_gdf.NAME.isin(['California'])]
@@ -105,7 +106,8 @@ def weighted_average(df, data_col, weight_col, by_col):
 
 def culumative_time_plot(dpi=300):
     # Purple Air Data (date_created for all current california sensors)
-    date_file = PATHS.data.test_data / 'date_created_california_manual_request_2021-11-09.txt'
+    date_ = '2021-11-29'
+    date_file = PATHS.data.test_data / f'date_created_california_manual_request_{date_}.txt'
     dict_ = json.load(open(date_file,))
     df1 = (pd.DataFrame(dict_['data'], columns=dict_['fields'])
           .assign(date_start=lambda x: pd.to_datetime(x['date_created'], unit='s')))
@@ -168,11 +170,12 @@ def intersect_pa_counties(sensor_df):
     sensor_gdf = sensor_gdf.reset_index()
     sensor_gdf.to_csv(PATHS.data.purpleair / 'current_pa_sensors_metadata.csv', index=False)
 
+
 def make_all_sensor_maps(sensor_df):
-    intersect_pa_counties(sensor_df)
-    # for area in ['world', 'us', 'california', 'benton county']:
-    #     sensor_plot(sensor_df, area=area)
-    culumative_time_plot()
+    for area in ['world', 'us', 'california', 'benton county']:
+        sensor_plot(sensor_df, area=area)
+    # intersect_pa_counties(sensor_df)
+    # culumative_time_plot()
 
 
 

### 2021-11-30 run.py

@@ -10,8 +10,17 @@ import acwatt_syp_code.build.purpleair_download as pad
 from acwatt_syp_code.utils.config import PATHS
 
 if __name__ == "__main__":
+    # Get available sensor locations from  from purpleair.network.SensorList API wrapper
     sensor_df = pad.load_current_sensor_data()
-    sensor_gdf, base_gdf = am.sensor_df_to_geo(sensor_df, 'california')
+    # Make sensor plots
+    # am.make_all_sensor_maps(sensor_df)
+    # Save sensor dataframe
+    sensor_gdf, base_gdf = am.sensor_df_to_geo(sensor_df, 'us')
+    (sensor_gdf
+     .reset_index()
+     .to_csv(PATHS.data.purpleair / 'location_us_pan_request_2021-11-29_geocoded.csv',
+             index=False))
+
+    # Start downloading PA sensor data
 
     pad.dl_in_geo_area(sensor_gdf, start_date, num_weeks=10)
-    am.make_all_sensor_maps(sensor_df)

### 2021-11-30 paper/5min-presentation/5min-presentation.Rmd
Updating 5 min presentation with map of EPA and PA sensors.

@@
 
+```{r packages, include=F}
+library(tidyverse)
+library(docstring)
+library(ggplot2)
+# library(mapdata)
+# Requires rgdal package to be installed, requires gdal to be installed on computer
+# Ubuntu 20.04: sudo aptitude install gdal-bin proj-bin libgdal-dev libproj-dev
+```
+
 # Motivation
 ## Clean Air
+
 - The Clean Air Act (1970) established National Ambient Air Quality Standards
@@
 
+- Air quality can vary significantly throughout a day, week, quarter.
+
+- NAAQS only requires each monitor to report 75\% of their readings, per quarter.
+
+- EPA pollution monitors can be shut off for unreported reasons.
+
 <!-- - Annual and Daily standards that both must be met. -->
@@
 
-- Air quality can vary significantly throughout a day, week, quarter.
-
-<!-- add space here -->
-- NAAQS only requires each monitor to report 75\% of their readings, per quarter.
-
-- EPA pollution monitors can be shut off for unreported reasons.
-
 ## Research Questions
 
-- How biased is *missing* air pollution data from self-reporting US EPA monitors?
+- How biased are EPA monitor-based measures of local air quality?
+<!-- - How biased are self-reported EPA pollution estimates? -->
+<!-- - How biased is *missing* air pollution data from self-reporting US EPA monitors? -->
 
@@
 
-# Project Overview
+# Context
 <!-- :
@@
 ## This project
-- Focus on EPA pollution data that is missing *in time*; limited to California.
-- Using new consumer-based pollution monitors to understand the bias in EPA data.
+- Using new consumer-based pollution monitors (PurpleAir).
+
+- Focus on EPA pollution data that is missing *in time*.
+
 - Avoids using satellite estimates (has been shown to have significant error).
@@
 
-# Purple Air Monitors
-[insert maps of California EPA and PA monitors, timelaps GIF? Timeline of adoption]
+<!-- [why California? Maybe should inlcude all EPA monitor - months that have at least one PA monitor within 10 miles] -->
 
-[insert pictures of PA outdoor monitors]
+<!-- [insert maps of California EPA and PA monitors, timelaps GIF? Timeline of adoption] -->
+
+<!-- [insert pictures of PA outdoor monitors] -->
+
+```{r geography intersection functions, include=F}
+library(sf)
+library(spData)
+
+lonlat_to_state <- function(df, lon_col, lat_col,
+                            states = spData::us_states,
+                            name_col = "NAME") {
+  #' State names for lon, lat vectors
+  #' 
+  #' @description Return a vector of state names for the lon-lat vectors
+  #' 
+  #' @param df: a data.frame containing lon lat pairs
+  #' 
+  #' @param lon_col: A character of the column name containing longitude
+  #' 
+  #' @param lat_col: A character of the column name containing latitude
+  #'
+  #' @param states:   An sf MULTIPOLYGON object with 50 states plus DC.
+  #'
+  #' @param name_col: Name of a column in `states` that supplies the states'
+  #'           names.
+  #' https://stackoverflow.com/a/8751965/16660792
+  ## Convert points data.frame to an sf POINTS object
+  pointsDF = data.frame(x = df[, lon_col], y = df[, lat_col])
+  pts <- st_as_sf(pointsDF, coords = 1:2, crs = 4326)
+
+  ## Transform spatial data to some planar coordinate system
+  ## (e.g. Web Mercator) as required for geometric operations
+  states <- st_transform(states, crs = 3857)
+  pts <- st_transform(pts, crs = 3857)
+
+  ## Find names of state (if any) intersected by each point
+  state_names <- states[[name_col]]
+  ii <- as.integer(st_intersects(pts, states))
+  state_names[ii]
+}
+
+lonlat_country_filter <- function(df, lon_col, lat_col,
+                                  country='United States',
+                                  name_col='name_long') {
+  #' Filter for lat-lons in a specific country
+  #' 
+  #' @description Return the dataframe filtered to inlcude only US points
+  #' 
+  #' @param df: a data.frame containing lon lat pairs
+  #' 
+  #' @param lon_col: A character of the column name containing longitude
+  #' 
+  #' @param lat_col: A character of the column name containing latitude
+  #'
+  #' @param country: A character of the country, matching name_col from spData::world
+  #'
+  #' @param name_col: A character of the column name containing the country name,
+  #' from spData::world
+  #'
+  ## Restrict to county boundaries
+  country = spData::world %>%
+    filter(!!as.symbol(name_col) == country)
+  ## Convert points data.frame to an sf POINTS object
+  pointsDF = data.frame(x = df[, lon_col], y = df[, lat_col])
+  pts <- st_as_sf(pointsDF, coords = 1:2, crs = 4326)
+
+  ## Transform spatial data to some planar coordinate system
+  ## (e.g. Web Mercator) as required for geometric operations
+  country <- st_transform(country, crs = 3857)
+  pts <- st_transform(pts, crs = 3857)
+
+  ## Find names of state (if any) intersected by each point
+  ii <- as.integer(st_intersects(pts, country))
+  df %>%
+    mutate(include = ii) %>%
+    filter(include == 1)
+}
+
+```
+
+# PM2.5 Air Pollution Monitors
+
+```{r get EPA sensor locations 2016, include=F, cache=T}
+sensor_types = c(88101, 88501, 88502)
+years = c()
+
+load("/media/a/E/Programming/github/are219/data/epa/epa_monitors/epa_monitors/data_from_api/88101/api_request_daily_88101_2016.Rdata")
+df = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88101, NAAQS_used='yes')
+load("/media/a/E/Programming/github/are219/data/epa/epa_monitors/epa_monitors/data_from_api/88501/api_request_daily_88501_2016.Rdata")
+df2 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88501, NAAQS_used='no')
+load("/media/a/E/Programming/github/are219/data/epa/epa_monitors/epa_monitors/data_from_api/88502/api_request_daily_88502_2016.Rdata")
+df3 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88502, NAAQS_used='no')
+epa_data = bind_rows(df, df2, df3) %>%
+  filter(type == 88101) %>% 
+  distinct(latitude, longitude, .keep_all = TRUE) %>%
+  select(longitude, latitude, everything())
+
+```
+
+```{r get PA sensor locations all, include=F}
+result = jsonlite::fromJSON("/media/a/E/Programming/github/are219/data/test_data/manual_webpage_download_USmap_2021-11-29.txt")
+library(zoo)
+# Get start and last_seen dates from all sensors near the US
+pa_data2 = as.data.frame(result$data) %>%
+  purrr::set_names(result$fields) %>%
+  mutate(date_start = lubridate::as_datetime(date_created),
+         date_last = lubridate::as_datetime(last_seen),
+         month_start = as.yearmon(date_start),
+         month_last = as.yearmon(date_last)) %>%
+  select(longitude, latitude, everything()) %>%
+  arrange(date_start) %>%
+  lonlat_country_filter(lon_col='longitude', lat_col='latitude') %>%
+  filter(private == 0,  # public sensor
+         location_type == 0,  # outdoor sensor
+         channel_state != 0)  # has at least one PM sensor
+```
+
+```{r Plot the locations, eval=F, include=F}
+min_lat = min(pa_data$lat)
+max_lat = max(pa_data$lat)
+min_lon = min(pa_data$lon)
+max_lon = max(pa_data$lon)
+print(paste(max_lon, max_lat, min_lon, min_lat))
+usa <- map_data('usa', projection='conic', parameters=c(min_lat))
+
+b = 204/255
+plot_usmap(fill='white', color='grey') +
+  geom_point(data = usmap::usmap_transform(pa_data),   # Purple Air
+             aes(x = lon.1, y = lat.1),
+             color = rgb(0, b, b), alpha = 1, size=0.4) +
+  geom_point(data = usmap::usmap_transform(epa_data),   # EPA
+             aes(x = longitude.1, y = latitude.1),
+             color = "red", alpha = 0.25, size=2) +
+  labs(title = "US EPA & PurpleAir Pollution Monitors",
+       subtitle = "Source: EPA, PurpleAir.comJan 1 to Jun 30 2016") +
+  theme(legend.position = "right")
+
+```
+
+```{r Monitor Map with AK and HI, include=F}
+# devtools::install_github("hrbrmstr/albersusa")
+# sudo apt-get install libudunits2-dev
+# install.packages('units')
+# install.packages('sf')
+
+# Composite map with AK & HI
+usa_map = albersusa::usa_composite(proj="laea") %>%
+  fortify(.)  # make it usable in ggplot2
+epa_trans <- usmap::usmap_transform(epa_data)
+pa_trans <- usmap::usmap_transform(pa_data2)
+
+# Plot PA and EPA sensor locations on US map
+ggplot() +
+  geom_map(data=usa_map, map=usa_map,
+                    aes(map_id=id),
+                    color="#2b2b2b", size=0.1, fill=NA) +
+  geom_point(data = pa_trans, 
+             aes(x = longitude.1, y = latitude.1),  # Purple Air
+             color = rgb(0, 204/255, 204/255), alpha = 1, size=0.4) +
+  geom_point(data = epa_trans, aes(x = longitude.1, y = latitude.1),  # EPA
+             color = "red", alpha = 0.25, size=2) +
+  labs(title = "US EPA & PurpleAir Pollution Monitors",
+       subtitle = "Source: EPA 2016, PurpleAir.com 2015-2021") +
+  ggthemes::theme_map()
+
+ggsave('monitor_map_2015-2021.png', width = 6, height = 4)
+
+```
+
+\vspace{-2em}
+
+```{r Map and images, echo=F, warning=F, message=F, fig.width=12, fig.cap='test'}
+p_pa = "../../docs/purpleair/PA_outdoor_monitor_border.png"
+p_epa = "../../docs/epa/air-monitoring-site_border.jpg"
+p_map = 'monitor_map_2015-2021.png'
+library(png)
+library(jpeg)
+library(grid)
+library(gridExtra)
+img_map = rasterGrob(as.raster(readPNG(p_map)), interpolate = FALSE)
+img1 <-  rasterGrob(as.raster(readPNG(p_pa)), interpolate = FALSE)
+img2 <-  rasterGrob(as.raster(readJPEG(p_epa)), interpolate = FALSE)
+
+lay = rbind(c(1,2),
+            c(1,3))
+grid.arrange(img_map, img1, img2,
+             layout_matrix = lay,
+             widths = c(3,1))
+```
+
+```{r animate purpleair adoption, eval=F, include=F}
+##Source: https://github.com/thomasp85/gganimate
+library(ggplot2)
+library(gganimate)
+library(gapminder)
+## Warning: package 'gapminder' was built under R version 3.4.4
+library(gifski)
+## Warning: package 'gifski' was built under R version 3.4.4
+
+ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
+  geom_point(alpha = 0.7, show.legend = FALSE) +
+  scale_colour_manual(values = country_colors) +
+  scale_size(range = c(2, 12)) +
+  scale_x_log10() +
+  facet_wrap(~continent) +
+  # Here comes the gganimate specific bits
+  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
+  transition_time(year) +
+  ease_aes('linear')
+
+month_counts = pa_data2 %>%
+  group_by()
+
+dates <- as.Date(df1$date, '%d/%m/%Y')
+yr <- year(dates)
+monyr <- as.yearmon(dates)
+lst <- lapply(list(dates, yr, monyr), function(x) 
+       transform(df1, Count=ave(seq_along(x), x, FUN= length)))
+names(lst) <- paste0('newdf', seq_along(lst))
+list2env(lst, envir=.GlobalEnv)
+
+```
 
@@
 
-1. Predict EPA pollution at missing times
-
-2. Estimate bias between predicted pollution at missing times and reported pollution at nonmissing times.
-
-3. Estimate California counties' counterfactual attainment status using included predicted missing pollution data.
+\begin{itemize}%[itemsep=1.5em]
+\item \textbf{Estimate missing pollution observations:} Use PurpleAir data to predict EPA pollution at missing times
+\item \textbf{Estimate bias of reported EPA pollution:} difference between predicted pollution at missing times and reported pollution at nonmissing times.
+\item \textbf{Estimate counties' counterfactual attainment status:} Include estimated missing pollution data.
+\end{itemize}
 
@@
 \[
-EPA_{i,t} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{j,k} PA_{j,t} + u_{i,t}
+EPA_{i,t} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{j,k} PA_{j,t} \cdot Winddir_{i,t,k} + u_{i,t}
 \]
@@
 - Analysis done at the month and quarter level; suppressing that subscript.
+
 - $t$ is a unique hour within a given month or quarter.
+
 - EPA monitor $i$ at time $t$ reads PM2.5 pollution $EPA_{i,t}$.
+
 - For each EPA monitor $i$, there are $J_i$ Purple Air monitors within a 10-mile radius.
-[insert diagram of two EPA monitors with PA monitors surrounding them]
+
 - Purple Air monitor $j\in J_i$ at time $t$ reads PM2.5 pollution $PA_{j,t}$.
-[insert diagram of one EPA monitor and surrounding PA monitors, with wind directions]
+
+<!-- [insert diagram of two EPA monitors with PA monitors surrounding them] -->
+
+- $Winddir_{i,t,k}$ is a wind direction indicator; 1 if the prevailing wind near station $i$ at time $t$ is in the $k^{th}$ bucket (of 8 buckets).
+
+<!-- [insert diagram of one EPA monitor and surrounding PA monitors, with wind directions] -->
 
@@
 \widetilde{Bias}_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t} 
-- \frac{1}{|\mm_{h,d}|}\sum\limits_{t\in \mm_{h,d}}\widehat{EPA}_{i,t} \\}
+- \frac{1}{|\nn_{h,d}|}\sum\limits_{t\in \nn_{h,d}}\widehat{EPA}_{i,t} \\}
 \uncover<3->{\intertext{We can also define $Bias_{j,h,d}$ and $\widetilde{Bias}_{j,h,d}$ for PA monitor $j$ (we'll come back to this).}}
@@
 \begin{align*}
-Attain^{annual}_c &= 1 \text{ if \textbf{reported} annual average PM2.5 below threshold}^*\\
-  &= 1[\text{equation here}]\\[1em]
-Attain^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{reported} daily average PM2.5 below threshold}^*\\
-  &= 1[\text{equation here}]\\[1em]
+Attain^{annual}_c &= 1 \text{ if \textbf{reported} annual average PM2.5 below 15.0 $\mu$g/m$^3$}^*\\[1em]
+Attain^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{reported} daily average PM2.5 below 35 $\mu$g/m$^3$}^*\\[1em]
 \widehat{Attain}^{annual}_c &= 1 \text{ if \textbf{predicted} annual average PM2.5 below threshold}^*\\[1em]
@@
 during non-missing times will also be good predictors during missing times.
+
   - Specifically, reasons for EPA data missingness are not correlated with missingness
@@
 
+# Proposed Statistical Test (Is the bias larger than by random chance?)
+
 - **Graphical test**: For EPA sensor $i$, compare graph of $Bias_{i,h,d}$ to placebo $Bias_{j,h,d}$ for $j\in J_i$.
@@
   - Calculate sum of squared missingness bias and sum of squared algorithm bias for EPA sensor $i$ and PA sensors $j\in J_i$.
-  \vspace{0.5em}
+  \vspace{1em}
   - Ratio$_k$ = sum of squared missingness bias / sum of squared algorithm bias
-  \vspace{0.5em}
+  \vspace{1em}
   - p-value = $\frac{\text{\# of PA sensors in $i$'s radius with Ratio$_j$ larger than Ratio$_i$}}{\text{\# of PA sensors in $i$'s radius}}$
+
+<!-- But how does this account for poor prediction of the monitor?
+Because the denominator is larger if poor prediction in non-missing observations -->
 
@@
 
-![PurpleAir correction equation for EPA monitor PM2.5 (RH = relative humidity, also measured by PA monitor)](PA_correction_equation.png){width=50%}
+![PurpleAir correction equation for EPA monitor PM2.5 (RH = relative humidity, also measured by PA monitor)](../../docs/purpleair/PA_correction_equation.png){width=50%}
 
@@
 
+# Appendix C: PurpleAir Takeup
+![Valid Purple Air Monitor Locations, Contiguous United States](../../output/summary_stats/graphs/purple_air_sensor_cum_time_california.png){width=60%}
+
 <!-- Extra

### 2021-12-07 paper/5min-presentation/5min-presentation.Rmd
Adding final version of presentation and some files

@@
 ---
-title: "Measuring bias in (strategically) missing EPA pollution data"
+title: "Estimation and Implications of Bias in EPA Pollution Measurement"
 author: Aaron C Watt
@@
 - The Clean Air Act (1970) established National Ambient Air Quality Standards
-(NAAQS) for US counties.
+(NAAQS) for US counties
 
-- Binary standards: either "attainment" or "non-attainment"
+- Either "attainment" or "non-attainment", \color{red}large penalties and costly restrictions
 
-- Air quality can vary significantly throughout a day, week, quarter.
+- Minimum requirement of 75\% of readings, per quarter
 
-- NAAQS only requires each monitor to report 75\% of their readings, per quarter.
+- Air quality can change quickly
 
-- EPA pollution monitors can be shut off for unreported reasons.
+- Monitor shutoffs are common
 
@@
 
-## Previous Works
-- Spotting Strategic Behavior in EPA Monitor Shutoffs:
-Mu et al. 2021
+## Previous Works in EPA Pollution Monitors
+
+**Satellite Data:**
+
+- Grainger et al. 2017
+- Sullivan, Krupnick 2018
+- Fowlie, Rubin, Walker 2019
+- Zou 2021
+
+**Location Pollution Alerts:**
+
+- Mu, Rubin, Zou 2021
+
+<!-- - Spotting Strategic Behavior in EPA Monitor Shutoffs: Mu, Rubin, Zou 2021 -->
 <!-- [ put more here on strategic behavior evidence that motivates the interest in measuring -->
@@
 
-- Strategic decisions of monitor locations: Grainger et al. 2017
+<!-- - Strategic decisions of monitor locations: Grainger et al. 2017 -->
 
-- Using satellite data to fill gaps in air pollution monitoring: Sullivan, Krupnick 2018
+<!-- - Using satellite data to fill gaps in air pollution monitoring: Sullivan, Krupnick 2018 -->
 
-- Prediction error is significant in satellite-based estimates of air pollution: Fowlie, Rubin, Walker 2019
+<!-- - Prediction error is significant in satellite-based estimates of air pollution: Fowlie, Rubin, Walker 2019 -->
 
@@
 
-- Focus on EPA pollution data that is missing *in time*.
+- Avoids using satellite estimates (has been shown to have significant error).
 
-- Avoids using satellite estimates (has been shown to have significant error).
+- Focus on EPA pollution data that is missing *in time*, specifically PM2.5
 <!-- - Understanding the distribution of bias in missing California air pollution data using consumer-based pollution -->
@@
 
-# PM2.5 Air Pollution Monitors
+# Use PM2.5 Air Pollution Monitors to Predict Missing EPA Data
 
@@
 \[
-EPA_{i,t} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{j,k} PA_{j,t} \cdot Winddir_{i,t,k} + u_{i,t}
+\alert<2>{EPA_{i,t}} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{j,k} 
+\alert<3>{PA_{j,t}} \cdot \alert<4>{Winddir_{i,t,k}} + u_{i,t}
 \]
 
-- Analysis done at the month and quarter level; suppressing that subscript.
-
-- $t$ is a unique hour within a given month or quarter.
-
-- EPA monitor $i$ at time $t$ reads PM2.5 pollution $EPA_{i,t}$.
-
-- For each EPA monitor $i$, there are $J_i$ Purple Air monitors within a 10-mile radius.
-
-- Purple Air monitor $j\in J_i$ at time $t$ reads PM2.5 pollution $PA_{j,t}$.
+\begin{itemize}
+\item<1-> Analysis done at the month and quarter level; suppressing that subscript.
+\item<1-> $t$ is a unique hour within a given month or quarter.
+\item<1-> EPA monitor $i$ at time $t$ reads PM2.5 pollution \alert<2>{$EPA_{i,t}$}.
+\item<1-> For each EPA monitor $i$, there are $J_i$ Purple Air monitors within a 10-mile radius.
+\item<1-> Purple Air monitor $j\in J_i$ at time $t$ reads PM2.5 pollution \alert<3>{$PA_{j,t}$}.
+\item<1-> $\alert<4>{Winddir_{i,t,k}} $ is a wind direction indicator; 1 if the prevailing wind near station $i$ at time $t$ is in the $k^{th}$ bucket (of 8 buckets).
+\end{itemize}
 
 <!-- [insert diagram of two EPA monitors with PA monitors surrounding them] -->
-
-- $Winddir_{i,t,k}$ is a wind direction indicator; 1 if the prevailing wind near station $i$ at time $t$ is in the $k^{th}$ bucket (of 8 buckets).
 
@@
 Bias_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t} 
-- \frac{1}{|\mm_{h,d}|}\sum\limits_{t\in \mm_{h,d}}\widehat{EPA}_{i,t}\\
+- \frac{1}{|\mm_{i,h,d}|}\sum\limits_{t\in \mm_{i,h,d}}\widehat{EPA}_{i,t}\\
 \text{where } \mm_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Missing}\};\\
@@
 \widetilde{Bias}_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t} 
-- \frac{1}{|\nn_{h,d}|}\sum\limits_{t\in \nn_{h,d}}\widehat{EPA}_{i,t} \\}
+- \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}\widehat{EPA}_{i,t} \\}
 \uncover<3->{\intertext{We can also define $Bias_{j,h,d}$ and $\widetilde{Bias}_{j,h,d}$ for PA monitor $j$ (we'll come back to this).}}
@@
 Attain^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{reported} daily average PM2.5 below 35 $\mu$g/m$^3$}^*\\[1em]
-\widehat{Attain}^{annual}_c &= 1 \text{ if \textbf{predicted} annual average PM2.5 below threshold}^*\\[1em]
-\widehat{Attain}^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{predicted} daily average PM2.5 below threshold}^*
+\uncover<2->{\widehat{Attain}^{annual}_c &= 1 \text{ if \textbf{predicted} annual average PM2.5 below 15.0 $\mu$g/m$^3$}^*}\\[1em]
+\uncover<2->{\widehat{Attain}^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{predicted} daily average PM2.5 below 35 $\mu$g/m$^3$}^*}
 \end{align*}
@@
 
-# Identification Strategy
+# Identification
 
-- Direct causal link between reported pollution levels, attainment status, and
-regulatory penalties / attainment requirements
+\uncover<2->{\textbf{Is the missing pollution data identified?}}
 
-- Assumption: nearby PurpleAir monitors that are good predictors for EPA monitors
+\begin{itemize}
+\item<3-> Assumption: nearby PurpleAir monitors that are good predictors for EPA monitors
 during non-missing times will also be good predictors during missing times.
+\item<3-> Specifically, reasons for EPA data missingness are not correlated with missingness
+  or measurement error in PurpleAir data
+\end{itemize}
 
-  - Specifically, reasons for EPA data missingness are not correlated with missingness
-  or measurement error in PurpleAir data
+\uncover<4->{\textbf{Are attainment status changes identified?}}
+
+\begin{itemize}
+\item<5-> Yes: simple rule based on pollution concentrations.
+\end{itemize}
 
@@
 
-- **The question of bias can be stated**: are the data observed during the times when the EPA monitor is turned off significantly different from the data observed when the monitor is turned on?
-Is it more different than by random chance?
+- **The question of bias in EPA monitor data can be stated**: is pollution concentration at
+the monitor during reported times
+significantly different from the concentration at
+the monitor at missing times? Is the difference greater than random variations?
 
-- Implies an Abadie et al. 2011 style permutation inference test for each EPA monitor $i$.
+- Implies an permutation inference test for each EPA monitor $i$.
 
@@
 
-- Calculate missingness bias and algorithm bias for the EPA monitor
+```{r Create bar plots, eval=F, include=F}
+library(RColorBrewer)
 
-- Pick PA monitor $j\in J_i$, temporarily remove data for original hours missing from EPA monitor ($\mm_{i,h,d}$).
+# create dummy data
+b = rgb(0.7,0.7,0.7)
+r = rgb(0.8,0.1,0.1)
+mm = "Missing"
+nn = "Reported"
+data1 <- data.frame(
+  hour=1:24,
+  value=sapply(1:24, function (x) rnorm(1, 20-0.05*(x-14)^2, 2)),
+  color=c(rep(b,15), rep(r,5), rep(b, 24-15-5)),
+  missing=c(rep(nn,15), rep(mm,5), rep(nn, 24-15-5))
+)
 
-- Construct similar placebo synthetic control for PA monitor $j$ (predict $\widehat{PA}_{j,t}$).
+data2 <- data.frame(
+  hour=1:24,
+  value=sapply(1:24, function (x) rnorm(1, 20-0.05*(x-14)^2, 2)),
+  color=c(rep(b,15), rep(r,5), rep(b, 24-15-5)),
+  missing=sample(c())  c(rep(nn,15), rep(mm,5), rep(nn, 24-15-5))
+)
 
-- Calculate missingness bias and algorithm bias for the PA monitor: $Bias_{j,h,d}$
+# Uniform color
+ggplot(data) +
+  geom_col(aes(x=hour, y=value, fill=missing)) +
+  scale_fill_manual(values = c(r,b)) +
+  xlab("Hour of day") +
+  ylab("PM2.5 Pollution") +
+  ggtitle("EPA Monitor") +
+  theme(text = element_text(size=20))
+barplot(height=data$value, names=data$hour, col=data$color, cex.names = 1.4,
+        cex.axis=1.5, cex=1.5, xlab="hour", ylab="PM2.5 Pollution", main="EPA Monitor")
+
+```
+
+- Calculate missingness bias and algorithm bias for the $i^{th}$ EPA monitor using $J_i$ PA monitors with $n_i$ missing hour observations.
+
+- Pick PA monitor $j\in J_i$, temporarily remove random $n_i$ hour observations.
+
+- Construct placebo synthetic control for PA monitor $j$ and predict $\widehat{PA}_{j,t}$.
+
+- Calculate missingness bias and algorithm bias for PA monitor $j$: $Bias_{j,h,d}$
 
@@
 # Extensions
+
+- Use PurpleAir data to create population-weighted pollution measure $\implies$ counterfactual attainment.
 

### 2021-12-11 paper/5min-presentation/feedback/shreya_model.Rmd
addign shreya s feedback

---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

```

Assume $X_{m,y}$ is AR(1) + random normal distribution
```{r}
library(tidyverse)
library(magrittr)
set.seed(1)
# Define DGP parameters
months = 12
years = 2
households = 4
starting_x = 20
rho = 0.8
beta0 = 10
beta1 = 1  # for every dollar of subsidy, we want one more unit of gas consumed
# Generate household, month, year shocks (abosorbed by gamma fixed effects)
e_h = comprehenr::to_vec(for(h in 1:(households)) rnorm(1))
e_m = comprehenr::to_vec(for(h in 1:(months)) rnorm(1, mean=0, sd=2))
e_y = comprehenr::to_vec(for(h in 1:(years)) rnorm(1, mean=-2, sd=2))
# Generate the AR(1) subsidies
X = arima.sim(model = list(order = c(1, 0, 0), ar = rho), n = months*years, sd=6) + 20
plot(X)
# Create dataframe
df = data.frame(h = factor(comprehenr::to_vec(for (h in 1:households) for (y in 1:years) for (m in 1:months) h)),
                y = factor(comprehenr::to_vec(for (h in 1:households) for (y in 1:years) for (m in 1:months) y)),
                m = factor(comprehenr::to_vec(for (h in 1:households) for (y in 1:years) for (m in 1:months) m)),
                X = comprehenr::to_vec(for (h in 1:households) X),
                e_h = rep(e_h, each = months*years),
                e_y = rep(e_y, each = months, times=households),
                e_m = rep(e_m, times = households*years),
                e_hmy = rnorm(n = households*months*years, mean = 0, sd = 0.5)) %>%
    mutate(ym = paste0(y, '-', format(m, width=2)),
           # Generate Outcomes Y
           Y = beta0 + beta1*X + e_h + e_m + e_y + e_hmy)

df %>%
    ggplot() + geom_point(aes(x=X, y=Y, color = ym))

# Regression
df %>%
    lm(Y ~ X + h + m + y, data=.)
e_h
e_m
e_y
```

### 2021-12-15 paper/5min-presentation/5min-presentation.Rmd
Made some chunks all appear at the same time.

@@
 classoption: "aspectratio=169"
-knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'acWatt_purpleair_5min-pres_2021-12-02.pdf')) })
+knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'acWatt_purpleair_5min-pres_2021-12-15.pdf')) })
 ---
@@
 
-- The Clean Air Act (1970) established National Ambient Air Quality Standards
+\begin{itemize}
+\item<2-> The Clean Air Act (1970) established National Ambient Air Quality Standards
 (NAAQS) for US counties
-
-- Either "attainment" or "non-attainment", \color{red}large penalties and costly restrictions
-
-- Minimum requirement of 75\% of readings, per quarter
-
-- Air quality can change quickly
-
-- Monitor shutoffs are common
+\item<2-> Either "attainment" or "non-attainment", \color{red}large penalties and costly restrictions
+\item<2-> Minimum requirement of 75\% of readings, per quarter
+\item<2-> Air quality can change quickly
+\item<2-> Monitor shutoffs are common
+\end{itemize}
 
@@
 ## Research Questions
-
-- How biased are EPA monitor-based measures of local air quality?
+\begin{itemize}
+\item<3-> How biased are EPA monitor-based measures of local air quality?
+\item<3-> Does this bias significantly change NAAQS attainment status?
+\end{itemize}
 <!-- - How biased are self-reported EPA pollution estimates? -->
 <!-- - How biased is *missing* air pollution data from self-reporting US EPA monitors? -->
-
-- Does this bias significantly change NAAQS attainment status?
 
@@
 
-**Satellite Data:**
+\uncover<2->{**Satellite Data:**}
+\begin{itemize}
+\setlength{\itemsep}{0em}
+\item<2-> Grainger et al. 2017
+\item<2-> Sullivan, Krupnick 2018
+\item<2-> Fowlie, Rubin, Walker 2019
+\item<2-> Zou 2021
+\end{itemize}
 
-- Grainger et al. 2017
-- Sullivan, Krupnick 2018
-- Fowlie, Rubin, Walker 2019
-- Zou 2021
-
-**Location Pollution Alerts:**
-
-- Mu, Rubin, Zou 2021
+\uncover<3->{**Location Pollution Alerts:**
+\begin{itemize}
+\setlength{\itemsep}{0em}
+\item<3-> Mu, Rubin, Zou 2021
+\end{itemize}
 
@@
 
-## This project
-- Using new consumer-based pollution monitors (PurpleAir).
+\uncover<4->{## This project}
+\begin{itemize}
+%\setlength{\itemsep}{0em}
+\item<4-> Using new consumer-based pollution monitors (PurpleAir).
+\item<4-> Avoids using satellite estimates (has been shown to have significant error).
+\item<4-> Focus on EPA pollution data that is missing *in time*, specifically PM2.5
+\end{itemize}
 
-- Avoids using satellite estimates (has been shown to have significant error).
-
-- Focus on EPA pollution data that is missing *in time*, specifically PM2.5
 <!-- - Understanding the distribution of bias in missing California air pollution data using consumer-based pollution -->
@@
 
-load("/media/a/E/Programming/github/are219/data/epa/epa_monitors/epa_monitors/data_from_api/88101/api_request_daily_88101_2016.Rdata")
+# The below data files are only on Aaron's desktop. Need to create another data submodule to host data
+# or host on s3 bucket and download from there
+load("../../data/epa/epa_monitors/epa_monitors/data_from_api/88101/api_request_daily_88101_2016.Rdata")
 df = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88101, NAAQS_used='yes')
-load("/media/a/E/Programming/github/are219/data/epa/epa_monitors/epa_monitors/data_from_api/88501/api_request_daily_88501_2016.Rdata")
+load("../../data/epa/epa_monitors/epa_monitors/data_from_api/88501/api_request_daily_88501_2016.Rdata")
 df2 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88501, NAAQS_used='no')
-load("/media/a/E/Programming/github/are219/data/epa/epa_monitors/epa_monitors/data_from_api/88502/api_request_daily_88502_2016.Rdata")
+load("../../data/epa/epa_monitors/epa_monitors/data_from_api/88502/api_request_daily_88502_2016.Rdata")
 df3 = yr_dt %>% distinct(latitude, longitude, .keep_all = TRUE) %>% mutate(type=88502, NAAQS_used='no')
@@
 ```{r get PA sensor locations all, include=F}
-result = jsonlite::fromJSON("/media/a/E/Programming/github/are219/data/test_data/manual_webpage_download_USmap_2021-11-29.txt")
+result = jsonlite::fromJSON("../../data/test_data/manual_webpage_download_USmap_2021-11-29.txt")
 library(zoo)
@@
 \begin{itemize}%[itemsep=1.5em]
-\item \textbf{Estimate missing pollution observations:} Use PurpleAir data to predict EPA pollution at missing times
-\item \textbf{Estimate bias of reported EPA pollution:} difference between predicted pollution at missing times and reported pollution at nonmissing times.
-\item \textbf{Estimate counties' counterfactual attainment status:} Include estimated missing pollution data.
+\item<2> \textbf{Estimate missing pollution observations:} Use PurpleAir data to predict EPA pollution at missing times
+\item<2> \textbf{Estimate bias of reported EPA pollution:} difference between predicted pollution at missing times and reported pollution at nonmissing times.
+\item<2> \textbf{Estimate counties' counterfactual attainment status:} Include estimated missing pollution data.
 \end{itemize}

### 2021-12-17 paper/ongoing_progress/winter_break_plan.Rmd
Added winterbreak plan for meeting with MF on 2021-12-17

---
title: "Estimation and Implications of Bias in EPA Pollution Measurement"
author: Aaron C Watt
date: December 2, 2021
output:
  pdf_document:
    toc: true
    toc_depth: 2
    includes:
      in_header: preamble.tex
classoption: "12pt"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'acWatt_SYP_winterbreakplan_2021-12-17.pdf')) })
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tinytex)
# Had to run tinytex::install_tinytex() to use XeLaTeX if using YGbeamer.sty
```

```{r packages, include=F}
library(tidyverse)
library(docstring)
library(ggplot2)
# library(mapdata)
# Requires rgdal package to be installed, requires gdal to be installed on computer
# Ubuntu 20.04: sudo aptitude install gdal-bin proj-bin libgdal-dev libproj-dev
```

# Agenda
- communication over break and next term
- status of project
- winter break next steps
- discuss statistical analysis

# Communication
- How far in advance of our meetings
  - send updated materials 2 days in advance
- How often should/can we meet
  - 2 weeks as baseline, scheduling more if things come up
- Email vs. zoom calls
  - zoom preferred

# Status of project
- Have tested downloading full date range of PurpleAir data (python code ready)
- Have tested uploading files to S3 bucket (python code ready)

# Winter break plans
## Data
- Download hourly PA PM2.5 data from all US sensors (just in range of EPA monitors?)
- Download hours EPA PM2.5 data from all 88101 sensors (if more granular than spreadsheets)
- Add to S3 bucket

## Analysis
- Develop understanding of regulatory process (going from data to attainment status)
  - Make "pseudo design values" using my data (just using thresholds, don't worry about perfectly replicating the chosen attainment status/design values, but could have a map/stats comparing real to pseudo)
- Write code to implement attainment status given hourly data and check against actual attainment status

### Datasets
- Hourly PM2.5 Pollution data from California EPA pollution monitors (2015-2020)
- Hourly PM2.5 Pollution data from California PurpleAir sensors, hourly averages taken

## Deadlines
- Dec. 24: PurpleAir is downloading hourly averages on 8 AWS tiny linux instances, sending CSVs to S3 bucket
- Jan. 5: Proof of concept of statistical analysis for 2 EPA sensors 
  - Send Meredith update
- Jan. 14: Draft paper sections (esp. intro and data section)
- Jan. 21: Draft of all paper sections
- Feb 18: Paper due
- Feb 18: Presentation + other paper reviews
- Apr 11: Final paper (incorporating feedback)

# Research Questions

- How biased are EPA monitor-based measures of local air quality?
<!-- - How biased are self-reported EPA pollution estimates? -->
<!-- - How biased is *missing* air pollution data from self-reporting US EPA monitors? -->

- Does this bias significantly change NAAQS attainment status?

## Particulars of this project
- Using PurpleAir data on PM2.5

- Avoids using satellite estimates (but has drawbacks of it's own -- can document the extent of these).

- Focus on EPA pollution data that is missing *in time* (starting with every day, hourly sensors).

- Extensions to summer:
  - Missingness in space (placement of EPA monitors)
  - How biased are current measures of PM2.5 pollution faced by residents (population-weighted pollution)
  - Optimal regulation design: Given imperfect/incomplete reporting, spatial nature of weather, and the range of quality and price of different pollution monitors, what is the optimal regulation mechanism? At least, what is the optimal distribution of cheap vs. expensive monitors given budget constraints?

\newpage
# Models

\begin{enumerate}%[itemsep=1.5em]
\item \textbf{Estimate missing pollution observations:} Use PurpleAir data to predict EPA pollution at missing times
\item \textbf{Estimate bias of reported EPA pollution:} difference between predicted pollution at missing times and reported pollution at nonmissing times.
\item \textbf{Estimate counties' counterfactual attainment status:} Include estimated missing pollution data.
\end{enumerate}

## Model 1: Predictive model of each EPA monitor PM2.5 pollution

**Two models**: 
- one with sensor-specific weights/regression ($\gamma_{i,j,k}$)
- one with pooled sensor weights/regression ($\gamma_{j,k}$)

\[
EPA_{i,t} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{i,j,k} 
  PA_{j,t} \cdot Winddir_{i,t,k} + u_{i,t}
\]

\begin{itemize}
\item Analysis done at the month and quarter level; suppressing that subscript.
\item $t$ is a unique hour within a given month or quarter.
\item EPA monitor $i$ at time $t$ reads PM2.5 pollution $EPA_{i,t}$.
\item For each EPA monitor $i$, there are $J_i$ Purple Air monitors within a 10-mile radius.
\item Purple Air monitor $j\in J_i$ at time $t$ reads PM2.5 pollution $PA_{j,t}$.
\item $Winddir_{i,t,k} $ is a wind direction indicator; 1 if the prevailing wind near station $i$ at time $t$ is in the $k^{th}$ bucket (of 8 buckets).
\end{itemize}

**Extreme Event from EPA data:** indicator for imputed data when there was some nearby extreme event (eg, car fire), only use real data not imputed to create $gamma$ weights.

<!-- [insert diagram of two EPA monitors with PA monitors surrounding them] -->

<!-- [insert diagram of one EPA monitor and surrounding PA monitors, with wind directions] -->

\newpage
## Model 2: Hour-by-Day-of-week Bias of Missing EPA Monitor Pollution Data
\def\mm{\mathcal{M}}
\def\nn{\mathcal{N}}
\vspace{-1em}
\begin{align*}
\intertext{\textbf{Missingness Bias:}}
Bias_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t} 
- \frac{1}{|\mm_{i,h,d}|}\sum\limits_{t\in \mm_{i,h,d}}\widehat{EPA}_{i,t}\\
\text{where } \mm_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Missing}\};\\
  \nn_{i,h,d}&= \{t: t\text{ is at hour $h$ and day $d$ and $EPA_{i,t}$ is Non-missing}\} \\
\intertext{We can also define the \textbf{algorithm bias}     as the Hour-by-Day-of-week prediction error}
\widetilde{Bias}_{i,h,d} &= \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}EPA_{i,t} 
- \frac{1}{|\nn_{i,h,d}|}\sum\limits_{t\in \nn_{i,h,d}}\widehat{EPA}_{i,t} \\
\intertext{We can also define $Bias_{j,h,d}$ and $\widetilde{Bias}_{j,h,d}$ for PA monitor $j$ (we'll come back to this).}
\end{align*}

- ML correction: include region, humidity, temperature, etc of PA monitors within radius.
- Come up with my own correction equation? How does this compare to ML correction?
- Compare both types of bias from uncorrected predictions, EPA corrected predictions, ML corrected predictions'

## Model 3: County Attainment Status

\begin{align*}
Attain^{annual}_c &= 1 \text{ if \textbf{reported} annual average PM2.5 below 15.0 $\mu$g/m$^3$}^*\\[1em]
Attain^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{reported} daily average PM2.5 below 35 $\mu$g/m$^3$}^*\\[1em]
\widehat{Attain}^{annual}_c &= 1 \text{ if \textbf{predicted} annual average PM2.5 below 15.0 $\mu$g/m$^3$}^*\\[1em]
\widehat{Attain}^{daily}_c &= 1 \text{ if 98$^{th}$ percentile of \textbf{predicted} daily average PM2.5 below 35 $\mu$g/m$^3$}^*
\end{align*}
$^*$averaged over 3 years in NAAQS standard.

\newpage
# Proposed Statistical Test

- The $J_i$ group of PurpleAir sensors is (in a sense) a synthetic control for the EPA sensor $i$.

- **The question of bias in EPA monitor data can be stated**: is pollution concentration at
the monitor during reported times
significantly different from the concentration at
the monitor at missing times? Is the difference greater than random variations?

- Implies an permutation inference test for each EPA monitor $i$.

- Calculate missingness bias and algorithm bias for the $i^{th}$ EPA monitor using $J_i$ PA monitors with $n_i$ missing hour observations.

- Pick PA monitor $j\in J_i$, temporarily remove random $n_i$ hour observations.

- Construct placebo synthetic control for PA monitor $j$ and predict $\widehat{PA}_{j,t}$.

- Calculate missingness bias and algorithm bias for PA monitor $j$: $Bias_{j,h,d}$

- Repeat for all PA sensors.

- **Graphical test**: For EPA sensor $i$, compare graph of $Bias_{i,h,d}$ to placebo $Bias_{j,h,d}$ for $j\in J_i$.

- **Permutation inference p-value**:

  - Calculate sum of squared missingness bias and sum of squared algorithm bias for EPA sensor $i$ and PA sensors $j\in J_i$.
  \vspace{1em}
  - Ratio$_k$ = sum of squared missingness bias / sum of squared algorithm bias
  \vspace{1em}
  - p-value = $\frac{\text{\# of PA sensors in $i$'s radius with Ratio$_j$ larger than Ratio$_i$}}{\text{\# of PA sensors in $i$'s radius}}$

<!-- But how does this account for poor prediction of the monitor?
Because the denominator is larger if poor prediction in non-missing observations -->

**Welfare implications:** use cheap comparison from Fowlie et al. 2019: look at counties that would be non-attainment if missing data were included, calibrate model (mortality, etc) on those specific counties and predict \# of lives that would be saved, etc

# Extensions

- Use PurpleAir data to create population-weighted pollution measure $\implies$ counterfactual attainment.

- Welfare analysis based on attainment status changes and required reductions in pollution.

- Comparing county population-weighted PM2.5 pollution to EPA sensors to estimate location-based bias.

\newpage
# Appendix A: PurpleAir monitor correction factor

![PurpleAir correction equation for EPA monitor PM2.5 (RH = relative humidity, also measured by PA monitor)](../../docs/purpleair/PA_correction_equation.png){width=50%}

Source: \url{https://www.epa.gov/air-sensor-toolbox/technical-approaches-sensor-data-airnow-fire-and-smoke-map}

# Appendix B: Data Plan

### Data Wearhouse
- AWS Linux Cassandra database (noSQL columnar, designed for large queries of columns)
- Python pushes and pulls

### Storage costs
- ~ 30,000 sensors, 50 variables, 2 minute intervals, 5 years of data = 107.55 Terabytes
- Depending on method of storage: \$2,900 - \$13,300 per month
- Only storing hourly means and SD: \$4 - \$15 per month

### 2021-12-28 paper/5min-presentation/5min-presentation.Rmd
Updated prediction model to include i subscript (so each set of gamma weights is specific to the i'th monitor)
removing some redundancy of filenames.

@@
 
-ggsave('monitor_map_2015-2021.png', width = 6, height = 4)
+p_map = 'monitor_map_2015-2021.png'
+ggsave(map_filename, width = 6, height = 4)
 
@@
 p_epa = "../../docs/epa/air-monitoring-site_border.jpg"
-p_map = 'monitor_map_2015-2021.png'
 library(png)
@@
 library(gridExtra)
-img_map = rasterGrob(as.raster(readPNG(p_map)), interpolate = FALSE)
+img_map = rasterGrob(as.raster(readPNG(map_filename)), interpolate = FALSE)
 img1 <-  rasterGrob(as.raster(readPNG(p_pa)), interpolate = FALSE)
@@
 \[
-\alert<2>{EPA_{i,t}} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{j,k} 
+\alert<2>{EPA_{i,t}} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{i,j,k} 
 \alert<3>{PA_{j,t}} \cdot \alert<4>{Winddir_{i,t,k}} + u_{i,t}
@@
 \begin{itemize}
-\item<1-> Analysis done at the month and quarter level; suppressing that subscript.
+\item<1-> Each EPA monitor $i$ has it's own set of weights for the PA sensors around it.
+\item<1-> Analysis done at the month and quarter level; suppressing quarter subscript.
 \item<1-> $t$ is a unique hour within a given month or quarter.

### 2021-12-28 acwatt_syp_code/build/purpleair_download.py
Making sensor week download function only for one sensor at a time. Adding timezone convertion to ThingsSpeak API call, getting timezone from PurpleAir API call.

@@ -3,17 +3,22 @@
 """Functions to download purple air data via their API"""
 
 # Built-in Imports
-from datetime import datetime, timedelta
+import datetime as dt
+# from datetime import datetime, timedelta
 from dateutil.relativedelta import relativedelta
 import json
 import pandas as pd
 import requests
 from io import StringIO
+from typing import Optional, Union
+from pytz import timezone as tz
 
 # Third-party Imports
 from ratelimiter import RateLimiter
 import purpleair.network as pan
 from purpleair.network import SensorList, Sensor
+from timezonefinder import TimezoneFinder
+
 # Local imports
 from ..utils.config import PATHS, PA
 
@@ -53,14 +58,17 @@ def pa_request(sensor_id):
     # api_key = PA.read_key  # TODO: use this
     api_key = '5498FF4F-1642-11EC-BAD6-42010A800017'
     url = f'https://api.purpleair.com/v1/sensors/{sensor_id}'
-    fields = 'name, latitude, longitude, pm2.5,primary_id_a, primary_key_a, secondary_id_a, secondary_key_a, primary_id_b, primary_key_b, secondary_id_b, secondary_key_b'
-    query = {'api_key': api_key, 'fields': fields.replace(' ','')}
+    fields = 'name, date_created, last_modified, latitude, longitude, position_rating, ' \
+             'pm2.5, primary_id_a, primary_key_a, secondary_id_a, secondary_key_a, ' \
+             'primary_id_b, primary_key_b, secondary_id_b, secondary_key_b'
+    query = {'api_key': api_key, 'fields': fields.replace(' ', '')}
     response = requests.get(url, params=query)
     return response.json()
 
 
 @RateLimiter(max_calls=1, period=1)
-def ts_requst(channel_id, start_date, api_key, end_date=None):
+def ts_requst(channel_id, start_date, api_key,
+              end_date=None, average=None, timezone=None):
     """Return dataframe of REST json data response for thingsspeak request.
 
     channel_id = id of the device to get data from
@@ -75,9 +83,13 @@ def ts_requst(channel_id, start_date, api_key, end_date=None):
     query = {'api_key': api_key,
              'start': start_date.strftime("%Y-%m-%d%%20%H:%M:%S"),
              'end': end_date.strftime("%Y-%m-%d%%20%H:%M:%S")}
+    if average is not None:
+        query['average'] = average
+    if timezone is not None:
+        query['timezone'] = timezone
[…]

### 2021-12-28 acwatt_syp_code/build/purpleair_download.py
remove 'useful' filtering from PurpleAir.Network function because it filters out any sensors that may have good historical data.

@@ -41,7 +41,7 @@ def df_sensors():
     """Return list of sensors that have valid data."""
     p = pan.SensorList()  # Initialized 22,877 sensors!22,836
     # Other sensor filters include 'outside', 'all', 'useful', 'family', and 'no_child'
-    df = p.to_dataframe(sensor_filter='useful', channel='parent')
+    df = p.to_dataframe(sensor_filter='outside', channel='parent')
     # TODO: Cache and check if this has already been downloaded in past 7 days
     #   if not, redownload and update the cache
     return df

### 2021-12-28 acwatt_syp_code/build/purpleair_download.py
Using api key stored in config.py

@@ -43,7 +43,7 @@ def df_sensors():
     # Other sensor filters include 'outside', 'all', 'useful', 'family', and 'no_child'
     df = p.to_dataframe(sensor_filter='outside', channel='parent')
     # TODO: Cache and check if this has already been downloaded in past 7 days
-    #   if not, redownload and update the cache
+    #       if not, redownload and update the cache
     return df
 
 
@@ -55,8 +55,7 @@ def rest_csv_to_df(url, query):
 
 
 def pa_request(sensor_id):
-    # api_key = PA.read_key  # TODO: use this
-    api_key = '5498FF4F-1642-11EC-BAD6-42010A800017'
+    api_key = PA.read_key
     url = f'https://api.purpleair.com/v1/sensors/{sensor_id}'
     fields = 'name, date_created, last_modified, latitude, longitude, position_rating, ' \
              'pm2.5, primary_id_a, primary_key_a, secondary_id_a, secondary_key_a, ' \
@@ -115,8 +114,7 @@ def ts_example():
 
 def dl_sensor_list():
     """Download sensor metadata for california"""
-    # api_key = PA.read_key  # TODO: use this
-    api_key = '5498FF4F-1642-11EC-BAD6-42010A800017'
+    api_key = PA.read_key
     url = "https://api.purpleair.com/v1/sensors"
     fields = "sensor_index,date_created,latitude,longitude,altitude,position_rating,private,location_type,confidence_auto,channel_state,channel_flags,pm2.5,pm2.5_a,pm2.5_b,pm2.5_24hour,pm2.5_1week,humidity,temperature,pressure,voc,ozone1"
     query = {'api_key': api_key, 'fields': fields.replace(" ", ""),

### 2021-12-28 run.py
Added timezone fix and checking hourly average.

@@ -3,6 +3,7 @@
 """Module Docstring"""
 
 # Built-in Imports
+import pandas as pd
 # Third-party Imports
 # Local Imports
 import acwatt_syp_code.analyze.maps as am
@@ -11,7 +12,18 @@ from acwatt_syp_code.utils.config import PATHS
 
 if __name__ == "__main__":
     # Get available sensor locations from  from purpleair.network.SensorList API wrapper
-    sensor_df = pad.load_current_sensor_data()
+    # sensor_df = pad.load_current_sensor_data()
+    # Get available sensor locations from thingsspeak API directly
+    sensors = [25999] #, 22751]
+    df_pa = pad.pa_request(sensors[0])
+    df_ts = pad.ts_example()
+    dfs = [pad.dl_sensor_weeks(id_, date_start='2021-12-25') for id_ in sensors]
+    df_total = pd.concat(dfs).sort_values('created_at')
+    df_total['datetime'] = pd.to_datetime(df_total['created_at'])
+    df_total.index = df_total['datetime']
+    df_hourly = df_total.resample('H').mean()
+    dfs = [pad.dl_sensor_weeks(id_, date_start='2021-12-25', average=60) for id_ in sensors]
+    df_total2 = pd.concat(dfs).sort_values('created_at')
     # Make sensor plots
     # am.make_all_sensor_maps(sensor_df)
     # Save sensor dataframe

### 2022-01-11 paper/5min-presentation/5min-presentation.Rmd
Day's end commit: lambda and S3 documentation, lab notes, reorganizing docs.

@@
          location_type == 0,  # outdoor sensor
-         channel_state != 0)  # has at least one PM sensor
+         channel_state == 3)  # has 2 PM sensors (so we can QA check)
+
+write.csv(pa_data2,
+          "../../data/test_data/manual_webpage_download_USmap_2021-11-29_filtered.csv",
+          row.names=F)
 ```

### 2022-01-11 acwatt_syp_code/build/aws/lambda_services.py
Initalizing lambda scripts.

@@ -1,0 +1,231 @@
+# Copyright Amazon.com, Inc. or its affiliates. All Rights Reserved.
+# SPDX-License-Identifier: Apache-2.0
+# Modified by Aaron Watt, Jan 2022
+
+"""
+PURPOSE:
+Create a lambda function service call to download a sensor's historical data.
+
+Uses the AWS SDK for Python (Boto3) to create an AWS Lambda function from `lambda_download_script.py,
+invoke it for a specific sensor ID, and delete it.
+"""
+
+import io
+import json
+import logging
+import random
+import time
+import zipfile
+import boto3
+from botocore.exceptions import ClientError
+
+logger = logging.getLogger(__name__)
+
+
+def exponential_retry(func, error_code, *func_args, **func_kwargs):
+    """
+    Retries the specified function with a simple exponential backoff algorithm.
+    This is necessary when AWS is not yet ready to perform an action because all
+    resources have not been fully deployed.
+
+    :param func: The function to retry.
+    :param error_code: The error code to retry. Other errors are raised again.
+    :param func_args: The positional arguments to pass to the function.
+    :param func_kwargs: The keyword arguments to pass to the function.
+    :return: The return value of the retried function.
+    """
+    sleepy_time = 1
+    func_return = None
+    while sleepy_time < 33 and func_return is None:
+        try:
+            func_return = func(*func_args, **func_kwargs)
+            logger.info("Ran %s, got %s.", func.__name__, func_return)
+        except ClientError as error:
+            if error.response['Error']['Code'] == error_code:
+                print(f"Sleeping for {sleepy_time} to give AWS time to "
+                      f"connect resources.")
+                time.sleep(sleepy_time)
+                sleepy_time = sleepy_time*2
+            else:
+                raise
+    return func_return
+
+
+def create_lambda_deployment_package(function_file_name):
+    """
+    Creates a Lambda deployment package in ZIP format in an in-memory buffer. This
+    buffer can be passed directly to AWS Lambda when creating the function.
+
+    :param function_file_name: The name of the file that contains the Lambda handler
+                               function.
+    :return: The deployment package.
+    """
+    buffer = io.BytesIO()
+    with zipfile.ZipFile(buffer, 'w') as zipped:
+        zipped.write(function_file_name)
+    buffer.seek(0)
+    return buffer.read()
+
+
+def create_iam_role_for_lambda(iam_resource, iam_role_name):
+    """
[…]

### 2022-01-11 acwatt_syp_code/build/aws/lambda_download_script.py
Adding pseudo code to get started.

@@ -1,0 +1,114 @@
+# Copyright Amazon.com, Inc. or its affiliates. All Rights Reserved.
+# SPDX-License-Identifier: Apache-2.0
+# Modified by Aaron Watt, Jan 2022
+
+"""
+PURPOSE:
+Writes IP address used during lambda function execution to S3 bucket CSV.
+"""
+
+import logging
+import pandas as pd
+import datetime as dt
+
+logger = logging.getLogger()
+logger.setLevel(logging.INFO)
+
+# Define a list of Python lambda functions that are called by this AWS Lambda function.
+
+
+def get_all_week_dates(date_beg, date_end):
+    # Get
+    date_start = dt.datetime.strptime(date_beg, '%Y-%m-%d')
+    date_end = dt.datetime.strptime(date_end, '%Y-%m-%d')
+    date_list = pd.date_range(date_start, dt.datetime.today(), freq='w')
+    pass
+
+
+def download_sensor_week(id_, week_beg, week_end):
+    # construct REST call to ThingsSpeak API
+
+    # Download JSON response
+
+    # Convert JSON to dataframe
+
+    pass
+
+
+def write_df_to_memcsv(df):
+    # Create CSV of df in memory
+
+    pass
+
+
+def upload_s3_csv(memcsv):
+    # Write fileobject to S3 bucket
+    pass
+
+
+def download_sensor(id_, date_beg, date_end):
+    # Get sensor's creation date and last modified date
+    date_beg, date_end = get_begin_end_dates(id_)
+
+    # Get beginning and end dates of all weeks between creation and last modified
+    weeks_tuple_list = get_all_week_dates(date_beg, date_end)
+
+    # For each week
+    df_list = []
+    for week_beg, week_end in weeks_tuple_list:
+        # Download the Purple Air sensor data to dataframe
+        df_list.append(download_sensor_week(id_, week_beg, week_end))
+
+    # Write df to CSV in memory
+    df = pd.concat(df_list)
+    write_df_to_memcsv(df)
+
+    # Write S3 file
+    pass
+
+
+def ip_test(id_):
+    # Get IP address
+    # Create df of IP address
+    # Create memCSV of df
+    # Write CSV to S3 bucket
+    # Return success or error code
+    pass
+
+
+def lambda_ip_s3_writer(params, bucket_dir = 'sensor_csvs'):
+    """
+    Accepts an action and a number, performs the specified action on the number,
+    and returns the result.
+
[…]

### 2022-01-11 acwatt_syp_code/build/aws/lambda_services.py
Edited lambda creation script for a demo run.

@@ -70,15 +70,15 @@ def create_lambda_deployment_package(function_file_name):
 def create_iam_role_for_lambda(iam_resource, iam_role_name):
     """
     Creates an AWS Identity and Access Management (IAM) role that grants the
-    AWS Lambda function basic permission to run. If a role with the specified
-    name already exists, it is used for the demo.
+    AWS Lambda function basic permission to run S3 bucket reads and writes.
+    If a role with the specified name already exists, it is used.
 
     :param iam_resource: The Boto3 IAM resource object.
     :param iam_role_name: The name of the role to create.
     :return: The newly created role.
     """
     lambda_assume_role_policy = {
-        'Version': '2012-10-17',
+        'Version': '2022-01-01',
         'Statement': [
             {
                 'Effect': 'Allow',
@@ -89,7 +89,7 @@ def create_iam_role_for_lambda(iam_resource, iam_role_name):
             }
         ]
     }
-    policy_arn = 'arn:aws:iam::aws:policy/service-role/AWSLambdaBasicExecutionRole'
+    policy_arn = 'arn:aws:iam::aws:policy/service-role/AmazonS3ObjectLambdaExecutionRolePolicy'
 
     try:
         role = iam_resource.create_role(
@@ -130,8 +130,8 @@ def deploy_lambda_function(
     try:
         response = lambda_client.create_function(
             FunctionName=function_name,
-            Description="AWS Lambda demo",
-            Runtime='python3.8',
+            Description="AWS Lambda demo for S3",
+            Runtime='python3.7',
             Role=iam_role.arn,
             Handler=handler_name,
             Code={'ZipFile': deployment_package},
@@ -187,13 +187,13 @@ def usage_demo():
     """
     logging.basicConfig(level=logging.INFO, format='%(levelname)s: %(message)s')
     print('-'*88)
-    print("Welcome to the AWS Lambda basics demo.")
+    print("Welcome to the AWS Lambda demo for S3 bucket writes.")
     print('-'*88)
 
-    lambda_function_filename = 'lambda_handler_basic.py'
-    lambda_handler_name = 'lambda_handler_basic.lambda_handler'
+    lambda_function_filename = 'lambda_download_script.py'
+    lambda_handler_name = 'lambda_download_script.lambda_ip_s3_writer'
     lambda_role_name = 'demo-lambda-role'
-    lambda_function_name = 'demo-lambda-function'
+    lambda_function_name = 'demo-lambda-function-s3'
 
[…]

### 2022-01-11 acwatt_syp_code/build/aws/lambda_download_script.py
adding notes

@@ -109,6 +109,6 @@ def lambda_ip_s3_writer(params, bucket_dir = 'sensor_csvs'):
 """
 NOTES
 - test very simple IP address writer asap
-- 
+- need to determine how to make / break calls to get different IP addresses.
 
 """

### 2022-01-12 acwatt_syp_code/utils/config.py
Adding class for AWS settings to config

@@ -71,6 +71,13 @@ class PurpleAirSettings:
         self.url = 'https://api.purpleair.com/v1/sensors'
 
 
+class AWSSettings:
+    """Class to hold settings for Amazon AWS info."""
+    def __init__(self):
+        self.bucket_arn = 'arn:aws:s3:::purpleair-data'
+        self.access_key = "placeholder"
+
+
 # FUNCTIONS --------------------------
 
 

### 2022-01-12 acwatt_syp_code/build/aws/lambda_download_script.py
removing call to already removed function (for getting date_created, date_last_modified, because those can be sent during the lambda call)

@@ -47,9 +47,6 @@ def upload_s3_csv(memcsv):
 
 
 def download_sensor(id_, date_beg, date_end):
-    # Get sensor's creation date and last modified date
-    date_beg, date_end = get_begin_end_dates(id_)
-
     # Get beginning and end dates of all weeks between creation and last modified
     weeks_tuple_list = get_all_week_dates(date_beg, date_end)
 

### 2022-01-12 acwatt_syp_code/build/aws/lambda_download_script.py
Adding code for getting IP address and uploading file to S3

@@ -6,10 +6,16 @@
 PURPOSE:
 Writes IP address used during lambda function execution to S3 bucket CSV.
 """
-
+import io
 import logging
 import pandas as pd
 import datetime as dt
+import requests
+import json
+import os
+
+import boto3
+from botocore.exceptions import ClientError
 
 logger = logging.getLogger()
 logger.setLevel(logging.INFO)
@@ -64,10 +70,49 @@ def download_sensor(id_, date_beg, date_end):
     pass
 
 
+def get_ip():
+    """Return public IP address. From https://pytutorial.com/python-get-public-ip"""
+    endpoint = 'https://ipinfo.io/json'
+    response = requests.get(endpoint, verify = True)
+
+    if response.status_code != 200:
+        return f'Error status code: {response.status_code}'
+        exit()
+
+    data = response.json()
+    return data['ip']
+
+
+def upload_file(file_path, bucket, object_name=None):
+    """Upload a file to an S3 bucket
+
+    :param file_path: File to upload
+    :param bucket: Bucket to upload to
+    :param object_name: S3 object name. If not specified then file_name is used
+    :return: True if file was uploaded, else False
+    """
+    # If S3 object_name was not specified, use file_name
+    if object_name is None:
+        object_name = os.path.basename(file_path)
+
+    # Upload the file
+    s3_client = boto3.client('s3')
+    try:
+        response = s3_client.upload_file(file_path, bucket, object_name)
+    except ClientError as e:
+        logging.error(e)
+        return False
+    return True
+
+
 def ip_test(id_):
     # Get IP address
+    ip = get_ip()
     # Create df of IP address
-    # Create memCSV of df
+    df = pd.DataFrame({'ip_address': ip}, index=[id_])
+    # Save the df to a CSV that only exists in memory (not written to disk)
+    mem_csv = io.StringIO()
+    df.to_csv(mem_csv)
     # Write CSV to S3 bucket
     # Return success or error code
     pass

### 2022-01-12 acwatt_syp_code/build/aws/lambda_services.py
AWS settings instantiation in config

@@ -19,6 +19,9 @@ import zipfile
 import boto3
 from botocore.exceptions import ClientError
 
+# Local imports
+from ...utils.config import PATHS, AWS
+
 logger = logging.getLogger(__name__)
 
 

### 2022-01-12 acwatt_syp_code/utils/config.py

@@ -86,3 +86,4 @@ class AWSSettings:
 PATHS = Paths()
 GIS = GISSettings()
 PA = PurpleAirSettings()
+AWS = AWSSettings()

### 2022-01-12 acwatt_syp_code/build/purpleair_download.py
style

@@ -22,7 +22,6 @@ from timezonefinder import TimezoneFinder
 # Local imports
 from ..utils.config import PATHS, PA
 
-
 # If retrieving data from multiple sensors at once, please send a single request
 # rather than individual requests in succession.
 
@@ -205,8 +204,8 @@ def dl_sensor_weeks(sensor_id: Union[str, int, float] = None,
     # iterate through all different channels of data for sensor
     df_list = []
     print(f'\nDownloading for sensor {sensor_id}')
-    for i in reversed(range(len(date_list)-1)):
-        start, end = date_list[i: i+2].date
+    for i in reversed(range(len(date_list) - 1)):
+        start, end = date_list[i: i + 2].date
         # offset = daylight_savings_offset(start, timezone)
         print(f'. . Date Range: {start} - {end}')
         for channel in ['a', 'b']:
@@ -330,11 +329,11 @@ def sensorid_to_df(sensor_id):
         suffix = {'parent': '_1', 'child': '_2'}[channel]
         for k1 in dict_[channel]:
             for k2 in dict_[channel][k1]:
-                row_dict[k2+suffix] = [dict_[channel][k1][k2]]
+                row_dict[k2 + suffix] = [dict_[channel][k1][k2]]
     # Add created datetime column
     created_unix = dict_['parent']['diagnostic']['created']
     return (pd.DataFrame.from_dict(row_dict)
-          .assign(created_date=datetime.utcfromtimestamp(created_unix)))
+            .assign(created_date=dt.utcfromtimestamp(created_unix)))
 
 
 def load_current_sensor_data(update_lookup=True):
@@ -355,7 +354,7 @@ def update_loc_lookup(df, output=False):
     except FileNotFoundError:
         loc_lookup = pd.DataFrame(columns=loc_cols)
     # Update lookup
-    today = datetime.today().date().strftime('%Y-%m-%d')
+    today = dt.today().date().strftime('%Y-%m-%d')
     loc_lookup = (loc_lookup.merge(df
                                    .reset_index()
                                    .filter(loc_cols, axis=1),

### 2022-01-12 acwatt_syp_code/build/aws/lambda_services.py
Adding conversion from unix time stamp, adding real unix times from data to test.

@@ -10,6 +10,7 @@ Uses the AWS SDK for Python (Boto3) to create an AWS Lambda function from `lambd
 invoke it for a specific sensor ID, and delete it.
 """
 
+import datetime as dt
 import io
 import json
 import logging
@@ -212,14 +213,19 @@ def usage_demo():
 
     print(f"Directly invoking function {lambda_function_name} a few times...")
     sensors = [25999, 26003, 26005, 26011, 26013]
-    for i in range(5):
+    date_created_list = [1632955574, 1632955612, 1632955594, 1446763462, 1632955644]
+    last_modified_list = [1635632829, 1634149424, 1634410114, 1633665195, 1635632829]
+    for id_, date_created, last_modified in zip(sensors, date_created_list, last_modified_list):
         # lambda_parms = {
         #     'number': random.randint(1, 100), 'action': random.choice(actions)
         # }
-        lambda_params = {'sensor_id': sensors[i]}
+        lambda_params = {'sensor_id': id_,
+                         'bucket_arn': AWS.bucket_arn,
+                         'date_created': dt.utcfromtimestamp(date_created),
+                         'last_modified': dt.utcfromtimestamp(last_modified)}
         response = invoke_lambda_function(
             lambda_client, lambda_function_name, lambda_params)
-        print(f"Downloading and saving of sensor {sensors[i]} resulted in "
+        print(f"Downloading and saving of sensor {id_} resulted in "
               f"{json.load(response['Payload'])}")
 
     for policy in iam_role.attached_policies.all():

### 2022-01-12 acwatt_syp_code/build/aws/lambda_services.py
notes for IP issues

@@ -206,6 +206,10 @@ def usage_demo():
           f"{lambda_handler_name} function in {lambda_function_filename}...")
     deployment_package = create_lambda_deployment_package(lambda_function_filename)
     iam_role = create_iam_role_for_lambda(iam_resource, lambda_role_name)
+    # Could use the  GetFunctionConfiguration State "Active" response to wait until function is ready to envoke
+    # while state != "Active" and timer < max_try_time: ...
+    # Could use PutFunctionConcurrency to set concurrency to 1 if there are IP issues
+    # put lambda_client instantiation inside for loop to create new for each sensor
     exponential_retry(
         deploy_lambda_function, 'InvalidParameterValueException',
         lambda_client, lambda_function_name, lambda_handler_name, iam_role,

### 2022-01-12 acwatt_syp_code/build/aws/lambda_download_script.py
final details on draft to upload IP csv to S3 using lambda

@@ -105,20 +105,20 @@ def upload_file(file_path, bucket, object_name=None):
     return True
 
 
-def ip_test(id_):
+def ip_test(id_, bucket):
     # Get IP address
     ip = get_ip()
     # Create df of IP address
     df = pd.DataFrame({'ip_address': ip}, index=[id_])
     # Save the df to a CSV that only exists in memory (not written to disk)
     mem_csv = io.StringIO()
-    df.to_csv(mem_csv)
-    # Write CSV to S3 bucket
-    # Return success or error code
-    pass
+    df.to_csv(mem_csv, index=False)
+    mem_csv.seek(0)  # need to set position of buffer back to begging before reading
+    # Write CSV to S3 bucket (True/False for success/failure)
+    return upload_file(mem_csv, bucket, object_name=f'{id_:06d}.csv')
 
 
-def lambda_ip_s3_writer(params, bucket_dir = 'sensor_csvs'):
+def lambda_ip_s3_writer(params, bucket_dir='sensor_csvs'):
     """
     Accepts an action and a number, performs the specified action on the number,
     and returns the result.
@@ -129,11 +129,13 @@ def lambda_ip_s3_writer(params, bucket_dir = 'sensor_csvs'):
     :return: The result of the specified action.
     """
     id_ = params['sensor_id']
-    created_date = params['']
+    bucket = params['bucket_arn']
+    created_date = params['date_created']
+    last_modified = params['last_modified']
     filename = bucket_dir + f'/{id_}.csv'
     logger.info(f'Sensor: {id_}')
 
-    result = ip_test(id_)
+    result = "Success" if ip_test(id_, bucket) else "Failure"
     logger.info(f'Result of uploading IP CSV: {result}')
 
     # result = download_sensor(id_, params['date_created'], params['date_end'])

### 2022-01-12 acwatt_syp_code/build/aws/lambda_services.py

@@ -196,7 +196,7 @@ def usage_demo():
 
     lambda_function_filename = 'lambda_download_script.py'
     lambda_handler_name = 'lambda_download_script.lambda_ip_s3_writer'
-    lambda_role_name = 'demo-lambda-role'
+    lambda_role_name = 'demo-lambda-role-S3-ip-upload'
     lambda_function_name = 'demo-lambda-function-s3'
 
     iam_resource = boto3.resource('iam')

### 2022-01-12 run.py

@@ -9,8 +9,10 @@ import pandas as pd
 import acwatt_syp_code.analyze.maps as am
 import acwatt_syp_code.build.purpleair_download as pad
 from acwatt_syp_code.utils.config import PATHS
+from acwatt_syp_code.build.aws import lambda_services
 
 if __name__ == "__main__":
+    lambda_services.usage_demo()
     # Get available sensor locations from  from purpleair.network.SensorList API wrapper
     # sensor_df = pad.load_current_sensor_data()
     # Get available sensor locations from thingsspeak API directly

### 2022-01-12 acwatt_syp_code/build/aws/lambda_services.py
implementing secure passwords keyring for AWS and PurpleAir keys

@@ -199,6 +199,11 @@ def usage_demo():
     lambda_role_name = 'demo-lambda-role-S3-ip-upload'
     lambda_function_name = 'demo-lambda-function-s3'
 
+    session = boto3.Session(
+        region_name=AWS.region,
+        aws_access_key_id=AWS.access_key,
+        aws_secret_access_key=AWS.secret_key
+    )
     iam_resource = boto3.resource('iam')
     lambda_client = boto3.client('lambda')
 

### 2022-01-12 acwatt_syp_code/utils/config.py

@@ -7,6 +7,8 @@
 
 # Built-in Imports
 from pathlib import Path
+import keyring
+import keyring.util.platform_ as keyring_platform
 
 
 # CLASSES --------------------------
@@ -66,16 +68,37 @@ class GISSettings:
 class PurpleAirSettings:
     """Class to hold settings for Purple Air API when downloading data."""
     def __init__(self):
-        self.read_key = '5498FF4F-1642-11EC-BAD6-42010A800017'
-        self.write_key = '5499CF6C-1642-11EC-BAD6-42010A800017'
         self.url = 'https://api.purpleair.com/v1/sensors'
+        namespace = "purpleair_api"
+        try:
+            self.read_key = keyring.get_credential(namespace, "read_key").password
+            # self.write_key = keyring.get_credential(namespace, "write_key").password
+            # Write key not used currently
+        except AttributeError:
+            self.read_key = input("Please paste your PurpleAir read key here\n"
+                               "This will be saved to your computer's encrypted keyring:")
+            keyring.set_password(namespace, "read_key", self.read_key)
+            # self.write_key = input("Please paste your PurpleAir write key here\n"
+            #                    "This will be saved to your computer's encrypted keyring:")
+            # keyring.set_password(namespace, "write_key", self.write_key)
 
 
 class AWSSettings:
     """Class to hold settings for Amazon AWS info."""
     def __init__(self):
+        namespace = "aws_purpleair_downloader"
         self.bucket_arn = 'arn:aws:s3:::purpleair-data'
-        self.access_key = "placeholder"
+        self.region = 'us-west-1'  # Northern CA
+        try:
+            self.access_key = keyring.get_credential(namespace, "access_key").password
+            self.secret_key = keyring.get_credential(namespace, "secret_key").password
+        except AttributeError:
+            self.access_key = input("Please paste your AWS IAM role access key here\n"
+                               "This will be saved to your computer's encrypted keyring:")
+            self.secret_key = input("Please paste your AWS IAM role secret key here\n"
+                               "This will be saved to your computer's encrypted keyring:")
+            keyring.set_password(namespace, "access_key", self.access_key)
+            keyring.set_password(namespace, "secret_key", self.secret_key)
 
 
 # FUNCTIONS --------------------------

### 2022-01-13 acwatt_syp_code/build/aws/lambda_download_script.py
night's commit: got lambda function working, but pandas not available in lambda function. Need to make pandas layer?

@@ -8,7 +8,7 @@ Writes IP address used during lambda function execution to S3 bucket CSV.
 """
 import io
 import logging
-import pandas as pd
+# import pandas as pd
 import datetime as dt
 import requests
 import json
@@ -23,51 +23,49 @@ logger.setLevel(logging.INFO)
 # Define a list of Python lambda functions that are called by this AWS Lambda function.
 
 
-def get_all_week_dates(date_beg, date_end):
-    # Get
-    date_start = dt.datetime.strptime(date_beg, '%Y-%m-%d')
-    date_end = dt.datetime.strptime(date_end, '%Y-%m-%d')
-    date_list = pd.date_range(date_start, dt.datetime.today(), freq='w')
-    pass
-
-
-def download_sensor_week(id_, week_beg, week_end):
-    # construct REST call to ThingsSpeak API
-
-    # Download JSON response
-
-    # Convert JSON to dataframe
-
-    pass
-
-
-def write_df_to_memcsv(df):
-    # Create CSV of df in memory
-
-    pass
-
-
-def upload_s3_csv(memcsv):
-    # Write fileobject to S3 bucket
-    pass
-
-
-def download_sensor(id_, date_beg, date_end):
-    # Get beginning and end dates of all weeks between creation and last modified
-    weeks_tuple_list = get_all_week_dates(date_beg, date_end)
-
-    # For each week
-    df_list = []
-    for week_beg, week_end in weeks_tuple_list:
-        # Download the Purple Air sensor data to dataframe
-        df_list.append(download_sensor_week(id_, week_beg, week_end))
-
-    # Write df to CSV in memory
-    df = pd.concat(df_list)
-    write_df_to_memcsv(df)
-
-    # Write S3 file
-    pass
+# def get_all_week_dates(date_beg, date_end):
+#     #
+#     date_list = pd.date_range(date_beg, date_end, freq='w')
+#     pass
+#
+#
+# def download_sensor_week(id_, week_beg, week_end):
+#     # construct REST call to ThingsSpeak API
+#
+#     # Download JSON response
+#
+#     # Convert JSON to dataframe
+#
+#     pass
+#
+#
+# def write_df_to_memcsv(df):
+#     # Create CSV of df in memory
+#
+#     pass
+#
+#
+# def upload_s3_csv(memcsv):
+#     # Write fileobject to S3 bucket
+#     pass
+#
+#
+# def download_sensor(id_, date_beg, date_end):
[…]

### 2022-01-13 acwatt_syp_code/build/aws/lambda_services.py

@@ -14,6 +14,7 @@ import datetime as dt
 import io
 import json
 import logging
+import os
 import random
 import time
 import zipfile
@@ -64,14 +65,17 @@ def create_lambda_deployment_package(function_file_name):
                                function.
     :return: The deployment package.
     """
+    cwd = os.getcwd()
+    os.chdir(PATHS.code / 'build' / 'aws')
     buffer = io.BytesIO()
     with zipfile.ZipFile(buffer, 'w') as zipped:
         zipped.write(function_file_name)
     buffer.seek(0)
+    os.chdir(cwd)
     return buffer.read()
 
 
-def create_iam_role_for_lambda(iam_resource, iam_role_name):
+def create_iam_role_for_lambda(iam_resource, iam_role_name, bucket_arn):
     """
     Creates an AWS Identity and Access Management (IAM) role that grants the
     AWS Lambda function basic permission to run S3 bucket reads and writes.
@@ -82,7 +86,7 @@ def create_iam_role_for_lambda(iam_resource, iam_role_name):
     :return: The newly created role.
     """
     lambda_assume_role_policy = {
-        'Version': '2022-01-01',
+        'Version': '2012-10-17',
         'Statement': [
             {
                 'Effect': 'Allow',
@@ -93,7 +97,20 @@ def create_iam_role_for_lambda(iam_resource, iam_role_name):
             }
         ]
     }
-    policy_arn = 'arn:aws:iam::aws:policy/service-role/AmazonS3ObjectLambdaExecutionRolePolicy'
+    s3_permissions_policy_doc = {
+        "Version": "2012-10-17",
+        "Statement": {
+            "Effect": "Allow",
+            "Action": "s3:PutObject",
+            "Resource": bucket_arn
+        }
+    }
+
+    policy_arn = 'arn:aws:iam::aws:policy/service-role/AWSLambdaBasicExecutionRole'
+    # policy_arn = 'arn:aws:iam::aws:policy/service-role/AmazonS3ObjectLambdaExecutionRolePolicy'
+
+    # TODO: find policy ARN that can be applied to lambda functions allowing them
+    #   to write to S3 buckets
 
     try:
         role = iam_resource.create_role(
@@ -103,6 +120,13 @@ def create_iam_role_for_lambda(iam_resource, iam_role_name):
         logger.info("Created role %s.", role.name)
 
         role.attach_policy(PolicyArn=policy_arn)
+
+        # Create and attach S3 policy
[…]

### 2022-01-13 acwatt_syp_code/utils/config.py

@@ -87,9 +87,10 @@ class AWSSettings:
     """Class to hold settings for Amazon AWS info."""
     def __init__(self):
         namespace = "aws_purpleair_downloader"
-        self.bucket_arn = 'arn:aws:s3:::purpleair-data'
+        self.bucket_arn = 'arn:aws:s3:::purpleair-data/*'
         self.region = 'us-west-1'  # Northern CA
         try:
+            self.account_id = keyring.get_credential(namespace, "account_id").password
             self.access_key = keyring.get_credential(namespace, "access_key").password
             self.secret_key = keyring.get_credential(namespace, "secret_key").password
         except AttributeError:

### 2022-01-13 acwatt_syp_code/build/aws/lambda_download_script.py
making notes

@@ -22,7 +22,7 @@ logger.setLevel(logging.INFO)
 
 # Define a list of Python lambda functions that are called by this AWS Lambda function.
 
-
+# TODO: get list of weeks from lambda_services.py
 # def get_all_week_dates(date_beg, date_end):
 #     #
 #     date_list = pd.date_range(date_beg, date_end, freq='w')
@@ -110,6 +110,7 @@ def ip_test(id_, bucket):
     df = pd.DataFrame({'ip_address': ip}, index=[id_])
     # Save the df to a CSV that only exists in memory (not written to disk)
     mem_csv = io.StringIO()
+    # TODO: write txt instead
     df.to_csv(mem_csv, index=False)
     mem_csv.seek(0)  # need to set position of buffer back to begging before reading
     # Write CSV to S3 bucket (True/False for success/failure)
@@ -144,16 +145,14 @@ def lambda_ip_s3_writer(params, bucket_dir='sensor_csvs'):
     return response
 
 
-
-
-
-
 """
 NOTES
 - test very simple IP address writer asap
 - need to determine how to make / break calls to get different IP addresses.
 
-- Would like to figure out easy way of including pandas
+- Would like to figure out easy way of including pandas (and requests?)
+https://duckduckgo.com/?q=python+write+txt+file+in+memory+io.StringIO&t=newext
+
 - Need to remove pandas dependency:
     - Need to pass list of weeks to lambda function
     - write json files directly to S3 instead of turning to csvs

### 2022-03-06 1_introduction.tex
Updates from Overleaf

%===================================================
%   Introduction
%===================================================

% Something about measurement error in environmental economics... robustness-to-measurement-error replication study, focusing on omitted measurement error from scientifically measured data (like temperature, elevation, pollution concentration, satellite approximated surface characteristics, etc).

% =====================================
%               1. HOOK (1-2)
% =====================================
% A good introduction starts with a good “hook,” i.e., something that grabs the reader’s attention and makes her want to keep reading. Here, the closer one can get to the reader, the better. Likewise, the broader one can go, the better. Bad hooks tend to appeal to the literature: “A long literature in economics has looked at ...” If that is the case, do you really want to make it any longer? Good hooks tend to relate to the real world: A lot of the food we buy at the grocery store is grown in the context of long value chains. What does the first link in that value chain look like? What does participating in those value chains do for the people who actually grow the food we eat? The hook should be one or two paragraphs long.

% Attract the reader’s interest by telling them that this paper relates to something interesting. What makes a topic interesting? Some combination of the following attributes makes Y something worth looking at.
% Y matters: When Y rises or falls, people are hurt or helped.
% Y is puzzling: it defies easy explanation.
% Y is controversial: some argue one thing while other say another.
% Y is big (like the service sector) or common (like traffic jams).

% Things to avoid:
% The bait and switch : promising an interesting topic but delivering something else, in particular, something boring.
% “all my friends are doing it” : presenting no other motivation for a topic than that other people have written papers on it.

A critical input to good air quality regulation is good air quality measurement.
%
Specifically, the efficiency of current pollution regulation hinges on our ability to accurately monitor air quality across the country.
%
In the United States, air quality is assessed by the government using a network of monitors that measure levels of ambient air pollution to a high degree of accuracy.
%
The Environmental Protection Agency (EPA) requires these monitors to measure average daily air quality at specific frequencies to ensure enough data is collected for effective regulation.\footnote{The three main measurement frequencies require measuring daily average air quality every 1, 3 or 6 days.}
%
During the days that are required to be measured, the goal is to accurately measure the daily average pollution concentration at the site of the monitor.
%
Statistics of these daily averages, called \textit{design values}, are then used to decide if a region is in or out of compliance with the National Ambient Air Quality Standards (NAAQS).

Though these air quality monitoring stations are regulated by the EPA, they are managed by local and state officials who control when the monitors are on or off.
%
For added flexibility, the EPA allows for some portion of air quality readings to be missing when calculating the design values that determine a region's compliance with the NAAQS.
% \citep{epa_appendix_2017}
For instance, when measuring particulate matter in the air (one of the most common types of pollution), the EPA allows more than 25\% of measurements to be missing or omitted \citep{epaAppendixPart502017}.\footnote{Design values are used to decide compliance with NAAQS and are statistics of daily averages. In calculating daily averages, the daily average is valid if at least 75\% of the hourly readings (18 of 24 hours) are reported and valid. In calculating the design values, the design value is valid if at least 75\% of daily averages in each year are reported and valid. Combined, the minimum reporting standard is actually 56-57\% of all required hourly PM2.5 readings. This is slightly different for each site depending on their reporting frequency (every 1, 3, or 6 days). }
%
In effect, this flexibility means that local managers of monitoring stations can choose up to 25\% of readings to omit -- readings that would otherwise be used to determine of compliance.
%
Though the measurements of air quality at the site of the monitor are fairly accurate when the monitor is on, omitting some measurements (by turning the monitor off) can bias the daily average and compliance statistics calculated from reported measurements.
% 
Additionally, if a region is out of compliance with the standards, the region or state can potentially face large penalties and forced adoption of expensive abatement technology.

The combination of large penalties and the discretion local officials have to drop measurements that could negatively impact compliance status leads to misaligned incentives between federal regulators and local officials in charge of monitoring air quality, potentially leading to biased air quality statistics.
%
Indeed, previous research suggests that there is mismeasurement of air quality statistics occurring;
%\cite{zou_unwatched_2021} and \cite{mu_whats_2021}
\cite{zouUnwatchedPollutionEffect2021}, \cite{muWhatMissingEnvironmental2021}, \cite{graingerRegulatorsStrategicallyAvoid2019} and \cite{graingerDiscriminationAmbientAir2019} provide evidence of strategic behavior in pollution measurement on behalf of local pollution regulators.
%
This paper focuses the size of mismeasurement occurring and the effects this mismeasurement has on determining compliance.

% =====================================
%       2. RESEARCH QUESTION (1)
% =====================================
% After hooking the reader in and setting the stage, it is time to state your research question as clearly as possible. I like to do so by stating my actual research question as the first sentence of this part of my introductions. “What is the impact of participation in contract farming on the welfare of those who participate?” The clearer this is stated, the better, because the fewer are the occasions for the reader to be disappointed. This should be one paragraph long.

% Tell the reader what this paper actually does. Think of this as the point in a trial where having detailed the crime, you now identify a perpetrator and promise to provide a persuasive case. The reader should have an idea of a clean research question that will have a more or less satisfactory answer by the end of the paper. Examples follow below. The question may take two paragraphs. At the end of the first (2nd paragraph of the paper) or possibly beginning of the second (3rd paragraph overall) you should have the “This paper addresses the question” sentence.

% Research question
Specifically, I explore the question: is there a bias in reported air quality data and how might this bias affect NAAQS compliance?
% Explanation
To explore these issues, I utilize a new dataset of air quality measurements collected from consumer products (PurpleAir sensors).
%
These data help provide an independent groundtruth comparison to air quality reported to the EPA.
%
The most promising new data coming from these consumer products are PM2.5 measurements -- the concentration of particles in air that are 2.5 micrometers and below.
%
Specifically, I combine PM2.5 measurements from multiple PurpleAir sensors that are near to federally-regulated monitoring stations to estimate the PM2.5 value at the monitoring station; I use inverse distance weighting to create a weighted average of PurpleAir measurements.\footnote{Inverse distance weighting has drawbacks: it can apply very large weight to sensors very near to the NAAQS monitor and it does not take into account that some PurpleAir sensors will be better predictors for the NAAQS monitor. See Appendix section \ref{sec:app-prediction} for an alternative strategy that I plan to implement.}
%
This allows me to construct predicted values of PM2.5 at the station during times when the station's readings would be used to calculated NAAQS compliance but when the station was shut down.
%
I first examine how these predicted missing PM2.5 values compare to the reported values -- if the missing data is missing at random, I would expect the data to be similar in distribution. Then I use the predicted values from PurpleAir sensors to fill in the gaps in the NAAQS monitor's reported data and use this reconstructed dataset to generate counterfactual NAAQS compliance statistics.

The NAAQS compliance statistics for PM2.5, called \textit{design values}, are functions of the daily averages reported by air quality monitors.
%
There are two primary design values for PM2.5: the ``annual'' design value is a three-year average of the daily averages; and ``24-hour'' design value is a three-year average of the annual 98$^{th}$ percentile of daily averages.\footnote{these statistics are discussed more in the Data section. Specific formulas for these statistics are listed in the appendix.}
%
Each quarter (3-month period), these two design values are calculated and compared to the NAAQS for PM2.5.
%
If a monitor's design value is above the standard, then the monitor (and associated region) is determined to be in \textit{non-attainment} (non-compliance) with the standard for that quarter.
%
Using the reconstructed dataset of PM2.5 (PM2.5 estimates for all hours that would be reported from a given NAAQS monitor), I construct counterfactual estimates of the design values that determine if a region is in or out of attainment.
%
I use these counterfactuals to determine which regions would have changed compliance status if they reported 100\% of their PM2.5 measurements -- I call these ``flipped regions''.
%
I also examine how close these flipped regions were to the regulatory threshold and report a measure of the bias related to the station's missing PM2.5 readings.

% =====================================
%           3. ANTECEDENTS
% =====================================
% After stating your research question, it is time to relate it and what you are doing to the existing literature. Here, relate your work to the five to ten closest studies (the closer to five, the better) in the literature. What the relevant literature -- the antecedents—is will obviously depend on the question at hand. If you are lucky enough to work in a literature that has seen a lot of activity, you may have a hard time narrowing it down, and you will need to judiciously pick the five to ten closest studies. If you are working on a problem that no one has really looked at, or that no one has looked at in a long time, you might have to go back in time a bit further or expand your parameters for what counts as antecedents. Here, what counts is to tell a bit of a story; no one wants to read a bland enumeration of studies: “Johnson (2011) found this. Wang (2012) found that. Smith (2013) found something else. Patel (2015) found something else altogether.” For every topic, the intellectual history of that topic can be told in an interesting way.

% Relate your work to the five closest studies in the literature
% tell a bit of a story

Though I am examining the effect of pollution data that are missing from a monitor's record (data missing \textit{in time}), there is also the issue of attempting to measure a region's ambient air quality using spatially sparse locations of monitors (you could consider this and issue of data missing \textit{in space}).
%
Previous literature has examined the sparse distribution of regulation-grade monitors and the resulting sensitivity of CAA air quality regulation.
%\cite{grainger_regulators_2019} and \cite{grainger_discrimination_2019}
\cite{graingerRegulatorsStrategicallyAvoid2019} and \cite{graingerDiscriminationAmbientAir2019} identify a principle-agent problem with the initial spatial placement of sparse pollution monitors; they find evidence that local regulators may be strategically locating their air quality monitors based on pollution, and possibly socioeconomic characteristics. 
%\citealt{sullivan_using_2018}, \citealt{fowlie_bringing_2019}
To address the issue of sparse data and fill in the gaps, several authors have used satellite data products to provide finer resolution pollution data (\citealt{sullivanUsingSatelliteData2018}, \citealt{fowlieBringingSatelliteBasedAir2019}).
%\cite{zou_unwatched_2021}
Moving to more time-based issues, \cite{zouUnwatchedPollutionEffect2021} also uses satellite estimates to discuss the issue of strategic behavior in reaction to the timing of pollution monitoring. He provides evidence that some areas have significantly worse air quality on unmonitored days.
% \cite{mu_whats_2021}
In related work, \cite{muWhatMissingEnvironmental2021} show potential for strategic monitor shutdowns on days of expected high pollution, contributing to air quality data that is missing \textit{in time}.

% =====================================
%           4. VALUE ADDED
% =====================================
% This is where you need to shine. What is your contribution? How does your paper change people’s priors about your topic? Ideally, your paper will have three contributions. For instance, you may be improving on the internal validity front for the question you are looking at by having a better identification strategy. You may also be improving on the external validity front by having data that cover a broader swath of the real world; or you may be performing a mediation analysis that allows identifying what mechanism m the treatment variable D operates through in causing changes in y. Lastly, you may also be bringing a small methodological improvement to the table. But even papers with one contribution deserve to be published, provided that contribution is important enough.

% Contribution 1 \cite{fowlie_bringing_2019}
This paper is most similar to the analysis in \cite{fowlieBringingSatelliteBasedAir2019} where they use PM2.5 estimates generated from satellite data to examine counterfactual compliance status.
%
However, they end their analysis noting that the satellite-based data commonly used in these applications has significant prediction error in some areas; this can cause result in incorrect conclusions about design values.
%\citeauthor{mu_whats_2021}
This paper compliments their analysis and that of \citeauthor{muWhatMissingEnvironmental2021}, where I use a different form of ground-truth PM2.5 data to also address missing air quality data \textit{in time}. \footnote{PurpleAir data, and other on-the-ground pollution sensors, also have the potential to examine issues of spatial distributions of monitor networks -- work left for future research.} In contrast to \citeauthor{muWhatMissingEnvironmental2021} however, I am examining pollution at times when it is missing in the data but required to be reported, whereas their work was on pollution at times that are not required to be reported.
%
While satellite-based PM2.5 estimates have potential for large prediction errors, PurpleAir sensors can be fairly accurate measures of their local air quality\footnote{PurpleAir sensors have specifically been shown to be less accurate than regulation-grade monitors at high levels of PM2.5 concentration. 
However, the EPA has developed a correction technique that result in PurpleAir readings within 5\% of co-located EPA monitors. 
This correction technique is used here and explained in more detail in the appendix.} and can be averaged over multiple nearby sensors.
%
PurpleAir data also have drawbacks however -- the sensors are highly non-uniform in coverage across the US and are sensitive to specific placement by the consumer, perhaps leading to hyper-local estimates of air quality.

For these reasons, this analysis should be seen as a compliment to previous works.
%
As consumer sensors become more widespread, we can augment reliable federal air quality measurements with a growing number of auxiliary data points to better understand the shape of mismeasurement in air quality.
%
In this paper, I explore one way of leveraging these data to test for issues with biased reporting of air quality.
%
After predicting missing observations using PurpleAir measurements, I do not find any statistically significant flipping of NAAQS compliance status in the 15 California NAAQS monitors. However, for a monitor in Fresno, CA, I do find differences between design values calculated on reported NAAQS data and design values calculated on imputed data; these differences persist for five quarters in 2020 and 2021 and are statistically significant at more than the 95\% level. 
%
The largest discrepancy for Fresno is in the 24-hour design values, where the difference is more than 2.5$\mu$g/m$^3$ of PM2.5 on average between 2018 and 2021. 
%\citep{currie_what_2020}
Fresno has been out of compliance for some time, and these results mimic previous literature that suggests larger pollution measurement problems in nonattainment areas.
%
There is also evidence from previous research that pollution in non-attainment areas has been decreasing at significantly faster rates since the introduction of the CAA \citep{currieWhatCausedRacial2020}.
%
Due to the small sample size of this study, there is not much evidence of widespread biased pollution standards. However, with the framework now setup, it would be possible to expand this type of analysis to all NAAQS monitors in the United States and all PurpleAir monitors available. 
% With continued growth in the adoption of consumer-based air quality monitoring, it is possible we will see increased resolution of air quality and

Ultimately, a possible outcome of this line of research is estimating the possible gains to be made in changing reporting standards. Increasing reporting standards to decrease allowable omitted observations may result in more non-attainment areas and further increases regulatory efficiency.

% Contribution 2

% Contribution 3

% =====================================
%           5. ROAD MAP
% =====================================
% Lastly, you should provide your reader with a roadmap to your paper. This section usually starts with “The remainder of this article is organized as follows,” and it lists section and what they do in order. So for a typical paper, it would go: “The remainder of this paper is organized as follows. Section 2 presents the theoretical framework used to study the research question and derives this paper’s core testable prediction. In section 3, the empirical framework is presented, first by discussing the estimation strategy, and then by discussing the identification strategy. Section 4 presents the data and discusses some summary statistics. In section 5, the empirical results are presented and discussed, followed by a battery of robustness checks and a discussion of the limitations of the results. Section 6 concludes with policy recommendations and suggestions for future research.” I have seen some economists on social media state that they have gotten papers rejected for many reasons, but never for want of a roadmap section. Fair enough. In most cases, however, it is simply easier to include such a roadmap section and delete it at a reviewer’s request than to not have one and have to write one when asked to revise and resubmit a paper, not to mention the fact that some readers will simply expect there to be a roadmap, since the majority of applied economics articles include them. Anything that signals that you know what the unspoken rules and norms of the profession are is a good thing for your article’s chances of getting published.

% Section 1.b (optional): background (not lit review)
% Section 2: theoretical framework, testable prediction
% Section 3: empirical framework: estimation strategy, identification strategy
% Section 4: data, summary stats
% Section 5: empirical results, discussion, robustness checks, limitations of the results
% Section 6: policy recommendations, future research

The remainder of this article is organized as follows. Section \ref{background} briefly reviews the history of air quality standard in the US and some key details of current regulations. Section \ref{data} then discusses the data used and section \ref{emprical} describes the theoretical and empirical framework that will be applied to estimate the missing pollution and effects on reference levels for national standards. Section \ref{results} reviews the results of the empirical study and concludes.

### 2022-03-06 2_background.tex
Updates from Overleaf

%\citep{91st_us_congress_clean_1970}
Amid growing public concern about air quality and pollution, the United States Congress passed the Clean Air Act of 1963 (CAA). Later additions to the CAA, the Clean Air Amendments of 1970, granted the Environmental Protection Agency (EPA) the regulatory authority to create and enforce air quality standards in the US. One major way air quality is regulated is through the National Ambient Air Quality Standards (NAAQS), which set concentration thresholds for a list of different ``criteria'' pollutants
\citep{91stuscongressCleanAirAmendments1970}. The EPA has since been in charge of setting and updating the NAAQS and require states to submit plans to bring their air quality to within NAAQS limits. An important aspect of enforcing the NAAQS is measuring criteria pollutants across the US by requiring states to install pollution monitoring stations in areas of questionable air quality. Because these monitoring stations are used for potentially costly enforcement, the equipment within each station must abide by specific regulations and are relatively costly to install and run.

Over the last decade, commercially available scientific equipment in measuring various air pollutants has evolved. There is now relatively cheap\footnote{e.g., a PurpleAir outdoor air quality sensor is about \$250 to purchase with little upkeep from the end user, compared to roughly \$100,000-200,000 to install EPA regulation-grade criteria pollutant monitors and trained staff to upkeep and record measurements. The cost alone is not a good comparison because the EPA monitors use different technology that is known to be more accurate across a wider range of pollution concentrations, have a better sense the sensor error, and measure more pollutants than the PurpleAir monitors. For the purposes of this analysis, PurpleAir monitors should be seen as a compliment to EPA monitors, not a potential replacement.} equipment available to measure particulate matter (one of the criteria pollutants that regulated by the NAAQS). Specifically, the PurpleAir company produces devices that can measure particulate matter that has a diameter of less than 2.5 micrometers (designated as PM2.5).\footnote{PurpleAir devices can measure a few other criteria pollutants (namely ozone and PM10) but the comparability of the PM2.5 measurements between PurpleAir and EPA monitors are currently better understood.} PurpleAir is of particular interest because they have built an opt-out mechanism for end-users to allow their ambient air quality data to be stored in the cloud. They also provide multiple ways for researchers and the general public to use this crowd-sourced air quality data.

This paper is primarily concerned with the minimum reporting requirement. As with many federal regulations, there are many ways that states or emitters can cleverly navigate the rules to emit more than they are meant to according to the spirit of the regulation. One way of navigating the CAA regulations is through the choice of what data to report. The EPA currently requires a minimum threshold of air quality data to be reported -- for PM 2.5, 75\% of daily measurements need to be reported, and each day must have 75\% of hours reported. That leaves many choices of which hours to turn the monitor off for cleaning, calibration, or other reasons. I wish to understand how these timing decisions are affecting the distribution of reported data -- specifically how it might be affecting a statistic of that distribution: the design value.

### 2022-03-06 pics/appendix/tables/attainment_counties.tex
Updates from Overleaf

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{multirow}
% \usepackage{longtable}
% Note: It may be necessary to compile the document several times to get a multi-page table to line up properly
\begin{longtable}{@{}lllll@{}}
\caption{Current NAAQS Nonattainment Counties}
\label{tab:naaqs_attain_counties}\\
\toprule
\textbf{State} & \textbf{Area Name} & \textbf{\begin{tabular}[c]{@{}l@{}}EPA Designated \\ Nonattainment Counties\end{tabular}} & \textbf{FIPS} & \textbf{\begin{tabular}[c]{@{}l@{}}\# of monitors\\ in sample\end{tabular}} \\* \midrule
\endfirsthead
%
\endhead
%
\multicolumn{1}{|l|}{\multirow{14}{*}{CA}} & \multicolumn{1}{l|}{Imperial County, CA} & \multicolumn{1}{l|}{Imperial, CA (p)} & \multicolumn{1}{l|}{025} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{\multirow{8}{*}{\begin{tabular}[c]{@{}l@{}}San Joaquin Valley \\ Air Basin, CA\end{tabular}}} & \multicolumn{1}{l|}{Fresno, CA} & \multicolumn{1}{l|}{019} & \multicolumn{1}{l|}{2} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Kern, CA (p)} & \multicolumn{1}{l|}{029} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Kings, CA} & \multicolumn{1}{l|}{031} & \multicolumn{1}{l|}{2} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Madera, CA} & \multicolumn{1}{l|}{039} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Merced, CA} & \multicolumn{1}{l|}{047} & \multicolumn{1}{l|}{} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{San Joaquin, CA} & \multicolumn{1}{l|}{077} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Stanislaus, CA} & \multicolumn{1}{l|}{099} & \multicolumn{1}{l|}{} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Tulare, CA} & \multicolumn{1}{l|}{107} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{\multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Los Angeles-South \\ Coast Air Basin, CA\end{tabular}}} & \multicolumn{1}{l|}{Los Angeles, CA (p)} & \multicolumn{1}{l|}{037} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Orange, CA} & \multicolumn{1}{l|}{059} & \multicolumn{1}{l|}{1} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Riverside, CA (p)} & \multicolumn{1}{l|}{065} & \multicolumn{1}{l|}{} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{San Bernardino, CA (p)} & \multicolumn{1}{l|}{071} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{Plumas County, CA} & \multicolumn{1}{l|}{Plumas, CA (p)} & \multicolumn{1}{l|}{063} & \multicolumn{1}{l|}{} \\* \midrule
\multicolumn{1}{|l|}{ID} & \multicolumn{1}{l|}{West Silver Valley, ID} & \multicolumn{1}{l|}{Shoshone, ID (p)} & \multicolumn{1}{l|}{079} & \multicolumn{1}{l|}{} \\* \midrule
\multicolumn{1}{|l|}{\multirow{2}{*}{OH}} & \multicolumn{1}{l|}{\multirow{2}{*}{Cleveland, OH}} & \multicolumn{1}{l|}{Cuyahoga, OH} & \multicolumn{1}{l|}{035} & \multicolumn{1}{l|}{} \\* \cmidrule(l){3-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{Lorain, OH} & \multicolumn{1}{l|}{093} & \multicolumn{1}{l|}{} \\* \midrule
\multicolumn{1}{|l|}{\multirow{3}{*}{PA}} & \multicolumn{1}{l|}{Delaware County, PA} & \multicolumn{1}{l|}{Delaware, PA} & \multicolumn{1}{l|}{035} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{Lebanon County, PA} & \multicolumn{1}{l|}{Lebanon, PA} & \multicolumn{1}{l|}{075} & \multicolumn{1}{l|}{} \\* \cmidrule(l){2-5} 
\multicolumn{1}{|l|}{} & \multicolumn{1}{l|}{Allegheny, PA} & \multicolumn{1}{l|}{Allegheny, PA} & \multicolumn{1}{l|}{005} & \multicolumn{1}{l|}{} \\* \bottomrule
\end{longtable}

### 2022-03-06 paper.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lipsum}
\usepackage{titling}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\doublespacing
\usepackage{graphicx}
\usepackage{placeins} % \FloatBarrier
\usepackage{wrapfig}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{lscape}
\usepackage{caption}
\graphicspath{{pics/}}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }

\urlstyle{same}

% Let overleaf run longer to compile many pictures
\maxdeadcycles=200

\title{Filling in the Gaps: Using Consumer Products to Replace Missing Pollution Data}
\author{Aaron Watt}
\date{February 2022}
\setlength{\droptitle}{-10em}

% REVIEW COUNT -- goal is 100 by April
% increase the count each time the intro is reviewed
% 8

% Helpful docs:
% CONVERSABLE ECONOMIST: Writing the Intro to Your Economics Research Paper (Timothy Taylor)
% Bellemare, “How to Write Applied Papers in Economics.”
% McCloskey, Economical Writing.

% Task: Submission of your complete paper. Typically a complete paper will include an introduction, a model, an econometric equation, a description of the data, discussion of results, and a conclusion. The paper should be no more than 30 pages using a 12 pt font, 1-inch margins, including tables, graphs and reference, and with the core text double-spaced.

\begin{document}

\maketitle

%============================================
\begin{abstract}
%============================================
% Typically, it is possible to write a solid draft of your abstract by keeping only the first sentence of the hook, research question, and value added sections of your introduction, and by polishing up the resulting paragraph some.
% Except for the requisite terminology (e.g., randomized controlled trial, difference-in-differences, regression discontinuity), your abstract should be intelligible to any smart, college-educated person who is not an economist. This is especially true for an applied economics paper. After all, we are writing about real-world phenomena that are of interest to policy makers or business managers, so your abstract should be intelligible to someone with a master’s degree in public policy or in business administration, depending on what you are doing. Do not make the mistake of confusing lack of intelligibility with intellectual rigor; this is economics, not French postmodern philosophy.
% If your title is not repellent, and if your abstract is intelligible to people who are not experts in your field and to people in other disciplines, you have just expanded the scope of your citations tenfold, because whether one likes it or not, a lot of people cite stuff they have only read the abstract of.

% Research Question
[abstract here]

% Proposed Model

% Dataset

% Estimation and testing plan

\vspace{2em}
\end{abstract}

% Research Question
% How much do prediction errors matter in pollution regulation? Does incorporating prediction errors of machine-learning-produced pollution data affect the policy categorization of areas without a pollution monitor?

%==========================================
\section{Introduction} 
\label{introduction}
%==========================================
\input{1_introduction}

%==========================================
\section{Background} 
\label{background}
%==========================================
\input{2_background}

%===========================================
\section{Data} \label{data}
%===========================================

\subsection{EPA Regulation-grade Monitors}
There are currently 388 air quality monitoring stations around the US that are used for NAAQS determination for PM2.5; I will refer to these monitors as \textit{NAAQS monitors}. There are more regulation-grade monitors that meet or approach the regulatory accuracy standards set by the EPA, but these 388 are the monitors that are officially used to calculate the design values that decide NAAQS attainment status. Of the 388 NAAQS monitors in the US, I limit my preliminary analysis on the 15 monitors in California that take hourly readings every day. Future analysis will include the full set of NAAQS monitors, which include monitors that only report daily averages (as opposed to hourly averages) and monitors that only report every 2, 3, 6, or 12 days.

%\citep{fowlie_bringing_2019}
\textbf{Design Values.} PM2.5 \textit{design values} are statistics of hourly PM2.5 concentrations reported by the NAAQS monitors. In reality, design value determination for a monitor begins by calculating the initial design value on the non-missing data and then includes a negotiation step between EPA and the local regulator to decide the final, publicly-reported design value. I could use final design values for each monitor that are listed in EPA reports. However, I am interested in directly comparing the design values calculated from only reported data to design values that include predictions of unreported data. Because I cannot replicate the final negotiation process, I replicate work done in \citep{fowlieBringingSatelliteBasedAir2019} to create \textit{pseudo design values} by calculating the statistic on the data and making comparisons based on this initial design value. There are two NAAQS design values for PM2.5 explained in Section \ref{design_value_equations}: annual and 24-hour. These design value calculations only use valid daily and annual averages, where validity is determined based on the number of reported and non-excluded observations.

% \cite{epa_event_1990}
\textbf{Excluded Readings.} There are a number of events that create air quality measurements that cannot be used in NAAQS determination; wildfires or machine calibrations (for example) can cause hour- or day-long readings to be invalid for the purposes of NAAQS determination. These times, referred to as \textit{exceptional events} (EE), are events that are ``not expected to recur routinely at a given location, or that [are] possibly uncontrollable or unrealistic to control through the [NAAQS regulatory] process''\citep{epaEventQualifier1990}. These events are identified in the NAAQS monitor data and removed from the analysis: hours that have been labeled as EE are removed from both the PurpleAir and NAAQS monitor data before calculating design values.\footnote{See Appendix Section \ref{sec:tables} Table \ref{tab:excluded_qualifiers} for a detailed list of reasons that observations are excluded from NAAQS determinations.} These are not considered ``missing'' or ``unreported'' data for the sake of predicting missing values, however these are considered invalid observations in the design value calculation. Removing EE provides more realistic pseudo design value estimates.

\subsection{PurpleAir Consumer Sensors}
The last ten years have seen a growing interest in consumer-based air quality measurement. PurpleAir air quality sensors are designed to mainly measure PM2.5, but also measure other pollutants (PM10, ozone) and environmental factors (humidity, temperature).\footnote{See Appendix section \ref{sec:monitor_pictures} for pictures of both a NAAQS monitoring station and a typical PurpleAir outdoor pollution sensor}. In my analysis, PurpleAir PM2.5 data plays a ground-truth role -- it gives me an alternative source of PM2.5 measurements to rely on when the NAAQS monitor is shut off.

To examine how design values might be influenced by missing data, I predict missing PM2.5 hourly average concentrations from EPA NAAQS monitors using nearby PurpleAir PM2.5 sensors. For an initial analysis, I limit the sample to include PurpleAir sensors within 5 miles of each NAAQS monitor, or extending up to 25 miles to get 10 PurpleAir sensors minimum for each monitor.

This is a fairly new and rich dataset: there have been more than 16,000 public PurpleAir sensors brought online in the United States since 2015. When a consumer is setting up their sensor, they have the choice to make the sensor public or private. All sensors upload their PM2.5 readings to an online server, but only public sensors have data available for research use. The company asks consumers to make their data public if possible, attempting to contribute to more citizen science. Of the 16,000+ US sensors, there are 10,401 in California, Oregon, Nevada, and Arizona and I limit my sample to the 592 unique PurpleAir sensors within 5 miles of 15 NAAQS-primary monitors.

\textbf{Correction of PurpleAir Readings.} PurpleAir sensors are known to have worse readings at higher levels of pollution. I modify PurpleAir PM2.5 values using the EPA's correction equation for PurpleAir sensors. The calibrated this equation by studying co-located PurpleAir and NAAQS monitors.
$$
\widetilde{PA}_{j,t}=\begin{cases}
			0.52*PA_{j,t} - 0.086*H_{j,t} + 5.75, & \text{if $PA_{j,t} \leq 343 \mu$g/m$^3$}\\
            0.46*PA_{j,t} + 0.(3.93e-4)PA_{j,t}^2 + 2.97, & \text{otherwise}
		 \end{cases}
$$
where $PA_{j,t}$ is the ambient PM2.5 measured by PurpleAir sensor $j$ at time $t$ and $H_{j,t}$ is the relative humidity (between 0 and 1) also measured by the PurpleAir device. This correction helps reduce concerns about heteroskedasticity due to larger errors in PurpleAir readings at high levels of PM2.5. Future work involves a more complex predictive model.

% Highlights:
% \begin{itemize}
% \item Goal: predict missing
% \item Limited the analysis to include PurpleAir sensors within 5 miles of each NAAQS-primary monitor.
% \item 16,038 PurpleAir outdoor, publicly-shared PM2.5 sensors in the United States have some accessible data.
% \item 10,401 sensors in California, Oregon, Nevada, and Arizona.
% \item 11,205 instances of PurpleAir sensors within 50 miles of the 15 NAAQS-primary monitors (double counting permitted).
% \item 7,777 unique PurpleAir sensors within 50 miles of 15 NAAQS-primary monitors.
% \item 592 unique PurpleAir sensors within 5 miles of 15 NAAQS-primary monitors (final sample of PA sensors contributing to estimated EPA monitor values).
% \item I correct PurpleAir PM2.5 values using EPA's correction equation.
% \item For each hour that there are valid PurpleAir sensor readings within 5 miles of the EPA monitor, I calculate an inverse-distance weighted average PM2.5 level.
% \item Future work includes a better PurpleAir prediction for EPA monitor PM2.5 measurements, using wind speed and direction.
% \item There are some PurpleAir sensors that are biasing the prediction of PM2.5 -- seem to be measuring localized pollution that the EPA monitor does not pick up. These would be down-weighted or removed in the future version of the prediction mentioned above.
% \item I show plots for an example NAAQS monitor in Los Angeles, CA. Plots for other monitors are in the appendix.
% \end{itemize}

% \FloatBarrier
% \begin{figure}[ht]
% \centering
% \includegraphics[width=0.8\textwidth]{purple_air_sensor_cum_time_california.png}
% \caption{(Red) 20-day rolling average of the Air Quality Index in California. (Blue) Cumulative PurpleAir outdoor sensors posting public PM2.5 data.}
% \label{fig:ca_purpleair_adoption}
% \end{figure}

\FloatBarrier
\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth]{purple_air_sensor_map_california.png}
\caption{Map of PurpleAir sensors offering public, outdoor PM2.5 measurements. These are sensors that have offered any data in the past, so many are now inactive. The historical data is used in this analysis.}
\label{fig:ca_purpleair_map}
\end{figure}

\FloatBarrier
\textbf{PurpleAir and a NAAQS Monitor.} As an example, figure \ref{fig:concentric_purpleair_037-4004} depicts the high number of PurpleAir sensors in the vacinity of one of Los Angeles's NAAQS monitors. I select the PurpleAir sensors in pink, those within 5 miles of the monitor.
\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth]{county-037_site-4004_epa-pa-concentric-ranges.png}
\caption{Map of an EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii. This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers).}
\label{fig:concentric_purpleair_037-4004}
\end{figure}

% \FloatBarrier
% \begin{figure}[ht]
% \centering
% \includegraphics[width=0.8\textwidth]{site-037-4004_pa-daily-covereage.png}
% \caption{Scatter plot indicating the number of hours in each day that this NAAQS-primary monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles.}
% \label{fig:hourly_coverage_037-4004}
% \end{figure}

\FloatBarrier
\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth]{site-037-4004_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. For this site, we can see the PurpleAir average is skewed to the right for readings from 2021. This is likely from a PurpleAir sensor coming online that was placed near a source of localized pollution that is not being picked up by the NAAQS-primary monitor.}
\label{fig:pa-epa-compare_037-4004}
\end{figure}

\FloatBarrier
% \subsection{Wind Speed and Direction}

\subsection{Estimating Regulation-grade Readings with PurpleAir Sensors} 
I use inverse-distance weighting (IDW) with a power of 1 on the denominator. In their discussion of IDW in ambient pollution estimation, \cite{demesnardPollutionModelsInverse2013} derives that a power between 1 and 3 is appropriate for diffuse particle distributions. I use a power of 1 here because I find evidence that some PurpleAir sensors that are very close to the NAAQS monitor are not very good predictors of the monitor's PM2.5 levels. These PurpleAir sensors seem to still have reliable estimates\footnote{Each PurpleAir sensor has two internal sensors that measure PM2.5. Reliability of the PurpleAir sensors is determined by the agreement of the two sensors' hourly averages.}, and anecdotally seem to have very high PM2.5 readings when they disagree with the NAAQS monitor. This suggests they are measuring localized pollution that is out of the range of the NAAQS monitor (these sensors could be located next to a highway, for example).

I plan to fix this issue with a future implementation of a better prediction model\footnote{See Appendix section \ref{sec:app-prediction}}. In this iteration, I have implemented the sub-optimal IDW average to avoid excluding entire PurpleAir sensors and removing potentially useful sensor data.

%===========================================
\section{Theoretical \& Empirical Framework} 
\label{theoretical}
%===========================================

\subsection{Estimating PM2.5 at the NAAQS Monitor with PurpleAir Sensors}
\FloatBarrier

To examine the difference between reported pollution and that which is missing from the NAAQS monitors' dataset, I need some measure of ambient PM2.5 levels around the monitor. A good place to start is inverse distance weighting (IDW) to create a weighted average of PurpleAir sensors that can tell us about the PM2.5 levels near the NAAQS monitor.\footnote{This is not a good place to end, however. IDW produces fairly poor estimates of PM2.5 at the location of the NAAQS monitor (before OLS). I plan to implement a more rich prediction model using wind speed and direction -- see the Appendix section \ref{sec:app-prediction} for model and data notes on this method. Ultimately, both satellite and PurpleAir data could be combined with NAAQS monitor data to provide a more accurate depiction of pollution in the US.} To help make the estimation process concrete, I will use a monitor in Los Angeles, CA as an example (see Fig. \ref{fig:idw_diagram} for a visual representation).

\begin{figure}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{IDW_diagram}
  \end{center}
  \caption{Example distances of two selected PurpleAir sensors near a NAAQS monitor in Los Angeles, CA.}
  \label{fig:idw_diagram}
\end{figure}

Consider an EPA NAAQS monitor that measures PM2.5 concentration $EPA_T$ at time $t$. Let $PA_{j,t}$ be PM2.5 concentration as measured by PurpleAir sensor $j$ at time $t$ at a distance $d_j$ from the NAAQS monitor. Assume that there are $J_t$ active PurpleAir sensors in the vicinity of the NAAQS monitor at some time $t$, each with their own PM measurements and distance. Then the Inverse-Distance Weighted average PurpleAir PM reading $PA^{IDW}_{j,t}$ is
\begin{equation}
    PA^{IDW}_t= \sum\limits_{j=1}^{J_t} \dfrac{\frac{1}{d_j}\cdot PA_{j,t}}{\sum\limits_j^{J_t}\frac{1}{d_j}}
    = \sum\limits_{j=1}^{J^t} w_{j,t}\cdot PA_{j,t}
\end{equation}
Note here that I am explicitly using the exponent of one on the distance to balance the desire to have closer sensors provide more weight but avoid having extraordinarily large weights on sensors that are relatively close to the monitor. Also note that the number of PurpleAir sensors $J_t$ changes over time as sensors come online and exit. This means the weights of the weighted average need to be calculated separately for each period. This IDW average PurpleAir measurement provides me with a measure of ambient PM2.5 variation in the vicinity of the NAAQS monitor at all the possible times there exists at least one PurpleAir monitor in the area. This helps provide good coverage and make implementing OLS in the next step easier.

\textbf{OLS Prediction.} Even if the PurpleAir sensors are highly accurate at measuring PM2.5 concentrations at their location, they may still biased as measures of PM2.5 near the NAAQS monitor. For example, someone might put up a PurpleAir sensor on a telephone pole right next to a highway, which might have high average PM compared to where the NAAQS monitor is placed. To help ensure an unbiased approximation of the missing monitor measurements \textit{at the location of the NAAQS monitor}, I use OLS to regress the NAAQS monitor data on the weighted average PurpleAir data.

\begin{equation}
EPA_t = \beta_0 + \beta_1 PA^{IDW}_t + \varepsilon_t
\end{equation}

\def\M{\mathcal{M}}\def\N{\mathcal{N}}\def\R{\mathbbm{R}}
\def\up{\widehat{EPA}^U_t} \def\low{\widehat{EPA}^L_t}
I then predict missing NAAQS monitor data (out of sample) and combine the predicted PM estimate with the reported readings. Formally, suppose $\M$ is the set of \textbf{missing} times that we do not have a PM record from the NAAQS monitor (i.e., $EPA_t$ does not exist for $t\in\M$). Let $\N$ be the \textbf{non-missing} (reported) times for the NAAQS monitor ($EPA_t\in\R\ \forall t\in\N$). For simplicity, assume that some PurpleAir data are available at all times ($PA^{IDW}_t\in\R_+\ \forall t\in\M\bigcup\N$). Using the PurpleAir data during the missing times, I predict the missing NAAQS data:

\begin{equation}
\widehat{EPA}_t = \hat\beta_0 + \hat\beta_1 PA^{IDW}_t \quad \forall t\in\M
\end{equation}

I also estimate the 95\% lower and upper bound on each predicted value; $\low$ and $\up$ to later using in lower and upper bounds on the design value.

\subsection{Estimating Design Values} \label{design_value_equations}
The NAAQS specify two primary statistics to determine if an area is in or out of attainment. These two statistics are referred to as the Annual and 24-hour Design Values. The annual design value is a 3-year average of annual averages of daily average PM2.5 levels. The 24-hour design value is a 3-year average of the annual 98th percentile of daily averages. There are also considerations about the proportion of allowed missing recordings (75\%). See Appendix Section \ref{sec:design_values} for details on constructing these design values.

\def\dva{\text{DV}_A}\def\dvh{\text{DV}_H}
\def\dvaa{\widetilde{\text{DV}}_A}\def\dvhh{\widetilde{\text{DV}_H}}
I first construct design values using the reported NAAQS monitor data. In the style of \cite{fowlieBringingSatelliteBasedAir2019}, I will call these annual and 24-hour \textit{pseudo design values}; $\dva$ and $\dvh$. I do not use the reported design values because there is a more complex negotiation that happens between the EPA and state regulators that can change some of the numbers. In order to compare design values between reported and reported + imputed datasets, I must calculate them on the actual reported PM2.5 readings from the NAAQS monitor.

I then create predicted NAAQS PM values using PurpleAir data as described in the previous section. I replace all missing NAAQS monitor values possible with the predicted values, leaving the original valid NAAQS readings. With this new imputed dataset, I calculate the new imputed design values: $\dvaa$ and $\dvhh$. Subtracting the original design value from the imputed design value gives an estimate of design value bias caused by missing data.

\begin{align}
    \text{bias}^{miss}(\dva) &\approx \dvaa - \dva\\
    \text{bias}^{miss}(\dvh) &\approx \dvhh - \dvh
\end{align}

If this quantity is positive, then there is support that there is under-reporting of pollution via allowed missing data. To gain an understanding of significance of the results, I also propagate the upper and lower bounds of the predicted observations through the design value calculation (similar to \cite{fowlieBringingSatelliteBasedAir2019}, but they were estimating out of sample prediction errors). This provides a 95\% confidence interval -- upper and lower bounds on the imputed design values: $\dvaa^{upper}$,  $\dvaa^{lower}$, $\dvhh^{upper}$, and $\dvhh^{lower}$.

\textbf{NAAQS.} The standards set out in the National Ambient Air Quality Standards give specific thresholds to compare the design values to. Above the thresholds are considered nonattainment. The primary standard for the annual design value is 12 $\mu$g/m$^3$, and has a secondary standard of 12 $\mu$g/m$^3$. The 24-hour standard is 35 $\mu$g/m$^3$.

\textbf{Exceptional Events.} As mentioned in section 3, there are many events for which the EPA allows local regulators to exclude their readings from design value calculations. These readings were removed from the dataset before any design value calculations and do not get imputed.

%==============================================
\section{Results and Discussion}
\label{results}
%==============================================
I conducted 15 design value tests on 15 different NAAQS monitors. To create the predicted design value for each NAAQS monitor, I regressed the NAAQS monitor PM2.5 readings on the weighted average PurpleAir readings. An example of this regression is below for one of the Los Angeles monitors. I ran regions both with an without a constant, but because I wanted the prediction errors to have mean zero, I chose to use model (2) in the creation of the design value. Note that the $R^2$ value in model (1) is much higher, however $R^2$ values between regressions with constant and those without are no comparable due to the difference in denominators. 
\input{pics/appendix/tables/epa_OLS_idw_pa_site-037-4004}

The number of observations here are the number of hours between 2016 and 2021 where neither the NAAQS monitor or the weighted average PurpleAir readings were missing. The slope in both models is less than one, indicating that PurpleAir tends to measure higher PM2.5 concentrations. This could be a selection issue -- consumers who choose to buy a PurpleAir monitor and place it outside their house are probably concerned about pollution -- or a measurement difference between the types of devices.

 For all sites, I generated kernel density plots like those in Fig. \ref{fig:missing-density_019-0500}. Only one of the 15 sites I tested had noticeable differences in the PM2.5 distributions for missing and reported hours (see the appendix for all sites' plots). However, the design values are the policy-relevant statistic of those distributions. Looking at the top and bottom figures below however, we can guess that the means and 98th percentile might be similar in the top case, and could plausibly be different in the bottom case.

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{site-037-4004_epa-pa-missing-density.png}
\includegraphics[width=0.6\textwidth]{appendix/site_plots/site-019-0500_epa-pa-missing-density.png}
\caption{Comparison of estimated kernel densities of PM2.5 concentration for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. The top image is for a monitor at a site in LA, and the bottom is at a site in Fresno, CA.}
\label{fig:missing-density_019-0500}
\end{figure}

\begin{table}[]
\caption{Design Value Comparison for Fresno, CA. (98\% CI Bounds)}
\label{tab:fresno_dvs}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}l|lll|lll@{}}
Year-Quarter & \begin{tabular}[c]{@{}l@{}}Annual DV\\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower\\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Hour DV \\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower \\ Bound\end{tabular} \\ \midrule
2018-4 & Invalid DV &  &  & Invalid DV &  &  \\
2019-1 & -0.005 & 0.133 & -0.143 & 0.000 & 0.252 & 0.000 \\
2019-2 & -0.003 & 0.141 & -0.147 & 0.000 & 0.252 & 0.000 \\
2019-3 & 0.015 & 0.170 & -0.139 & 0.058 & 2.202 & 0.000 \\
2019-4 & 0.002 & 0.190 & -0.185 & 0.000 & 1.460 & -0.024 \\
2020-1 & -0.012 & 0.182 & -0.207 & 0.000 & 0.335 & 0.000 \\
2020-2 & -0.010 & 0.191 & -0.211 & 0.000 & 0.376 & 0.000 \\
2020-3 & 0.679 & 0.954 & 0.403 & 8.718 & 11.704 & 8.556 \\
2020-4 & 0.647 & 1.006 & 0.288 & 5.979 & 8.281 & 5.851 \\
2021-1 & 0.564 & 1.036 & 0.091 & 3.007 & 4.184 & 2.903 \\
2021-2 & 0.533 & 1.024 & 0.042 & 3.007 & 4.225 & 2.903 \\
2021-3 & 0.630 & 1.129 & 0.132 & 7.607 & 10.557 & 7.444
\end{tabular}%
}
\end{table}

Table \ref{tab:fresno_dvs} reports the difference between the pseudo design value (calculated using the reported data) and the imputed design value -- what I believe is an estimate on the design value bias due to low reporting standards. We can see the first row is invalid. This is common in the full results in the Appendix. This occurs because one of the design values have too many invalid days in the quarter (or the 11 quarters before it) the completeness criteria are violated. Examining the Annual design value difference column, the numbers do not seem striking though the lower half do seem to be statistically significant. However, these are not materially significant -- the EPA has rounding conventions (to the ones place) that would make zero effectively within the lower bound since a design value falling below 0.5 would be rounded to 0.

These unremarkable results for the annual design value bias are fairly representative of the batch of 15 sites. When we look at the hour design value results, we can see in 2020 and 2021, there were fairly large and significant differences between the design values. This indicates that having more restrictive data collection standards could have increased Fresno's design value significantly in these quarters. Because OLS was used to impute the missing values, we would expect the hourly imputed values to be unbiased on average. Since we are taking 3-year averages, it seems plausible to believe most bias in the imputation method would average out. These combined with the lower bounds being well above zero suggest there was indeed under-reporting of pollution in Fresno. I cannot say whether or not that it happened intentionally or by mere chance of turning off the monitor at times that would likely have recorded higher pollution.

Is this meaningful for Fresno though? Fresno has had a nonattainment status for many years and is listed on several lists within this literature as having notable pollution measurement issues. So though Fresno's recorded design value would have been higher, they were already well over the 24-hour design threshold of 35 $\mu$g/m$^3$, having pseudo design values in the 50s and 60s in 2020 and 2021.

This is also just one of 15 cases, where the other 14 did not show interesting results. So is this just p-hacking to find the case study that is, by chance, abnormal? In regulation of air pollution, one might be concerned with individual actors or regions trying to get around the regulations. Since the Clean Air Act and the NAAQS are designed to ensure that all American residents can share in a minimum standard, it seems relevant to me to be able to diagnose mismeasurement.

This work is (hopefully) just the beginning of a line of research regarding the distribution of pollution and pollution policies in the US. There is a to-do list in the Appendix.

\newpage
\bibliographystyle{chicago}
\bibliography{references}
% Citation command	Output
% \citet{goossens93}      Goossens et al. (1993)
% \citep{goossens93}      (Goossens et al., 1993)	
% 
% \citet*{goossens93}     Goossens, Mittlebach, and Samarin (1993)
% \citep*{goossens93}     (Goossens, Mittlebach, and Samarin, 1993)
%
% \citeauthor{goossens93}    Goossens et al.
% \citeauthor*{goossens93}   Goossens, Mittlebach, and Samarin

% \citeyear{goossens93}      1993
% \citeyearpar{goossens93}  (1993)

% \citealt{goossens93}     Goossens et al. 1993
% \citealp{goossens93}     Goossens et al., 1993
% \citetext{priv.\ comm.}	(priv. comm.)

\newpage
\section{Appendix}

\subsection{Research Project To-do's} \label{sec:todo}

\textbf{Future Work}
\begin{itemize}
    \item I have only tested 15 of 388 possible sites. Now that I have written the bulk of the code the manage the data, most of the future work lies in acquiring the rest of the data for the NAAQS monitors and PurpleAir monitors.
    \item I am in the process of negotiating a data usage agreement with PurpleAir staff to get the entire historical dataset of all US sensors. This is the main bottleneck.
    \item I have written code to download any arbitrary NAAQS sensor data between two dates. However, some of the NAAQS monitors report on a less frequent basis than hourly, so the analysis code will need to be generalized for other reporting frequencies.
    \item I have written code to download wind velocity data, but parsing the files to get wind velocity at a particular location and time requires more work. This will be used in my predictive model for missing PM2.5 data at the NAAQS monitor.
\end{itemize}

\subsection{Improving Prediction of PM2.5 at the NAAQS Monitor Location} \label{sec:app-prediction}

% \begin{itemize}
%     \item Add short description of the implementation
%     \item Add description of the NOAA NCEP NARR wind data -- u-v components and extraction needed from layer.
% \end{itemize}

A more realistic model of predicting pollution at the EPA monitor could be used. Using wind direction (and possibly wind speed), we can intuitively put more weight on PurpleAir sensors that are North of the EPA monitor when the wind is blowing South. This model allows for mroe flexibility in dropping some sensors that may be measuring hyper-local pollution and are not good predictors of the EPA monitor's PM2.5 readings.

\[
EPA_{i,t} = \gamma_{i,0} + \sum\limits_{j\in J_i}\sum\limits_{k=1}^7\gamma_{i,j,k} 
PA_{j,t} \cdot Winddir_{i,t,k} + u_{i,t}
\]

\begin{itemize}
\item Each EPA monitor $i$ has it's own set of weights for the PA sensors around it.
\item Analysis is done at the quarter level; suppressing quarter subscript.
\item $t$ is a unique hour within a given quarter.
\item EPA monitor $i$ at time $t$ reads PM2.5 pollution $EPA_{i,t}$.
\item For each EPA monitor $i$, there are $J_i$ Purple Air monitors within a 5-mile radius.
\item Purple Air monitor $j\in J_i$ at time $t$ reads PM2.5 pollution $PA_{j,t}$.
\item $Winddir_{i,t,k} $ is a wind direction indicator; 1 if the prevailing wind near station $i$ at time $t$ is in the $k^{th}$ bucket (of 8 buckets).
\item I will also estimate a version with wind speed interacted in the sum. This could allow for sensors further away to have more predictive power when the winds are strong.
\item This regression could be run as a LASSO first to determine which of the interactions for each PurpleAir sensor have the most predictive power.
\end{itemize}

% \newpage
% \subsection{Data Cleaning Steps}
% \input{pics/appendix/tables/data_cleaning_steps}

\newpage
\subsection{Design Values}
\label{sec:design_values}

Notes about design values:
\begin{itemize}
    \item Valid days: a day that has 18 or more valid hours in it.
    \item Valid quarters: a 3-month period that has at least 67. 68, or 69 valid days in it.
    \item Valid 3-year period: a 12-quarter period that has 12 valid quarters in it.
    \item A design value is only valid if its 3-year period is valid.
    \item All averages below assume there might be some data missing.
    \item Design values are based on quarters, so each quarter has a rolling average over the last 3 years before.
\end{itemize}

\textbf{To construct the annual design value for a given quarter:}
\begin{itemize}
    \item Construct an average for every day.
    \item Construct an annual average of daily averages for every previous 12-month period before this quarter.
    \item Construct the 3-year average of those annual averages. Because some years have a different number of real or valid days in them, you cannot take a simple 3-year average of all available days or hours.
\end{itemize}

\textbf{To construct the 24-hour design value:}
\begin{itemize}
    \item Construct an average for every day.
    \item Construct a 98th percentile of all daily averages in the 4-quarter period ending in this quarter. To avoid ambiguous design value construction, the EPA provides a lookup table for which $n^{th}$ maximum daily value to take when constructing the 98th percentile.
    \item Take a 3-year average of the 98th percentiles.
\end{itemize}

\newpage
\subsection{Tables}
\label{sec:tables}

\input{pics/appendix/tables/dv_differences}

\input{pics/appendix/tables/attainment_counties}

\newpage

% \input{pics/appendix/tables/epa_OLS_idw_pa_site-037-4004}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-001-0013}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-019-0500}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-019-5001}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-027-0002}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-029-0018}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-031-0004}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-031-1004}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-039-2010}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-057-0005}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-059-0007}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-067-5003}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-077-2010}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-083-0011}
\input{pics/appendix/tables/epa_OLS_idw_pa_site-103-0007}

\newpage
\input{pics/appendix/tables/excluded_qualifiers_table}

\newpage
\subsection{Pictures of PM2.5 monitors}
\label{sec:monitor_pictures}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{appendix/sensor_pics/air-monitoring-site.jpg}
\caption{A typical NAAQS-primary grade air quality monitoring station.}
\label{fig:pic-epa-site}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{appendix/sensor_pics/PA_outdoor_monitor.png}
\caption{One of PurpleAir's two main outdoor air pollution monitors.}
\label{fig:pic-ps-sensor}
\end{figure}

\FloatBarrier
\newpage
\subsection{Plots for Other California Hourly NAAQS Monitors} \label{app-additional-plots}

\input{z_appendix_site_plots}
\end{document}

### 2022-03-07 acwatt_syp_code/build/calculate_pm.py
Added plot of US epa monitors

@@ -8,8 +8,11 @@ import datetime as dt
 
 import botocore.exceptions
 import pandas as pd
+import geopandas as gpd
+from shapely.geometry import Point
 import numpy as np
 import matplotlib.pyplot as plt
+from pathlib import Path
 import seaborn as sns
 import statsmodels.api as sm
 from statsmodels.iolib.summary2 import summary_col
@@ -24,7 +27,9 @@ from botocore.exceptions import ClientError
 from ..utils.config import PATHS, AWS
 
 logger = logging.getLogger(__name__)
-DTYPES = {"county": str, "site": str, "Qualifier Type Code": str}
+DTYPES = {"county_code": str, "county": str, "County Code": str, "State Code": str,
+          "site_number": str, "site": str, "Site Number": str, "Site Num": str,
+          "Qualifier Type Code": str}
 
 
 def make_hourly_avg_plots(df_pa, df_epa, sensor_id):
@@ -688,6 +693,100 @@ def create_sample_dvs(left='epa', right_list=None):
 
 def generate_predictions(df_):
     pass
+
+
+################################################################################
+#                       Presentation Plots
+################################################################################
+def add_latlon_points(df, crs, lat="Latitude", lon="Longitude"):
+    points = [Point(lon, lat) for lon, lat in zip(df[lon], df[lat])]
+    return gpd.GeoDataFrame(df, geometry=points, crs=crs)
+
+
+def plot_us_epa():
+    """Plot all US EPA NAAQS sensors, mark my sample"""
+    # Plot US shape
+    p1 = PATHS.data.gis / 'cb_2018_us_state_5m' / 'cb_2018_us_state_5m.shp'
+    gdf = gpd.read_file(p1)
+    states_to_drop = ['02', '15']  # remove alaska and hawaii
+    mask = "STATEFP < '60'"  # remove all non-states
+    gdf = gdf[~gdf.STATEFP.isin(states_to_drop)].query(mask)
+    fig = plt.figure()
+    ax = gdf.plot(color='grey', figsize=(15, 8), antialiased=False)
+
+    # Plot all EPA NAAQS monitors in black
+    p2 = PATHS.data.epa_monitors / 'aqs_monitors.csv'
+    df1 = pd.read_csv(p2, dtype=DTYPES)
+    df1 = df1.query("`NAAQS Primary Monitor` == 'Y' and `Parameter Code` == 88101")
+    df1 = df1.query("`Last Sample Date` > '2017-01-01' and `First Year of Data` < 2018")
[…]

### 2022-03-07 run.py

@@ -34,7 +34,8 @@ if __name__ == "__main__":
     # purpleair_download.dl_us_sensors()
     # purpleair_download.test_lambda()
     # lambda_services.save_pa_data_to_s3()
-    # calculate_pm.test_15_sites(run_all=False)
+    # calculate_pm.test_15_sites(run_all=True)
     # calculate_pm.make_plots_15_sites()
-    calculate_pm.create_sample_dvs()
+    # calculate_pm.create_sample_dvs()
+    calculate_pm.generate_presentation_plots()
     logger.info('DONE')

### 2022-03-07 acwatt_syp_code/build/calculate_pm.py
removed unused code from US epa plot

@@ -724,39 +724,6 @@ def plot_us_epa():
     gdf_epa_primary = add_latlon_points(df1, gdf.crs)
     gdf_epa_primary.plot(ax=ax, color='black', markersize=11)
 
-    #### Ended up not using these hourly points because all NAAQS primary monitors
-    #### are labeled as hourly and report the daily average of their hourly measurements
-    #### but not all NAAQS monitors report each hourly measurement to the AQS server,
-    #### they only report the daily average of the hourly measurements. So all monitors
-    #### are labeled as 1 HOUR, but fewer actualy have hourly AQS measurements
-    #### which I only found after downloading the measurements from AQS and counting
-    #### up all the measurements -- something I don't have time to do right now for
-    #### the whole US.
-    # Plot all hourly monitors in black
-    # Create indicator for hourly monitors
-    # Combine daily summaries tables that have "1 HOUR" samply duration indicator
-    # csvs = ['2016', '2017', '2018', '2019']
-    # dir_ = str(PATHS.data.epa_monitors / 'summary_data' / '88101')
-    # dfs = [pd.read_csv(dir_ + f"/daily_88101_{year}.csv") for year in csvs]
-    # keep_cols = ['State Code', 'County Code', 'Site Num', 'Parameter Code', 'Sample Duration']
-    # df_merge = pd.concat(dfs, ignore_index=True)
-    # df_merge = (df_merge
-    #             .query("`Parameter Code` == 88101 and `Sample Duration` == '1 HOUR'")
-    #             .drop_duplicates(keep_cols + ['Date Local']))
-    # df_merge.hist('Observation Count', bins=df_merge['Observation Count'].max()-1)
-    # df_temp = df_merge.query("`Observation Count` > 24")
-    # df_merge = (df_merge[keep_cols]
-    #             .query("`Parameter Code` == 88101 and `Sample Duration` == '1 HOUR'")
-    #             .drop_duplicates()
-    #             .rename(columns={'Site Num': 'Site Number'})
-    #             .assign(hourly=1)[['State Code', 'County Code', 'Site Number', 'hourly']])
-    # # Merge hourly measurement indicator
-    # gdf_epa_hourly = (gdf_epa_primary
-    #                   .merge(df_merge,
-    #                          how='left',
-    #                          on=['State Code', 'County Code', 'Site Number']))
[…]

### 2022-03-07 acwatt_syp_code/build/calculate_pm.py
Added Cali EPA monitor plot

@@ -741,10 +741,58 @@ def plot_us_epa():
     plt.xlim([-124.8, -66.9])
     plt.ylim([24.5, 49.4])
     plt.tight_layout()
+    plt.legend(fontsize=17, frameon=True, facecolor='white')
     p_fig = PATHS.output / 'figures' / 'epa' / f'all_us_and_15_epa_monitors.png'
     plt.savefig(p_fig, dpi=200)
     plt.close(fig)
-    
+
+
+def plot_ca_epa():
+    """Plot all ca EPA NAAQS sensors, mark my sample"""
+    # Plot US shape
+    p1 = PATHS.data.gis / 'cb_2018_us_state_5m' / 'cb_2018_us_state_5m.shp'
+    gdf = gpd.read_file(p1)
+    gdf = gdf.query("STATEFP == '06'")  # keep only CA
+    ax = gdf.plot(color='grey', figsize=(8, 8), edgecolor="face", linewidth=0.4)
+
+    # Plot all EPA NAAQS monitors in black
+    p2 = PATHS.data.epa_monitors / 'aqs_monitors.csv'
+    df1 = pd.read_csv(p2, dtype=DTYPES)
+    df1 = df1.query("`NAAQS Primary Monitor` == 'Y' and `Parameter Code` == 88101")
+    df1 = df1.query("`Last Sample Date` > '2017-01-01' and `First Year of Data` < 2018")
+    df1 = df1.drop_duplicates(['State Code', 'County Code', 'Site Number'])
+    df1 = df1.query("`State Code` == '06'")
+    # 853 unique monitoring sites after this
+    gdf_epa_primary = add_latlon_points(df1, gdf.crs)
+    gdf_epa_primary.plot(ax=ax, color='black', markersize=12)
+
+    # Plot selected EPA NAAQS monitors in red
+    p3 = PATHS.data.epa_monitors / 'aqs_monitors_88101_hourly-ca-monitors.csv'
+    df3 = pd.read_csv(p3, dtype=DTYPES)
+    df3 = (df3
+           .assign(state="06", in_sample="In Sample")
+           .rename(columns={'site_number': 'Site Number',
+                            'county_code': 'County Code',
+                            'state': "State Code"}))
+    gdf_epa_sample = (gdf_epa_primary
+                      .merge(df3,
+                             how='left',
+                             on=['State Code', 'County Code', 'Site Number']))
[…]

### 2022-03-07 acwatt_syp_code/build/calculate_pm.py
Creating PA california plot for all PA vs 15 sample

@@ -10,6 +10,7 @@ import botocore.exceptions
 import pandas as pd
 import geopandas as gpd
 from shapely.geometry import Point
+from shapely import wkt
 import numpy as np
 import matplotlib.pyplot as plt
 from pathlib import Path
@@ -703,6 +704,38 @@ def add_latlon_points(df, crs, lat="Latitude", lon="Longitude"):
     return gpd.GeoDataFrame(df, geometry=points, crs=crs)
 
 
+def load_ca_pa_locations():
+    # Load county shapefile to get CRS for purpleair locations
+    p_shp = PATHS.data.gis / 'cb_2018_us_county_500k' / 'cb_2018_us_county_500k.shp'
+    gdf = gpd.read_file(p_shp)
+    p_pa = PATHS.data.temp / 'sensors_filtered.csv'
+    cols = ['sensor_index', 'date_created', 'lat', 'lon', 'STUSPS', 'geometry']
+    df = pd.read_csv(p_pa, usecols=cols)
+    df['geometry'] = df['geometry'].apply(wkt.loads)
+    gdf = gpd.GeoDataFrame(df, crs=gdf.crs)
+    gdf = gdf.rename(columns={'lat': 'latitude', 'lon': 'longitude', 'STUSPS': 'state'})
+    gdf = gdf.query("state in ['CA']")
+    return gdf
+
+
+def load_ca_pa_insample(min_pa_sensors=5, dist_threshold=5):
+    df = pd.read_csv(PATHS.data.tables / 'epa_pa_lookups' / f'aqs_monitors_to_pa_sensors.csv', dtype=DTYPES)
+    df = df.sort_values(['county', 'site', 'dist_mile'])
+
+    df['in_sample'] = "Other PA Sensors"
+
+    def set_sample(df_):
+        df_ = df_.reset_index(drop=True)
+        if len(df_.query(f'dist_mile < {dist_threshold}')) > min_pa_sensors:
+            df_['in_sample'].iloc[df_['dist_mile'] < dist_threshold] = "In Radius"
+        else:
+            df_.iloc[:min_pa_sensors]['in_sample'] = "In Radius"
+        return df_
+
+    df = df.groupby(['county', 'site']).apply(set_sample).reset_index(drop=True)
+    return df
+
+
 def plot_us_epa():
     """Plot all US EPA NAAQS sensors, mark my sample"""
     # Plot US shape
[…]

### 2022-03-07 z_appendix_site_plots.tex

%=========================================
%  Concentric Radii Maps
%=========================================
\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-001_site-0013_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 0013 in county 001 (FIPS code).}
\label{fig:concentric_purpleair_001-0013}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-019_site-0500_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 0500 in county 019 (FIPS code).}
\label{fig:concentric_purpleair_019-0500}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-019_site-5001_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 5001 in county 019 (FIPS code).}
\label{fig:concentric_purpleair_019-5001}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-027_site-0002_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 0002 in county 027 (FIPS code).}
\label{fig:concentric_purpleair_027-0002}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-029_site-0018_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 0018 in county 029 (FIPS code).}
\label{fig:concentric_purpleair_029-0018}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-031_site-0004_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 0004 in county 031 (FIPS code).}
\label{fig:concentric_purpleair_031-0004}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-031_site-1004_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 1004 in county 031 (FIPS code).}
\label{fig:concentric_purpleair_031-1004}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-039_site-2010_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 2010 in county 039 (FIPS code).}
\label{fig:concentric_purpleair_039-2010}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-057_site-0005_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 0005 in county 057 (FIPS code).}
\label{fig:concentric_purpleair_057-0005}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-059_site-0007_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 0007 in county 059 (FIPS code).}
\label{fig:concentric_purpleair_059-0007}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-067_site-5003_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 5003 in county 067 (FIPS code).}
\label{fig:concentric_purpleair_067-5003}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-077_site-2010_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 2010 in county 077 (FIPS code).}
\label{fig:concentric_purpleair_077-2010}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-083_site-0011_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 0011 in county 083 (FIPS code).}
\label{fig:concentric_purpleair_083-0011}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/county-103_site-0007_epa-pa-concentric-ranges.png}
\caption{Map of EPA NAAQS-primary monitoring station (red) surrounded by PurpleAir monitors within 5-mile (pink), 10-mile (yellow), and 25-mile (green) radii.This preliminary analysis uses the PurpleAir sensors within 5 miles (pink markers). This monitor is at site 0007 in county 103 (FIPS code).}
\label{fig:concentric_purpleair_103-0007}
\end{figure}
%=========================================
%  PurpleAir Hourly Observation Coverage
%=========================================
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-059-0007_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 0007 in county 059 (FIPS code).}
\label{fig:hourly_coverage_059-0007}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-001-0013_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 0013 in county 001 (FIPS code).}
\label{fig:hourly_coverage_001-0013}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-019-0500_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 0500 in county 019 (FIPS code).}
\label{fig:hourly_coverage_019-0500}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-019-5001_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 5001 in county 019 (FIPS code).}
\label{fig:hourly_coverage_019-5001}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-027-0002_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 0002 in county 027 (FIPS code).}
\label{fig:hourly_coverage_027-0002}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-029-0018_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 0018 in county 029 (FIPS code).}
\label{fig:hourly_coverage_029-0018}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-031-1004_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 1004 in county 031 (FIPS code).}
\label{fig:hourly_coverage_031-1004}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-039-2010_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 2010 in county 039 (FIPS code).}
\label{fig:hourly_coverage_039-2010}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-057-0005_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 0005 in county 057 (FIPS code).}
\label{fig:hourly_coverage_057-0005}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-067-5003_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 5003 in county 067 (FIPS code).}
\label{fig:hourly_coverage_067-5003}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-077-2010_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 2010 in county 077 (FIPS code).}
\label{fig:hourly_coverage_077-2010}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-083-0011_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 0011 in county 083 (FIPS code).}
\label{fig:hourly_coverage_083-0011}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-103-0007_pa-daily-covereage.png}
\caption{Scatter plot indicating the number of hours in each day that this NAAQS monitor has PurpleAir coverage. An hour has PurpleAir coverage if there are any PurpleAir sensor readings within the 5-mile radius of the monitor site for that hour. The weighted average is calculated for that hour using all the available PurpleAir readings within 5 miles. This monitor is at site 0007 in county 103 (FIPS code).}
\label{fig:hourly_coverage_103-0007}
\end{figure}
%=========================================
%  PurpleAir EPA Non-missing Comparison
%=========================================
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-027-0002_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 0002 in county 027 (FIPS code).}
\label{fig:pa-epa-compare_027-0002}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-001-0013_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 0013 in county 001 (FIPS code).}
\label{fig:pa-epa-compare_001-0013}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-019-0500_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 0500 in county 019 (FIPS code).}
\label{fig:pa-epa-compare_019-0500}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-019-5001_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 5001 in county 019 (FIPS code).}
\label{fig:pa-epa-compare_019-5001}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-029-0018_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 0018 in county 029 (FIPS code).}
\label{fig:pa-epa-compare_029-0018}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-031-1004_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 1004 in county 031 (FIPS code).}
\label{fig:pa-epa-compare_031-1004}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-039-2010_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 2010 in county 039 (FIPS code).}
\label{fig:pa-epa-compare_039-2010}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-057-0005_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 0005 in county 057 (FIPS code).}
\label{fig:pa-epa-compare_057-0005}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-059-0007_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 0007 in county 059 (FIPS code).}
\label{fig:pa-epa-compare_059-0007}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-067-5003_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 5003 in county 067 (FIPS code).}
\label{fig:pa-epa-compare_067-5003}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-077-2010_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 2010 in county 077 (FIPS code).}
\label{fig:pa-epa-compare_077-2010}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-083-0011_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 0011 in county 083 (FIPS code).}
\label{fig:pa-epa-compare_083-0011}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-103-0007_epa-pa-hourly-plot.png}
\caption{Scatter plot comparing reported hourly PM2.5 measurements: the x-axis represents the IDW-weighted average of PurpleAir measurements, the y-axis represents reported NAAQS-primary monitor measurements. The red line is a 45$^\circ$ line, representing perfect correlation between the PurpleAir average and the NAAQS-primary monitor. This monitor is at site 0007 in county 103 (FIPS code).}
\label{fig:pa-epa-compare_103-0007}
\end{figure}
%=========================================
%  Kernel Density Comparison
%=========================================
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-001-0013_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 0013 in county 001 (FIPS code).}
\label{fig:missing-density_001-0013}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-019-5001_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 5001 in county 019 (FIPS code).}
\label{fig:missing-density_019-5001}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-027-0002_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 0002 in county 027 (FIPS code).}
\label{fig:missing-density_027-0002}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-029-0018_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 0018 in county 029 (FIPS code).}
\label{fig:missing-density_029-0018}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-039-2010_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 2010 in county 039 (FIPS code).}
\label{fig:missing-density_039-2010}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-057-0005_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 0005 in county 057 (FIPS code).}
\label{fig:missing-density_057-0005}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-059-0007_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 0007 in county 059 (FIPS code).}
\label{fig:missing-density_059-0007}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-067-5003_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 5003 in county 067 (FIPS code).}
\label{fig:missing-density_067-5003}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-077-2010_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 2010 in county 077 (FIPS code).}
\label{fig:missing-density_077-2010}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-083-0011_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 0011 in county 083 (FIPS code).}
\label{fig:missing-density_083-0011}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{appendix/site_plots/site-103-0007_epa-pa-missing-density.png}
\caption{Comparison of PM2.5 concentration densities for two sets of hours: reported (blue) and missing (red) hourly observations of the NAAQS monitor. Both densities use the hourly PurpleAir PM2.5 concentration estimates for this site, calculated using the IDW average of PurpleAir sensors within 5 miles of the NAAQS monitor location. This monitor is at site 0007 in county 103 (FIPS code).}
\label{fig:missing-density_103-0007}
\end{figure}

### 2022-03-07 paper/15min-presentation/15min-presentation.tex
Updates from Overleaf

\documentclass{beamer}
\usepackage[utf8]{inputenc}

\usetheme{Madrid}
\usecolortheme{default}

%------------------------------------------------------------
%This block of code defines the information to appear in the
%Title page
\title[Missing Pollution Data] %optional
{Filling in the Gaps: Using Consumer Products to Replace Missing Pollution Data}

% \subtitle{A short story}

\author[Watt, Aaron] % (optional)
{A.~C.~Watt\inst{1}}

\institute[UCB] % (optional)
{
  \inst{1}%
  Agricultural \& Resource Economics\\
  University of California, Berkeley
}

\date[SYP 2022] % (optional)
{Second Year Paper, March 2022}

%---------------------------------------------------------

%---------------------------------------------------------
% Introduction

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection]
  \end{frame}
}
%------------------------------------------------------------

\begin{document}

%The next statement creates the title page.
\frame{\titlepage}

%---------------------------------------------------------
%This block of code is for the table of contents after
%the title page
\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}
%---------------------------------------------------------

\section{First section}

%---------------------------------------------------------
%Changing visivility of the text
\begin{frame}
\frametitle{Sample frame title}
This is a text in second frame. For the sake of showing an example.

\begin{itemize}
    \item<1-> Text visible on slide 1
    \item<2-> Text visible on slide 2
    \item<3> Text visible on slides 3
    \item<4-> Text visible on slide 4
\end{itemize}
\end{frame}

%---------------------------------------------------------

%---------------------------------------------------------
%Example of the \pause command
\begin{frame}
In this slide \pause

the text will be partially visible \pause

And finally everything will be there
\end{frame}
%---------------------------------------------------------

\section{Second section}

%---------------------------------------------------------
%Highlighting text
\begin{frame}
\frametitle{Sample frame title}

In this slide, some important text will be
\alert{highlighted} because it's important.
Please, don't abuse it.

\begin{block}{Remark}
Sample text
\end{block}

\begin{alertblock}{Important theorem}
Sample text in red box
\end{alertblock}

\begin{examples}
Sample text in green box. The title of the block is ``Examples".
\end{examples}
\end{frame}
%---------------------------------------------------------

%---------------------------------------------------------
%Two columns
\begin{frame}
\frametitle{Two-column slide}

\begin{columns}

\column{0.5\textwidth}
This is a text in first column.
$$E=mc^2$$
\begin{itemize}
\item First item
\item Second item
\end{itemize}

\column{0.5\textwidth}
This text will be in the second column
and on a second tought this is a nice looking
layout in some cases.
\end{columns}
\end{frame}
%---------------------------------------------------------

\end{document}

### 2022-03-07 1_introduction.tex

@@
 Section \ref{background} briefly reviews the history of air quality standard in the US and some key details of current regulations.
-Section \ref{data} then discusses the data used and section \ref{emprical} describes the theoretical and empirical framework that will be applied to estimate the missing pollution and effects on reference levels for national standards.
+Section \ref{data} then discusses the data used and section \ref{theoretical} describes the theoretical and empirical framework that will be applied to estimate the missing pollution and effects on reference levels for national standards.
 Section \ref{results} reviews the results of the empirical study and concludes.

### 2022-03-07 paper.tex

@@
 % Research Question
-[abstract here]
+The Clean Air Act (CAA) is one of the largest-scale pollution policies in US history.
+Through the CAA, the National Ambient Air Quality Standards (NAAQS) provide thresholds for a basic minimum standard of air quality in the US.
+However, the EPA only requires 75\% completeness of air quality measurements.
+This minimum standard of completeness provides a plausible mechanism for locations to under-report high pollution to avoid expensive NAAQS penalties, thus biasing the air quality measurements that are compared to the NAAQS.
+This paper explores this issue of bias from missing data by bringing in new consumer-based outdoor air pollution data to fill in the gaps in reported data.
+I find that though most tested locations do not have a significant bias in their reported measurements, one of 15 sites tested has a statistically and economically large bias due to missing data that would be reported under a higher completeness standard.
+While this is a test on a relatively small number of NAAQS sites, this methodology can be employed in future work to examine most or all of US air quality monitors.
 
@@
 
-I also estimate the 95\% lower and upper bound on each predicted value; $\low$ and $\up$ to later using in lower and upper bounds on the design value.
+I also estimate the 95\% lower and upper bound on each predicted value ($\low$ and $\up$) to use later in calculating lower and upper bounds on the design values for each location.
 
@@
 
-These unremarkable results for the annual design value bias are fairly representative of the batch of 15 sites.
-When we look at the hour design value results, we can see in 2020 and 2021, there were fairly large and significant differences between the design values.
+These unremarkable results for the annual design value bias (columns 2-4) are fairly representative of the batch of 15 sites.
+When we look at the hour design value results above (columns 5-7), we can see in 2020 and 2021, there were fairly large and significant differences between the design values.
 This indicates that having more restrictive data collection standards could have increased Fresno's design value significantly in these quarters.
@@
 Is this meaningful for Fresno though?
-Fresno has had a nonattainment status for many years and is listed on several lists within this literature as having notable pollution measurement issues.
+Fresno has had nonattainment status for many years and is listed on several lists within this literature as having notable pollution measurement issues.
 So though Fresno's recorded design value would have been higher, they were already well over the 24-hour design threshold of 35 $\mu$g/m$^3$, having pseudo design values in the 50s and 60s in 2020 and 2021.
@@
     \item I have only tested 15 of 388 possible sites.
-Now that I have written the bulk of the code the manage the data, most of the future work lies in acquiring the rest of the data for the NAAQS monitors and PurpleAir monitors.
+Now that I have written the bulk of the code to manage the data, most of the future work lies in acquiring the rest of the data for the NAAQS monitors and PurpleAir monitors.
     \item I am in the process of negotiating a data usage agreement with PurpleAir staff to get the entire historical dataset of all US sensors. This is the main bottleneck.
-    \item I have written code to download any arbitrary NAAQS sensor data between two dates.
-However, some of the NAAQS monitors report on a less frequent basis than hourly, so the analysis code will need to be generalized for other reporting frequencies.
+    \item Some of the NAAQS monitors report on a less frequent basis than hourly, so the analysis code will need to be generalized for other reporting frequencies.
     \item I have written code to download wind velocity data, but parsing the files to get wind velocity at a particular location and time requires more work.
-This will be used in my predictive model for missing PM2.5 data at the NAAQS monitor.
+This will be used in my predictive model for missing PM2.5 data at the NAAQS monitor (see next section).
 \end{itemize}

### 2022-03-07 paper/15min-presentation/15min-presentation.tex

@@
 \usepackage[utf8]{inputenc}
+\setbeamertemplate{navigation symbols}{}
 
@@
 \begin{frame}
-\frametitle{Table of Contents}
-\tableofcontents
+\frametitle{Motivation}
+\textbf{Clean Air}
+
+\begin{itemize}
+    \item The Clean Air Act (1970) established National Ambient Air Quality Standards
+    (NAAQS) for US counties
+    \item Either "attainment" or "non-attainment", penalties/forced adoption
+    \item Minimum requirement of 75\% of readings, per quarter
+    \item Air quality can change quickly
+    \item Monitor shutoffs are common
+\end{itemize}
+\vspace{2em}
+\textbf{Research Questions}
+\begin{itemize}
+\item How biased are EPA monitor-based measures of local air quality?
+\item Does this bias significantly change NAAQS attainment status?
+\end{itemize}
+
 \end{frame}
@@
 
-\section{First section}
+%---------------------------------------------------------
+\begin{frame}
+\frametitle{EPA National Ambient Air Quality Standards Monitors}
 
-%---------------------------------------------------------
-%Changing visivility of the text
-\begin{frame}
-\frametitle{Sample frame title}
-This is a text in second frame. For the sake of showing an example.
-
-\begin{itemize}
-    \item<1-> Text visible on slide 1
-    \item<2-> Text visible on slide 2
-    \item<3> Text visible on slides 3
-    \item<4-> Text visible on slide 4
-\end{itemize}
+\includegraphics[width=\textwidth]{output/figures/epa/combined_uc_cali_epa.png}
 \end{frame}
@@
 \begin{frame}
+\frametitle{National Ambient Air Quality Standards for PM2.5}
+Two standards for assessing compliance for ambient PM2.5 concentrations at the location of a monitor:
+\vspace{2em}
+
+\textbf{Daily Design Value} (average)
+\begin{itemize}
+    \item[$\Rightarrow$] 3-year rolling average of 1-year average of daily averages
+    \item[$\Rightarrow$] standard currently set at 15.0 $\mu$g/m$^3$
+\end{itemize}
+\vspace{1em}
+
+\textbf{24-Hour Design Value} (98$^{th}$ percentile)
+\begin{itemize}
+    \item[$\Rightarrow$] 3-year rolling average of 1-year 98$^{th}$ percentile of daily averages
+    \item[$\Rightarrow$] standard currently set at 35 $\mu$g/m$^3$
+\end{itemize}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Highlighting text
+\begin{frame}
+\frametitle{NAAQS Completeness Criteria for daily monitors}
+
+\textbf{Valid day:}
+minimum of 75\% hours reported  (18 hours)
+
+\vspace{2em}
+\textbf{Valid quarter:}
+minimum of 75\% valid days  (22-23 days\footnote{fewer for monitors that report less frequent observations})
+
+\vspace{2em}
+\textbf{Valid quarterly design value:}
+all 12 quarters in 3-year period (12 quarters)
+
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Observed EPA Completeness
+\begin{frame}
+\frametitle{Observed Completeness of NAAQS Monitors in sample}
+\includegraphics[width=0.8\textwidth]{example-image-a}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Purple Air Outdoor Monitors
+\begin{frame}
+\frametitle{Purple Air Outdoor Monitors}
+\includegraphics[width=0.48\textwidth]{output/figures/pa/all_ca_and_15_pa_monitors.png}
+\includegraphics[width=0.48]{0500}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Alternate measure of ambient air quality
+\begin{frame}
+\frametitle{Alternate measure of ambient air quality}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Predicting Missing EPA Data
+\begin{frame}
+\frametitle{Predicting Missing EPA Data}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Design Values for an EPA site
+\begin{frame}
+\frametitle{Calculating Design Values for an EPA Site}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Results: Sample EPA Sites
+\begin{frame}
+\frametitle{Results: Sample EPA Sites}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Results: Fresno
+\begin{frame}
+\frametitle{Results: Fresno}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Conclusions & Discussion
+\begin{frame}
+\frametitle{Conclusions \& Discussion}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Future Work
+\begin{frame}
+\frametitle{Future Work}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Extra Tables
+\begin{frame}
+\frametitle{Table: }
+\end{frame}
+%---------------------------------------------------------
+
+\end{document}
+
+%    Examples
+
+%%%%%%%%%%%%%%%%%% multi-slide text
+\begin{frame}
 In this slide \pause
@@
 \end{frame}
-%---------------------------------------------------------
 
-\section{Second section}
-
-%---------------------------------------------------------
-%Highlighting text
+%%%%%%%%%%%%%%%%%% highlighting text and theorem boxes
 \begin{frame}
-\frametitle{Sample frame title}
+\frametitle{NAAQS Completeness Criteria}
 
@@
 \end{frame}
-%---------------------------------------------------------
 
-%---------------------------------------------------------
 %Two columns
@@
 \end{frame}
-%---------------------------------------------------------
-
-\end{document}

### 2022-03-07 paper/15min-presentation/15min-presentation.tex
Updates from Overleaf

@@
 \frametitle{Purple Air Outdoor Monitors}
-\includegraphics[width=0.48\textwidth]{output/figures/pa/all_ca_and_15_pa_monitors.png}
-\includegraphics[width=0.48]{0500}
+\hspace{1em}
+Purple Air Sensors in California 
+\hspace{4em}
+LA Site Example \\
+
+\includegraphics[width=0.49\textwidth]{output/figures/pa/all_ca_and_15_pa_monitors.png}
+\includegraphics[width=0.49\textwidth]{output/figures/concentric_ranges/county-037_site-4004_epa-pa-concentric-ranges.png}
 \end{frame}

### 2022-03-07 paper/15min-presentation/15min-presentation.tex
Adding _short OLS table for presentation

@@
 \begin{frame}
-\frametitle{Purple Air Outdoor Monitors}
+\frametitle{PurpleAir Outdoor Monitors}
 \hspace{1em}
-Purple Air Sensors in California 
+PurpleAir Sensors in California 
 \hspace{4em}
@@
 \begin{frame}
-\frametitle{Alternate measure of ambient air quality}
+\frametitle{Alternate measure of ambient PM2.5 Concentration}
+
+\hspace{7em}
+\includegraphics[width=0.5\textwidth]{output/figures/diagrams/IDW_diagram.png}
+\\
+\textbf{Inverse-distance Weighted Average Ambient PM2.5}
+
+\begin{equation*}
+    PA^{IDW}_t= \sum\limits_{j=1}^{J_t} \dfrac{\frac{1}{d_j}\cdot PA_{j,t}}{\sum\limits_j^{J_t}\frac{1}{d_j}}
+    = \sum\limits_{j=1}^{J^t} w_{j,t}\cdot PA_{j,t}
+\end{equation*}
+\begin{itemize}
+    \item $J_t$ = active PurpleAir sensors around the NAAQS monitor at time $t$
+\end{itemize}
+
 \end{frame}
@@
 \frametitle{Predicting Missing EPA Data}
+\begin{equation*}
+EPA_t = \beta_0 + \beta_1 PA^{IDW}_t + \varepsilon_t
+\end{equation*}
+
+\input{pics/appendix/tables/epa_OLS_idw_pa_site-037-4004_short}
 \end{frame}
@@
 \frametitle{Future Work}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+% Purple Air EPA Correction
+\begin{frame}
+\frametitle{Correction of PurpleAir Readings}
+$$
+\widetilde{PA}_{j,t}=\begin{cases}
+			0.52*PA_{j,t} - 0.086*H_{j,t} + 5.75, & \text{if $PA_{j,t} \leq 343 \mu$g/m$^3$}\\
+            0.46*PA_{j,t} + 0.(3.93e-4)PA_{j,t}^2 + 2.97, & \text{otherwise}
+		 \end{cases}
+$$
+\begin{itemize}
+    \item $PA_{j,t}$ = ambient PM2.5 measured by PurpleAir sensor $j$ at time $t$
+    \item $H_{j,t}\in[0,1]$ is the relative humidity
+\end{itemize}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+% Future Prediction Equation
+\begin{frame}
+\frametitle{Better Prediction of EPA PM2.5}
+
+$$
+\widetilde{PA}_{j,t}=\begin{cases}
+			0.52*PA_{j,t} - 0.086*H_{j,t} + 5.75, & \text{if $PA_{j,t} \leq 343 \mu$g/m$^3$}\\
+            0.46*PA_{j,t} + 0.(3.93e-4)PA_{j,t}^2 + 2.97, & \text{otherwise}
+		 \end{cases}
+$$
+\begin{itemize}
+    \item $PA_{j,t}$ = ambient PM2.5 measured by PurpleAir sensor $j$ at time $t$
+    \item $H_{j,t}\in[0,1]$ is the relative humidity
+\end{itemize}
 \end{frame}

### 2022-03-07 paper/15min-presentation/15min-presentation.tex
Updates from Overleaf

@@
 \frametitle{Calculating Design Values for an EPA Site}
+
+\def\dvp{\text{DV}_p}
+\def\dvpp{\widetilde{\text{DV}}_p}
+
+\textbf{Pseudo Design Values}\\[0.5em]
+$\dvp$ = $p$ design value for quarter, calculated using \textbf{reported} PM2.5 data from EPA monitor, $\forall p \in \{\text{Daily, 24-Hour}\}$
+\\[1em]
+
+\textbf{Imputed Design Values}\\[0.5em]
+$\dvpp$ = $p$ design value for quarter, calculated using \textbf{imputed} PM2.5 data from EPA monitor and PA sensors
+\\[2em]
+\textbf{Bias in Design Values}
+\begin{equation*}
+    \text{bias from missing data in }\dvp \approx \dvpp - \dvp
+\end{equation*}
+
 \end{frame}
@@
 \frametitle{Results: Sample EPA Sites}
+\includegraphics[width=\textwidth]{}
 \end{frame}
@@
 \frametitle{Results: Fresno}
+\begin{table}[]
+\caption{Design Value Comparison for Fresno, CA. (98\% CI Bounds)}
+\label{tab:fresno_dvs}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{@{}l|lll|lll@{}}
+Year-Quarter & \begin{tabular}[c]{@{}l@{}}Annual DV\\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower\\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Hour DV \\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower \\ Bound\end{tabular} \\ \midrule
+2018-4 & Invalid DV &  &  & Invalid DV &  &  \\
+2019-1 & -0.005 & 0.133 & -0.143 & 0.000 & 0.252 & 0.000 \\
+2019-2 & -0.003 & 0.141 & -0.147 & 0.000 & 0.252 & 0.000 \\
+2019-3 & 0.015 & 0.170 & -0.139 & 0.058 & 2.202 & 0.000 \\
+2019-4 & 0.002 & 0.190 & -0.185 & 0.000 & 1.460 & -0.024 \\
+2020-1 & -0.012 & 0.182 & -0.207 & 0.000 & 0.335 & 0.000 \\
+2020-2 & -0.010 & 0.191 & -0.211 & 0.000 & 0.376 & 0.000 \\
+2020-3 & 0.679 & 0.954 & 0.403 & 8.718 & 11.704 & 8.556 \\
+2020-4 & 0.647 & 1.006 & 0.288 & 5.979 & 8.281 & 5.851 \\
+2021-1 & 0.564 & 1.036 & 0.091 & 3.007 & 4.184 & 2.903 \\
+2021-2 & 0.533 & 1.024 & 0.042 & 3.007 & 4.225 & 2.903 \\
+2021-3 & 0.630 & 1.129 & 0.132 & 7.607 & 10.557 & 7.444
+\end{tabular}%
+}
+\end{table}
 \end{frame}

### 2022-03-07 paper/15min-presentation/15min-presentation.tex

@@
 \begin{frame}
-\frametitle{Results: Sample EPA Sites}
-\includegraphics[width=\textwidth]{}
+\frametitle{Results for Daily Design Value: Sample EPA Sites}
+\includegraphics[width=\textwidth]{output/figures/final_results/DV_plot_all_test_sites.png}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Results: Fresno}
-\begin{table}[]
-\caption{Design Value Comparison for Fresno, CA. (98\% CI Bounds)}
-\label{tab:fresno_dvs}
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{@{}l|lll|lll@{}}
-Year-Quarter & \begin{tabular}[c]{@{}l@{}}Annual DV\\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower\\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Hour DV \\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower \\ Bound\end{tabular} \\ \midrule
-2018-4 & Invalid DV &  &  & Invalid DV &  &  \\
-2019-1 & -0.005 & 0.133 & -0.143 & 0.000 & 0.252 & 0.000 \\
-2019-2 & -0.003 & 0.141 & -0.147 & 0.000 & 0.252 & 0.000 \\
-2019-3 & 0.015 & 0.170 & -0.139 & 0.058 & 2.202 & 0.000 \\
-2019-4 & 0.002 & 0.190 & -0.185 & 0.000 & 1.460 & -0.024 \\
-2020-1 & -0.012 & 0.182 & -0.207 & 0.000 & 0.335 & 0.000 \\
-2020-2 & -0.010 & 0.191 & -0.211 & 0.000 & 0.376 & 0.000 \\
-2020-3 & 0.679 & 0.954 & 0.403 & 8.718 & 11.704 & 8.556 \\
-2020-4 & 0.647 & 1.006 & 0.288 & 5.979 & 8.281 & 5.851 \\
-2021-1 & 0.564 & 1.036 & 0.091 & 3.007 & 4.184 & 2.903 \\
-2021-2 & 0.533 & 1.024 & 0.042 & 3.007 & 4.225 & 2.903 \\
-2021-3 & 0.630 & 1.129 & 0.132 & 7.607 & 10.557 & 7.444
-\end{tabular}%
-}
-\end{table}
+\frametitle{Results for Daily Design Value: Fresno}
+
+\includegraphics[width=\textwidth]{output/figures/final_results/DV_plot_site_019-0500.png}
+
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Results: Sample EPA Sites
+\begin{frame}
+\frametitle{Results for 24-hour Design Value: Sample EPA Sites}
+\includegraphics[width=\textwidth]{output/figures/final_results/DV}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
+%Results: Fresno
+\begin{frame}
+\frametitle{Results for 24-hour Design Value: Fresno}
+
+\includegraphics[width=\textwidth]{output/figures/final_results/DV_plot_site_019-0500.png}
+
 \end{frame}
@@
 \frametitle{Conclusions \& Discussion}
+\begin{enumerate}
+    \item Most sites tested show little evidence of bias from missing data, but some have large, meaningful bias
+\end{enumerate}
 \end{frame}
@@
 %---------------------------------------------------------
+% Fresno Site Data
+\begin{frame}
+\frametitle{Fresno Site Data}
+
+\hspace{5em}
+\includegraphics[width=0.8\textwidth]{pics/appendix/site_plots/site-019-0500_epa-pa-hourly-plot.png}
+\end{frame}
+%---------------------------------------------------------
+
+%---------------------------------------------------------
 %Extra Tables
 \begin{frame}
-\frametitle{Table: }
+\frametitle{Table: Fresno, California}
+
+\begin{table}[]
+\caption{Design Value Comparison for Fresno, CA. (98\% CI Bounds)}
+\label{tab:fresno_dvs}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{@{}l|lll|lll@{}}
+Year-Quarter & \begin{tabular}[c]{@{}l@{}}Annual DV\\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower\\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Hour DV \\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower \\ Bound\end{tabular} \\ \midrule
+2018-4 & Invalid DV &  &  & Invalid DV &  &  \\
+2019-1 & -0.005 & 0.133 & -0.143 & 0.000 & 0.252 & 0.000 \\
+2019-2 & -0.003 & 0.141 & -0.147 & 0.000 & 0.252 & 0.000 \\
+2019-3 & 0.015 & 0.170 & -0.139 & 0.058 & 2.202 & 0.000 \\
+2019-4 & 0.002 & 0.190 & -0.185 & 0.000 & 1.460 & -0.024 \\
+2020-1 & -0.012 & 0.182 & -0.207 & 0.000 & 0.335 & 0.000 \\
+2020-2 & -0.010 & 0.191 & -0.211 & 0.000 & 0.376 & 0.000 \\
+2020-3 & 0.679 & 0.954 & 0.403 & 8.718 & 11.704 & 8.556 \\
+2020-4 & 0.647 & 1.006 & 0.288 & 5.979 & 8.281 & 5.851 \\
+2021-1 & 0.564 & 1.036 & 0.091 & 3.007 & 4.184 & 2.903 \\
+2021-2 & 0.533 & 1.024 & 0.042 & 3.007 & 4.225 & 2.903 \\
+2021-3 & 0.630 & 1.129 & 0.132 & 7.607 & 10.557 & 7.444
+\end{tabular}%
+}
+\end{table}
 \end{frame}

### 2022-03-07 paper/15min-presentation/15min-presentation.tex
Updates from Overleaf

@@
 \frametitle{Results for Daily Design Value: Sample EPA Sites}
-\includegraphics[width=\textwidth]{output/figures/final_results/DV_plot_all_test_sites.png}
+\includegraphics[width=\textwidth]{output/figures/final_results/DV_annual_plot_all_test_sites.png}
+
+{\small Shaded regions are 95\% confidence intervals from interpolating the data.}
 \end{frame}
@@
 
-\includegraphics[width=\textwidth]{output/figures/final_results/DV_plot_site_019-0500.png}
+\includegraphics[width=\textwidth]{output/figures/final_results/DV_annual_plot_site_019-0500.png}
 
+{\small Shaded regions are 95\% confidence intervals from interpolating the data.}
 \end{frame}
@@
 \frametitle{Results for 24-hour Design Value: Sample EPA Sites}
-\includegraphics[width=\textwidth]{output/figures/final_results/DV}
+\includegraphics[width=\textwidth]{output/figures/final_results/DV_hour_plot_all_test_sites.png}
+
+{\small Shaded regions are 95\% confidence intervals from interpolating the data.}
 \end{frame}
@@
 
-\includegraphics[width=\textwidth]{output/figures/final_results/DV_plot_site_019-0500.png}
+\includegraphics[width=\textwidth]{output/figures/final_results/DV_hour_plot_site_019-0500.png}
 
+{\small Shaded regions are 95\% confidence intervals from interpolating the data.}
 \end{frame}
@@
 \begin{enumerate}
-    \item Most sites tested show little evidence of bias from missing data, but some have large, meaningful bias
+\setlength\itemsep{2em}
+    \item Most tested sites show little evidence of bias from missing data, but one has large, meaningful bias
+    \item Even one site can affect millions of people due to the sparsity of monitoring sites
+    \item As high-pollution locations continue to reduce pollution, this bias may play an important role in determining NAAQS compliance
+    \item Underlines importance of expanding the monitor network or exploring alternative measures of ambient air quality
 \end{enumerate}
@@
 \frametitle{Future Work}
+
+\begin{itemize}
+\setlength\itemsep{2em}
+    \item Optimal regulation of ambient pollution under monitor expense-accuracy tradeoff
+    \item Expand test to rest of US monitors
+    \item Explore spatial distribution of air quality in unmonitored locations
+\end{itemize}
+
 \end{frame}
@@
 \begin{frame}
-\frametitle{Correction of PurpleAir Readings}
+\frametitle{Appendix: Correction of PurpleAir Readings}
 $$
@@
 \begin{frame}
-\frametitle{Better Prediction of EPA PM2.5}
+\frametitle{Appendix: Better Prediction of EPA PM2.5}
 
@@
 \begin{frame}
-\frametitle{Fresno Site Data}
+\frametitle{Appendix: Fresno Site Data}
 
@@
 \begin{frame}
-\frametitle{Table: Fresno, California}
-
+\frametitle{Appendix: Fresno, California DV Bias Table}
+\vspace{-1em}
 \begin{table}[]
-\caption{Design Value Comparison for Fresno, CA. (98\% CI Bounds)}
+\caption{Design Value Comparison for Fresno, CA. (95\% CI Bounds)}
 \label{tab:fresno_dvs}

### 2022-03-07 acwatt_syp_code/build/calculate_pm.py
Added completeness plots

@@ -22,6 +22,7 @@ import time
 import os
 import io
 # Third-party Imports
+import altair as alt
 import boto3
 from botocore.exceptions import ClientError
 # Local Imports
@@ -625,28 +626,57 @@ def annual_data(df_daily: pd.DataFrame, pm_type: str):
     return pd.concat(df_list, ignore_index=True)
 
 
+def save_hourly_completeness(df: pd.DataFrame, site_dict):
+    # Hour Stats
+    total = len(df)
+    d = {'type': 'hour',
+         'total': total,
+         'complete': df['pm2.5_epa'].notna().sum(),
+         'complete_p': df['pm2.5_epa'].notna().sum() / total,
+         'complete_after': df['pm2.5_epa.olsyc.pa'].notna().sum(),
+         'complete_after_p': df['pm2.5_epa.olsyc.pa'].notna().sum() / total}
+    d.update({'filled': d['complete_after'] - d['complete']})
+    d.update(site_dict)
+    return d
+
+
+def save_daily_completeness(df, site_dict):
+    # Daily Stats
+    total = len(df)
+    d = {'type': 'daily',
+         'total': total,
+         'complete': df['epa_valid_daily'].sum(),
+         'complete_p': df['epa_valid_daily'].sum() / total,
+         'complete_after': df['epa.olsyc.pa_valid_daily'].sum(),
+         'complete_after_p': df['epa.olsyc.pa_valid_daily'].sum() / total}
+    d.update({'filled': d['complete_after'] - d['complete']})
+    d.update(site_dict)
+    return d
+
+
 def create_site_dvs(site_dict):
+    completeness_list = []
     # Load combined site data
     df = load_combined(site_dict)
     # Create qualifier / exceptional event indicator
     df = add_exceptional_indicator(df)
-    # Create Summary Statistics
-
     # Drop Exceptional Hours
     df = df.query("drop_qualifier == False")
     # Make new combined EPA-PA column with IDW avereage PA data
     df = fill_in_missing_with_idw(df)
     # Make new combined EPA-PA column with OLS prediction from IDW PA data
     df = fill_in_missing_with_OLS(df, site_dict)
+    completeness_list.append(save_hourly_completeness(df.copy(deep=True), site_dict))
     # Make Daily dataset (with indicators for valid > 75% complete)
[…]

### 2022-04-03 acwatt_syp_code/build/calculate_pm.py
documentation, naming

@@ -364,7 +364,7 @@ def load_qualifiers():
     return pd.read_csv(p, converters={'Qualifier Type Code' : str})[['Qualifier Code', 'Qualifier Type Code']]
 
 
-def test_15_sites(run_all=False):
+def combine_15_sites(run_all=False):
     threshold = 5  # miles
     idw_power = 1  # Inv Distance Weighting Denominator Exponent
     min_sensors = 10  # min # of PA sensors to grab near EPA monitor
@@ -590,7 +590,7 @@ def calculate_design_values(df_daily: pd.DataFrame, quarters: list, pm_type: str
     years = [quarters[0:4], quarters[4:8], quarters[8:12]]  # list of list of four year-quarters
     dva, dvh = [], []
     for year in years:  # list of four year-quarters
-        # filter to these years
+        # filter to these year-quarters
         df_temp = df[df.year_quarter.isin(year)]
         # Get design values for PM2.5 column
         agg_dict = {f"pm2.5_{pm_type}_daily": [dv_annual, dv_hour]}
@@ -656,15 +656,17 @@ def save_daily_completeness(df, site_dict):
 
 def create_site_dvs(site_dict):
     completeness_list = []
-    # Load combined site data
+    # Load combined site data (loads 'epa', 'pa' pm2.5 columns)
     df = load_combined(site_dict)
     # Create qualifier / exceptional event indicator
     df = add_exceptional_indicator(df)
     # Drop Exceptional Hours
     df = df.query("drop_qualifier == False")
     # Make new combined EPA-PA column with IDW avereage PA data
+    # adds 'epa.idw.pa' pm2.5 column
     df = fill_in_missing_with_idw(df)
     # Make new combined EPA-PA column with OLS prediction from IDW PA data
+    # adds 'epa.olsnc.pa', 'epa.olsyc.pa', 'epa.olsyc.pa.lower', 'epa.olsyc.pa.upper' pm2.5 columns
     df = fill_in_missing_with_OLS(df, site_dict)
     completeness_list.append(save_hourly_completeness(df.copy(deep=True), site_dict))
     # Make Daily dataset (with indicators for valid > 75% complete)
@@ -675,7 +677,8 @@ def create_site_dvs(site_dict):
     for pm_type in ['epa', 'pa', 'epa.idw.pa', 'epa.olsnc.pa', 'epa.olsyc.pa', 'epa.olsyc.pa.lower', 'epa.olsyc.pa.upper']:
[…]

### 2022-04-03 acwatt_syp_code/build/calculate_pm.py
Starting code for minimum possible DVs (max possible manipulation)

@@ -732,6 +732,63 @@ def generate_predictions(df_):
 
 
 ################################################################################
+#                       REFEREE RESPONSES
+################################################################################
+##############################################
+#    Minimum DVs (max possible manipulation)
+##############################################
+def create_minimum_dvs():
+    # Load the county-site pairs
+    aqs_tbl = load_15_sites()
+    diffs_list, dv_list, complete_list = [], [], []
+    # For each EPA site-county in list
+    for county, site in aqs_tbl:  # cs_list
+        site_dict = {'county': county, 'site': site}
+        df, completeness_list = create_minimum_site_dvs(site_dict)
+        diffs_list.append(generate_differences(df, left=left, right_list=right_list))
+        dv_list.append(df)
+        complete_list += completeness_list
+    df_dv = pd.concat(dv_list, ignore_index=True)
+    df_dv.to_csv(PATHS.data.temp / 'design_value_est.csv', index=False)
+    df_save = pd.concat(diffs_list, ignore_index=True)
+    df_save['invalid quarter DV due to too many missing days'] = df_save.isnull().any(axis=1)
+    df_save.to_csv(PATHS.data.temp / 'design_value_differences.csv', index=False)
+    agg_dict = {f'annual_dv_diff_{right}': ['mean', 'std', 'max', 'min'] for right in right_list}
+    agg_dict.update({f'hour_dv_diff_{right}': ['mean', 'std', 'max', 'min'] for right in right_list})
+    agg_dict.update({'invalid quarter DV due to too many missing days': ['mean', 'size']})
+    df_stats = df_save.groupby(['county', 'site']).agg(agg_dict)
+    df_stats.reset_index().to_csv(PATHS.data.temp / 'design_value_site-stats.csv', index=False)
+    df_complete = pd.DataFrame(complete_list)
+    df_complete.to_csv(PATHS.data.temp / 'completeness_stats.csv')
+
+
+def create_minimum_site_dvs(site_dict):
+    completeness_list = []
+    # Load combined site data (loads 'epa', 'pa' pm2.5 columns)
+    df = load_combined(site_dict)
+    # Drop highest 6 hours from all days
+    df = drop_highest_6_hours(df)
+    # Make Daily dataset (with indicators for valid > 75% complete)
+    df_daily = daily_data(df)
+    completeness_list.append(save_daily_completeness(df_daily.copy(deep=True), site_dict))
+    # Calculate DVs for all quarters
+    df_list = []
[…]

### 2022-04-03 run.py

@@ -34,8 +34,9 @@ if __name__ == "__main__":
     # purpleair_download.dl_us_sensors()
     # purpleair_download.test_lambda()
     # lambda_services.save_pa_data_to_s3()
-    # calculate_pm.test_15_sites(run_all=True)
+    # calculate_pm.combine_15_sites(run_all=True)
     # calculate_pm.make_plots_15_sites()
     # calculate_pm.create_sample_dvs()
+    calculate_pm.create_minimum_dvs()
     calculate_pm.generate_presentation_plots()
     logger.info('DONE')

### 2022-04-04 acwatt_syp_code/build/calculate_pm.py
Adding plots for referee questions.

@@ -442,6 +442,18 @@ def fill_in_missing_with_OLS(df, site_dict, alpha=0.05):
     df['pm2.5_epa.olsyc.pa.upper'] = np.where(df['pm2.5_epa'].isna(), df['upper.yc'], df['pm2.5_epa'])
     df['pm2.5_epa.olsyc.pa.lower'] = np.where(df['pm2.5_epa'].isna(), df['lower.yc'], df['pm2.5_epa'])
 
+    # Add columns for conservative multiple hypothesis testing significance (referee)
+    alpha_conservative = 0.00656
+    k = 8; M = 198  # k approximate number of rejected hypotheses out of M total tested
+    alpha_conservative2 = (alpha / (1+alpha)) * (k / (M-k))  # 0.002..
+    frame_conservative = predictions.summary_frame(alpha=alpha_conservative2)
+    # Get prediction confidence interval upper and lower bounds for each observations
+    df['upper.yc.conservative'] = frame_conservative.obs_ci_upper  # Full set of upper bounds
+    df['lower.yc.conservative'] = frame_conservative.obs_ci_lower  # Full set of lower bounds
+    # Create two more combined sets of EPA PM, filling in upper and lower predictions
+    df['pm2.5_epa.olsyc.pa.upper.conservative'] = np.where(df['pm2.5_epa'].isna(), df['upper.yc.conservative'], df['pm2.5_epa'])
+    df['pm2.5_epa.olsyc.pa.lower.conservative'] = np.where(df['pm2.5_epa'].isna(), df['lower.yc.conservative'], df['pm2.5_epa'])
+
     # Save the regression summaries
     c_s = f'{site_dict["county"]}-{site_dict["site"]}'
     stargazer = Stargazer([model1, model2])
@@ -667,6 +679,7 @@ def create_site_dvs(site_dict):
     df = fill_in_missing_with_idw(df)
     # Make new combined EPA-PA column with OLS prediction from IDW PA data
     # adds 'epa.olsnc.pa', 'epa.olsyc.pa', 'epa.olsyc.pa.lower', 'epa.olsyc.pa.upper' pm2.5 columns
+    # also adds 'pm2.5_epa.olsyc.pa.upper.conservative' and 'pm2.5_epa.olsyc.pa.lower.conservative' (referee)
     df = fill_in_missing_with_OLS(df, site_dict)
     completeness_list.append(save_hourly_completeness(df.copy(deep=True), site_dict))
[…]

### 2022-04-04 run.py

@@ -36,7 +36,7 @@ if __name__ == "__main__":
     # lambda_services.save_pa_data_to_s3()
     # calculate_pm.combine_15_sites(run_all=True)
     # calculate_pm.make_plots_15_sites()
-    # calculate_pm.create_sample_dvs()
-    calculate_pm.create_minimum_dvs()
+    calculate_pm.create_sample_dvs()
+    # calculate_pm.create_minimum_dvs()
     calculate_pm.generate_presentation_plots()
     logger.info('DONE')

### 2022-04-04 A_faculty_feedback.tex

\textbf{Comment 1.} The economic story behind an empirical observation is most often (though not always) a story of incentives. You describe data anomalies, but without setting the context in terms of incentives. Critically, we would like to know more if possible about who has the discretion to decide what data to omit, and what their incentives are. The introduction of the paper presumes a familiarity with local actor incentives to stay in attainment, but this may be assuming too much for a general economic audience, as opposed to someone specializing in air pollution. So, this comment pertains both to explaining the importance of attainment status and who the key stakeholders are to the broader audience, and offering additional insight on the details of who the “data managers” are, who hires them, and whether they are plausibly subject to pressure.

\noindent\textbf{Response to Comment 1.}

\noindent\textbf{Comment 2.} Another exercise would be to ask how much room is left for additional manipulation. That is, suppose a data controller cared about boosting numbers and wanted to drop as many data points as would be allowed. Given the data that are being reported and the slack in the number dropped, how much of a difference would additional manipulation make? This is sort of a scale exercise for asking how useful manipulation might be.

\noindent\textbf{Response to Comment 2.}

\noindent\textbf{Comment 3.} A related big picture comment is to ask why the objective is to infer an estimate for the missing data, rather than simply look for evidence of manipulation. (To be sure, they are related, but we think they are not identical.) You might have begun, for example, by predicting whether an EPA data point was missing (binary outcome), and asking whether high purple air readings make it more likely an observation is missing.

More generally, one might argue that this paper is a very interesting measurement exercise in search of an economic application. The possibility of flipping attainment status is one way to turn this into an economics/policy narrative, but you might also consider something more broad, like stating that “effective regulation requires reliable measurement,” and then asking if the measure here is manipulable. Obviously we know that there are other types of manipulation (from prior literature), so your contribution is to look for the possibility of manipulation in this particular mechanism.

\noindent\textbf{Comment 4.}  Isn't a kernel regression the right approach here?  Asymptotic performance of current approach (inverse distance) may not work?

\noindent\textbf{Comment 5.} First table seems to involve regression to the mean (Purple Air a rhs variable, with measurement error).

\noindent\textbf{Comment 6.} Multiple testing (however many sites...)?  If drawing inference that a *particular* site is engaged in manipulation, current version of inference seems problematic.

\noindent\textbf{Comment  7.} Confidence intervals for non-compliant site highly asymmetric---not clear why?

\noindent\textbf{Response to Comment 7.} A brief explanation of the asymmetry of the confidence intervals has been added to section \ref{results} and a detailed discussion has been added to appendix section \ref{sec:design_values}.

### 2022-04-04 paper.tex

@@
 \usepackage{graphicx}
+\usepackage[table,xcdraw]{xcolor}
 \usepackage{placeins} % \FloatBarrier
@@
 If this quantity is positive, then there is support that there is under-reporting of pollution via allowed missing data.
-To gain an understanding of significance of the results, I also propagate the upper and lower bounds of the predicted observations through the design value calculation (similar to \cite{fowlieBringingSatelliteBasedAir2019}, but they were estimating out of sample prediction errors).
-This provides a 95\% confidence interval -- upper and lower bounds on the imputed design values: $\dvaa^{upper}$,  $\dvaa^{lower}$, $\dvhh^{upper}$, and $\dvhh^{lower}$.
 
@@
 
+\subsection{Confidence Intervals \& Inference} \label{confidence_intervals}
+To gain an understanding of significance of the results, I also propagate the upper and lower bounds of the predicted observations through the design value calculation (similar to \cite{fowlieBringingSatelliteBasedAir2019}, but they were estimating out of sample prediction errors).
+Propagating the bounds of the prediction intervals this way provides a confidence interval -- upper and lower bounds on the imputed design values: $\dvaa^{upper}$,  $\dvaa^{lower}$, $\dvhh^{upper}$, and $\dvhh^{lower}$.
+
+\textbf{Multiple Hypothesis Testing and Significance Levels.} Suppose I am interested in confidence intervals at a level of significance $\alpha$.
+If I simply propagate prediction intervals of level $\alpha$ through the design value calculation, I would end up with $\alpha$-level confidence intervals that are valid for single-hypothesis inference -- overstating the significance of the results and underestimating the true $\alpha$-level confidence intervals.
+In total, I am conducting 198 different hypothesis tests, testing each of 15 sites for their two design values on up to 12 quarters.
+Following \cite{benjamini_adaptive_2006}, I use their two-stage procedure for selecting the correct significance level of the prediction intervals to propagate through the design value calculation, the correct level that will represent an $\alpha$-level confidence interval given the multiplicity of hypotheses.\footnote{See \cite{anderson_multiple_2008} for an helpful description of applying the two-stage procedure.}
+
+For this paper, I present 95\% confidence intervals (0.05 significance level), which correspond to a corrected significance level of approximately 0.002 applied to the prediction intervals and propagated through the design value calculation.
+See Appendix Section \ref{sec:multiple_hypothesis} for a full description of the iterative procedure used to attain the corrected significance level.
+
 %==============================================
@@
 I conducted 15 design value tests on 15 different NAAQS monitors.
-To create the predicted design value for each NAAQS monitor, I regressed the NAAQS monitor PM2.5 readings on the weighted average PurpleAir readings.
+To create the predicted design values for each NAAQS monitor, I regressed the NAAQS monitor PM2.5 readings on the weighted average PurpleAir readings.
 An example of this regression is below for one of the Los Angeles monitors.
-I ran regions both with an without a constant, but because I wanted the prediction errors to have mean zero, I chose to use model (2) in the creation of the design value.
-Note that the $R^2$ value in model (1) is much higher, however $R^2$ values between regressions with constant and those without are no comparable due to the difference in denominators.
+I ran regressions both with and without a constant, but because I wanted the prediction errors to have mean zero, I chose to use model (2) in the creation of the design values.
+Note that the $R^2$ value in model (1) is much higher, however $R^2$ values between regressions with constant and those without are not comparable due to the difference in denominators.
 \input{pics/appendix/tables/epa_OLS_idw_pa_site-037-4004}
@@
 
-\begin{table}[]
-\caption{Design Value Comparison for Fresno, CA. (98\% CI Bounds)}
-\label{tab:fresno_dvs}
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{@{}l|lll|lll@{}}
-Year-Quarter & \begin{tabular}[c]{@{}l@{}}Annual DV\\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower\\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Hour DV \\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower \\ Bound\end{tabular} \\ \midrule
-2018-4 & Invalid DV &  &  & Invalid DV &  &  \\
-2019-1 & -0.005 & 0.133 & -0.143 & 0.000 & 0.252 & 0.000 \\
-2019-2 & -0.003 & 0.141 & -0.147 & 0.000 & 0.252 & 0.000 \\
-2019-3 & 0.015 & 0.170 & -0.139 & 0.058 & 2.202 & 0.000 \\
-2019-4 & 0.002 & 0.190 & -0.185 & 0.000 & 1.460 & -0.024 \\
-2020-1 & -0.012 & 0.182 & -0.207 & 0.000 & 0.335 & 0.000 \\
-2020-2 & -0.010 & 0.191 & -0.211 & 0.000 & 0.376 & 0.000 \\
-2020-3 & 0.679 & 0.954 & 0.403 & 8.718 & 11.704 & 8.556 \\
-2020-4 & 0.647 & 1.006 & 0.288 & 5.979 & 8.281 & 5.851 \\
-2021-1 & 0.564 & 1.036 & 0.091 & 3.007 & 4.184 & 2.903 \\
-2021-2 & 0.533 & 1.024 & 0.042 & 3.007 & 4.225 & 2.903 \\
-2021-3 & 0.630 & 1.129 & 0.132 & 7.607 & 10.557 & 7.444
-\end{tabular}%
-}
-\end{table}
+% \begin{table}[]
+% \caption{Design Value Comparison for Fresno, CA. (98\% CI Bounds)}
+% \label{tab:fresno_dvs}
+% \resizebox{\textwidth}{!}{%
+% \begin{tabular}{@{}l|lll|lll@{}}
+% Year-Quarter & \begin{tabular}[c]{@{}l@{}}Annual DV\\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower\\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Hour DV \\ Difference\end{tabular} & \begin{tabular}[c]{@{}l@{}}Upper \\ Bound\end{tabular} & \begin{tabular}[c]{@{}l@{}}Lower \\ Bound\end{tabular} \\ \midrule
+% 2018-4 & Invalid DV &  &  & Invalid DV &  &  \\
+% 2019-1 & -0.005 & 0.133 & -0.143 & 0.000 & 0.252 & 0.000 \\
+% 2019-2 & -0.003 & 0.141 & -0.147 & 0.000 & 0.252 & 0.000 \\
+% 2019-3 & 0.015 & 0.170 & -0.139 & 0.058 & 2.202 & 0.000 \\
+% 2019-4 & 0.002 & 0.190 & -0.185 & 0.000 & 1.460 & -0.024 \\
+% 2020-1 & -0.012 & 0.182 & -0.207 & 0.000 & 0.335 & 0.000 \\
+% 2020-2 & -0.010 & 0.191 & -0.211 & 0.000 & 0.376 & 0.000 \\
+% 2020-3 & 0.679 & 0.954 & 0.403 & 8.718 & 11.704 & 8.556 \\
+% 2020-4 & 0.647 & 1.006 & 0.288 & 5.979 & 8.281 & 5.851 \\
+% 2021-1 & 0.564 & 1.036 & 0.091 & 3.007 & 4.184 & 2.903 \\
+% 2021-2 & 0.533 & 1.024 & 0.042 & 3.007 & 4.225 & 2.903 \\
+% 2021-3 & 0.630 & 1.129 & 0.132 & 7.607 & 10.557 & 7.444
+% \end{tabular}%
+% }
+% \end{table}
 
-Table \ref{tab:fresno_dvs} reports the difference between the pseudo design value (calculated using the reported data) and the imputed design value -- what I believe is an estimate on the design value bias due to low reporting standards.
-We can see the first row is invalid.
-This is common in the full results in the Appendix.
-This occurs because one of the design values have too many invalid days in the quarter (or the 11 quarters before it) the completeness criteria are violated.
-Examining the Annual design value difference column, the numbers do not seem striking though the lower half do seem to be statistically significant.
-However, these are not materially significant -- the EPA has rounding conventions (to the ones place) that would make zero effectively within the lower bound since a design value falling below 0.5 would be rounded to 0.
+\FloatBarrier
 
-These unremarkable results for the annual design value bias (columns 2-4) are fairly representative of the batch of 15 sites.
-When we look at the hour design value results above (columns 5-7), we can see in 2020 and 2021, there were fairly large and significant differences between the design values.
-This indicates that having more restrictive data collection standards could have increased Fresno's design value significantly in these quarters.
-Because OLS was used to impute the missing values, we would expect the hourly imputed values to be unbiased on average.
-Since we are taking 3-year averages, it seems plausible to believe most bias in the imputation method would average out.
+\begin{wrapfigure}{r}{0.5\textwidth}
+    \centering
+    \includegraphics[width=0.5\textwidth]{output/figures/final_results/DV_annual_plot_all_test_sites.png}
+    \smallskip\par
+    \includegraphics[width=\linewidth]{output/figures/final_results/DV_annual_plot_site_019-0500.png}
+    \caption{Difference between the annual pseudo design value from reported data and the predicted design value from filling in missing observations with predicted PM2.5 measurements.
+Top: all 15 sites; Bottom: site 019-0500.}
+    \label{fig:dv_annual}
+\end{wrapfigure}
+
+\noindent\textbf{Annual Design Values.} For each site tested, Figure \ref{fig:dv_annual} graphically shows the difference between the pseudo design value (calculated using the reported data) and the imputed design value -- what I believe is an estimate on the design value bias due to low reporting standards.
+We can see that not all sites have points for all quarters.
+This is common in the full results (Appendix Table \ref{tab:design_value_comparisons}) and this occurs because the site has too many invalid days in one of the 12 quarters used to calculate the design value for that quarter (the completeness criteria are violated).
+
+Note that design values are rounded to the nearest integer.
+So we can see that most of the differences in the graph are much smaller than possible rounding errors and have relatively wide confidence intervals containing zero difference.
+There is one site that jumps out (site 019-0500, Fresno, CA).
+Looking at that site in the bottom of Figure \ref{fig:dv_annual}, we can see that the confidence intervals are significantly different from zero, but are not meaningfully large because the predicted difference in quarter 3 of 2020 is still less than 1 $\mu$g/m$^3$ and are not likely to flip the attainment status.
+
+\FloatBarrier
+\begin{wrapfigure}{R}{0.5\textwidth}
+    \centering
+    \includegraphics[width=0.5\textwidth]{output/figures/final_results/DV_hour_plot_all_test_sites.png}
+    \smallskip\par
+    \includegraphics[width=\linewidth]{output/figures/final_results/DV_hour_plot_site_019-0500.png}
+    \caption{Difference between the 24-hour (98th percentile) pseudo design value from reported data and the predicted design value from filling in missing observations with predicted PM2.5 measurements.
+Top: all 15 sites; Bottom: site 019-0500.}
+    \label{fig:dv_hour}
+\end{wrapfigure}
+
+\noindent\textbf{24-hour Design Values.} Figure \ref{fig:dv_hour} depicts the difference in 24-hour design values.
+We can see in 2020 and 2021, there were fairly large and significant differences between the design values for site 019-0500 (Fresno).
+A positive value indicates that the imputed design value is larger than the design value based only on reported data.
+This means that having a higher data collection standard could have increased Fresno's design value significantly in these quarters.
+The increase in the Fresno 24-hour design value for the third quarter of 2020 is between 8.55 and 11.7 $\mu$g/m$^3$.
+This represents a meaningful difference because the 24-hour design value threshold is 35 $\mu$g/m$^2$ and if an area is near the threshold, a difference of 9 or 10 units could change the attainment status.
+
+% Also note the quick peaking and decline of Fresno's 24-hour design value between 2020 Q3 and 2021 Q2.
+Because the design values are based on 12 quarters, this shape is indicative of high quarterly observations in 2020 Q3 that increased the annual 98th percentile, and the following quarters have lower observations that
+
+\noindent\textbf{24-hour Design Value Confidence Intervals}
+The design value confidence intervals for the annual design values are nearly perfectly symmetric, while the 24-hour design value confidence intervals are highly asymmetric.
+This is expected because the annual design value is very close to a simply average of the daily observations, whereas the 24-hour design value is an order statistic that is sensitive to the placement of just a few predicted values that are higher in PM2.5 than the reported data.
+See Appendix Section \ref{sec:design_values} for a full discussion and example of why we can expect the lower bound of the 24-hour design value confidence interval to be much closer to the predicted value while the upper bound can be further away.
+
+Using OLS to predict the missing EPA hourly observations using PurpleAir data results in prediction errors that are mean 0.
+Since we are taking 3-year averages, it is plausible that the errors from the imputation method would average out.
 These combined with the lower bounds being well above zero suggest there was indeed under-reporting of pollution in Fresno.
@@
 
-Is this meaningful for Fresno though?
-Fresno has had nonattainment status for many years and is listed on several lists within this literature as having notable pollution measurement issues.
+\noindent\textbf{Did Flipping Occur?}
+Is the magnitude of the bias from missing data meaningful for Fresno though?
+Fresno has had nonattainment status for many years and is mentioned within the literature as having notable pollution measurement issues.
 So though Fresno's recorded design value would have been higher, they were already well over the 24-hour design threshold of 35 $\mu$g/m$^3$, having pseudo design values in the 50s and 60s in 2020 and 2021.
+However, pollution in nonattainment counties generally decreases over time.
+So if Fresno were to fall below the threshold and into attainment based on reported values, a downward bias of 8-11 $\mu$g/m$^3$ in the 24-hour design value could be enough to flip the attainment status.
+So while this bias does not play a pivotal role in Fresno at this time, it may be important in the future and may be important for other areas not currently in this study that are closer to the nonattainment threshold.
 
@@
 In regulation of air pollution, one might be concerned with individual actors or regions trying to get around the regulations.
-Since the Clean Air Act and the NAAQS are designed to ensure that all American residents can share in a minimum standard, it seems relevant to me to be able to diagnose mismeasurement.
+Since the Clean Air Act and the NAAQS are designed to ensure that all US residents can share in a minimum standard, it seems relevant to be able to diagnose the occurrence and size of mismeasurement, even if it is occurring at only a few sites in the network.
+Especially considering how sparse regulatory air quality monitors are in the US, a single systematic mismeasurement of pollution can have impacts on millions of people.
 
@@
 
+\subsection{Faculty Feedback} \label{sec:faculty_feedback}
+\input{A_faculty_feedback}
+
+\newpage
 \subsection{Research Project To-do's} \label{sec:todo}
-
 \textbf{Future Work}
@@
 \begin{itemize}
+\setlength{\itemsep}{0.5em}
     \item Valid days: a day that has 18 or more valid hours in it.
@@
 \end{itemize}
-
+%
 \textbf{To construct the annual design value for a given quarter:}
@@
 \end{itemize}
-
+%
 \textbf{To construct the 24-hour design value:}
@@
 \end{itemize}
+
+\newpage\noindent
+\textbf{Asymmetry of the design value confidence intervals:}
+
+The confidence intervals for the annual design value are fairly symmetric (in fact, not perfectly symmetric).
+This is expected because the the confidence interval for each OLS prediction is symmetric and the average design value is nearly a simple average (very close to the first moment of the hourly observations).
+``Nearly a simple average'' because the daily design value is not a simple average of all hourly observations; it takes daily averages over 18 to 24 possibly valid hourly observations, and is a simple average of all valid daily averages.
+Since some daily averages are calculated on fewer than 24 hourly observations and therefore have different denominators, it is close to, but not exactly, a simple average.
+
+The confidence intervals for the 24-hour design value are highly asymmetric.
+This is also expected as the 24-hour design value is an order statistic of the distribution of daily averages, not a moment of the distribution.
+The 24-hour design value takes the 98th percentile of daily averages, thus it is a function of a single daily observation of the right tail of the distribution and is highly sensitive to inserting new values further to the right of it (inserting higher PM2.5 daily observations).
+
+For clarity, consider the following simplied example: assume that we only need one year's worth of daily PM2.5 observations to calculate a 24-hour design value for the fictional Albatross County.
+Further assume that we start with 100 valid daily PM2.5 observations in a single year, used to calculate the 24-hour design value.
+The design value will be the 98th percentile -- \textbf{the third highest} -- of the daily observations.
+
+Say we can fill in 100 missing days with predicted values from PurpleAir measurements.
+The total is now 200 daily observations, of which, we would choose the sixth highest value for the 98th percentile.
+Assume Table \ref{tab:hypothetical-data} display the top three daily observations of the original 100 EPA daily observations and top six daily observations of the 100 filled in missing days.
+The table also includes symmetric upper and lower bounds for the prediction intervals of the daily
+
+% Please add the following required packages to your document preamble:
+% \usepackage{graphicx}
+\begin{table}[h!]
+\caption{Hypothetical PM2.5 daily averages from EPA reported measurements and PurpleAir-imputed daily averages that were missing from the EPA data.
+Column 1: top three original EPA PM2.5 measurements.
+Columns 2-4: The six added daily averages that were greater than the EPA top 3 daily averages, where added daily averages were filled-in from PurpleAir.}
+\label{tab:hypothetical-data}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{EPA original} & PA Upper Bound & PA Predicted & PA Lower Bound \\ \hline
+\multicolumn{1}{|c|}{30} & 50 & 40 & 30 \\ \hline
+\multicolumn{1}{|c|}{29} & 50 & 40 & 30 \\ \hline
+\multicolumn{1}{|c|}{15} & 50 & 40 & 30 \\ \hline
+ & 50 & 40 & 30 \\ \cline{2-4} 
+ & 45 & 35 & 25 \\ \cline{2-4} 
+ & 40 & 30 & 20 \\ \cline{2-4} 
+\end{tabular}%
+}
+\end{table}
+
+\FloatBarrier
+
+Now we can examine the 98th percentile in each of the following four situations: 
+\begin{enumerate}
+    \item the original reported EPA data (column 1) which gives us the original design value
+    \item the EPA data with the upper bound of predicted data (columns 1 and 2) -- the upper bound of the predicted design value
+    \item the EPA data with the predicted data (columns 1 and 3) -- the predicted design value
+    \item the EPA data with the lower bound of predicted data (columns 1 and 4) -- the lower bound of the predicted design value
+\end{enumerate}
+
+In Table \ref{tab:hypothetical-data-sorted}, the data is combined and sorted for each of the four cases.
+We can see that the 98th percentile (bolded) for all three combined datasets (columns 2-4) is larger than the 98th percentile of the original data (column 1).
+We can also see that both the predicted 98th percentile and the lower bound of the prediction are equal to 30, while the upper bound of the prediction is 40.
+This hypothetical example details a simple property of the order statistics for this paper's use case: when adding new predicted data points to the original data, the confidence interval of the 98th percentile (any order statistics in fact) can be highly asymmetric, especially when the predicted values have relatively large prediction intervals.
+
+% Please add the following required packages to your document preamble:
+% \usepackage{graphicx}
+% \usepackage[table,xcdraw]{xcolor}
+% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
+\begin{table}[]
+\caption{Hypothetical top PM2.5 daily observations, combined and sorted from Table \ref{tab:hypothetical-data}.
+Column 1: Sorted top three original EPA daily observations (no new data, 100 days).
+Columns 2-4: Sorted top observations, combining original EPA data with predicted data (the upper bound of the predictions, the prediction, and the lower bound of the predictions, all with 200 total days).
+The original EPA data is in red and the 98th percentile of each dataset is bolded.}
+\label{tab:hypothetical-data-sorted}
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{c|c|c|c|}
+\hline
+\multicolumn{1}{|c|}{EPA original} & Upper Bound & Predicted & Lower Bound \\ \hline
+\multicolumn{1}{|c|}{{\color[HTML]{FF0000} 30}} & 50 & 40 & {\color[HTML]{FF0000} 30} \\ \hline
+\multicolumn{1}{|c|}{{\color[HTML]{FF0000} 30}} & 50 & 40 & {\color[HTML]{FF0000} 30} \\ \hline
+\multicolumn{1}{|c|}{{\color[HTML]{FF0000} \textbf{15}}} & 50 & 40 & 30 \\ \hline
+ & 50 & 40 & 30 \\ \cline{2-4} 
+ & 45 & 35 & {\color[HTML]{FE0000} 30} \\ \cline{2-4} 
+ & \textbf{40} & \textbf{30} & {\color[HTML]{FE0000} \textbf{30}} \\ \cline{2-4} 
+ & {\color[HTML]{FF0000} 30} & {\color[HTML]{FF0000} 30} & 25 \\ \cline{2-4} 
+ & {\color[HTML]{FF0000} 30} & {\color[HTML]{FF0000} 30} & 20 \\ \cline{2-4} 
+ & {\color[HTML]{FF0000} 15} & {\color[HTML]{FF0000} 15} & {\color[HTML]{FF0000} 15} \\ \cline{2-4} 
+\end{tabular}%
+}
+\end{table}
+
+\newpage
+\subsection{Adjusting Confidence Intervals for Multiple Hypotheses}
+\label{sec:multiple_hypothesis}
+\cite{benjamini_adaptive_2006} show that we can use a two-stage procedure for correcting our p-value threshold for the issue of multiple hypothesis testing.
+In this procedure, you first estimate your statistics and test all hypotheses at the uncorrected level of significance.
+You can then use the number of total and rejected hypotheses to correct the significance threshold and conduct the hypothesis tests again.
+Given that the practitioner has p-values of the estimated statistics, this allows one to only run the estimation procedure once.
+Due to the complexity of the design values and the mixture of non-random data with random data,\footnote{In this application, the raw data reported to the EPA are treated as fixed and the missing data are estimated and have uncertainty.} it is significantly less cumbersome to produce confidence intervals at a given level of significance during the estimation process than to produce p-values.
+\noindent The original two-stage method is as follows:
+\begin{align*}
+\intertext{Let $\alpha$ be the desired level of significance, $m$ be the total number of hypotheses being tested, and $m_0$ be the true number of correct null hypotheses. Define}
+\alpha' &= \frac{\alpha}{1+\alpha}
+\intertext{\textbf{Stage 1:} Conduct all $m$ single-hypothesis tests at a level of significance $\alpha$. Denote $r_1$ as the number of rejected null hypotheses.}
+\intertext{\textbf{Stage 2:} Let $\hat m_0 = m - r_1$ be the estimated number of correct null hypotheses. Define}
+\alpha^* &= \alpha' \frac{m}{\hat m_0}
+\intertext{Denote $p_i$ as the p-value from the $i^{th}$ statistic, and let the ordered p-values be $p_{(1)}\leq ...
+\leq p_{(m)}$.
+Then define the number of multiple-hypothesis-corrected rejected hypotheses as}
+k &= \max \{i:  p_{(i)} \leq \alpha^* \frac{i}{m}\}
+\intertext{This $k$ can also be found using the step-down procedure, starting at the largest p-value and iterating to lower p-values until $p_{(i)} \leq \alpha^* \frac{i}{m} = (\alpha'\frac{m}{\hat m_0}) \frac{i}{m} = \alpha'\frac{i}{\hat m_0} $.
+}
+\end{align*}
+In the case of this paper, I did not attain p-values during estimation.
+But it is relatively simple to re-estimate confidence intervals.
+So I have created a slightly modified version of the above procedure in order to iterate through successively larger confidence intervals until I converge on the correct estimate of the number of correct null hypotheses.
+The modified procedure is as follows:
+
+\begin{enumerate}
+    \item Estimate the design value confidence intervals for all $m$ design values using naive $\alpha$-level prediction upper and lower bounds.
+(I have $m=198$ separate design values and use an $\alpha=0.05$ level of confidence.)
+    \item Denote $r_1$ as the number of design values with naive confidence intervals outside of zero.
+Estimate the number of correct null hypotheses in iteration 1 as $\hat m_0^1 = m - r_1$.
+(I had $r_1=24$ confidence intervals that did not include zero, and $\hat m_0^1 = 198 - 24 = 174$ estimated correct null hypotheses.)
+    \item I now need to know if the number of rejected hypotheses $r_1$ change if I adjust the significance level of the prediction intervals.
+But I need to know what level of significance to compare to, since the original procedure makes the comparison $p_{(i)} \leq \alpha'\frac{i}{\hat m_0}$.
+I know the $r_1^{th}$ statistic is the least-significant statistic and is closest to becoming insignificant.
+So testing if $r_1$ (the number of rejected null hypotheses) changes is the same as testing if the $r_1^{th}$ statistics stays significant.
+For the $r_1^{th}$ statistic to be significant, we want $p_{(r_1)} \leq \alpha'\frac{r_1}{\hat m_0^1}$.
+This holds if the $r_1^{th}$ single-hypothesis confidence interval of level $\alpha'\frac{r_1}{\hat m_0^1}$ does not contain zero.
+So I rerun the estimation process, creating prediction intervals at the $\alpha'\frac{r_1}{\hat m_0^1}$ level.
+(Here, $\alpha'\frac{r_1}{\hat m_0^1}  \approx 0.048 \frac{24}{198} \approx 0.0066$)
+    \item Denote $r_2$ as the new number of rejected null hypotheses (confidence intervals not containing zero).
+If $r_2=r_1$, we have applied the correct confidence intervals and the $r_1$ statistics are still significant.
+If $r_2=0$, then we can stop and none of the null hypotheses are rejected.
+If $r_2<r_1$, then we must re-estimate confidence intervals at the new corrected level using $r_2$ and  $\hat m_0^2 = m - r_2$, resulting in $r_3$ rejected null hypotheses.
+We then need to check if $r_3=r_2$.
+    \item Repeat this until $r_k = r_{k-1}$.
+We now have a consistent estimate of the true number of rejected null hypotheses.
+Then, to get the $\alpha$-level multiple hypothesis confidence intervals, the appropriate single-hypothesis confidence interval to apply is of level $\alpha'\frac{r_{k-1}}{\hat m_0^{k-1}}$.
+(Here, this converges after 2 iterations, going from 24 rejected null hypotheses, to 10, then to 8.
+The final level of significance used on the prediction intervals is $\alpha'\frac{r_3}{\hat m_0^3} \approx 0.048 \frac{8}{190} \approx 0.002$.
+This is a $\frac{0.05}{0.002}=25$ factor reduction in the significance level applied to the prediction intervals, which are then propagated through the design value calculation to get 95\% confidence intervals.
+)
+\end{enumerate}
 
@@
 
-\input{z_appendix_site_plots}
+% \input{z_appendix_site_plots}
 \end{document}

### 2022-04-04 paper/ongoing_progress/progress_2022-03-18.tex

--- 2022-03-07 paper/15min-presentation/15min-presentation.tex
@@
 \begin{frame}
-\frametitle{Motivation}
-\textbf{Clean Air}
+\frametitle{State of project}
+\textbf{Data (2016-2021)}
 
 \begin{itemize}
-    \item The Clean Air Act (1970) established National Ambient Air Quality Standards
-    (NAAQS) for US counties
-    \item Either "attainment" or "non-attainment", penalties/forced adoption
-    \item Minimum requirement of 75\% of readings, per quarter
-    \item Air quality can change quickly
-    \item Monitor shutoffs are common
+    \item have: 15 CA EPA sites
+    \item have: all hourly data for PA sites within 50 miles of all 15 EPA sites
+    \item want: all US PA data
+    \item can before April 11: get 10-50 more EPA sites
 \end{itemize}
 \vspace{2em}
-\textbf{Research Questions}
+\textbf{Analysis}
 \begin{itemize}
-\item How biased are EPA monitor-based measures of local air quality?
-\item Does this bias significantly change NAAQS attainment status?
+\item have: Inverse distance weighted mean of PA
+\item have: OLS to predict EPA with PA, give prediction intervals\\
+$EPA_t = \beta_0 + \beta_1 PA^{IDW}_t + \varepsilon_t
+$
+\item want: prediction (1): more complicated OLS with windspeed and direction
+\item prediction (2): machine learning model
+\item can before April 11: probably windspeed and direction model
 \end{itemize}
@@
 \begin{frame}
-\frametitle{EPA National Ambient Air Quality Standards Monitors}
+\frametitle{Data Use Agreement}
 
-\includegraphics[width=\textwidth]{output/figures/epa/combined_uc_cali_epa.png}
+\begin{itemize}
+    \item Gave Adrian Dybwad draft person DUA
+    \item Adrian said legal folks ask if we could do this at the University level
+    \item I asked if we could do a temporary DUA while we get the university level one figured out
+\end{itemize}
+
+What are the first steps to getting a university level DUA?
+
 \end{frame}
@@
 
-%---------------------------------------------------------
-%Example of the \pause command
-\begin{frame}
-\frametitle{National Ambient Air Quality Standards for PM2.5}
-Two standards for assessing compliance for ambient PM2.5 concentrations at the location of a monitor:
-\vspace{2em}
+% %---------------------------------------------------------
+% %Example of the \pause command
+% \begin{frame}
+% \frametitle{National Ambient Air Quality Standards for PM2.5}
+% Two standards for assessing compliance for ambient PM2.5 concentrations at the location of a monitor:
+% \vspace{2em}
 
-\textbf{Daily Design Value} (average)
-\begin{itemize}
-    \item[$\Rightarrow$] 3-year rolling average of 1-year average of daily averages
-    \item[$\Rightarrow$] standard currently set at 15.0 $\mu$g/m$^3$
-\end{itemize}
-\vspace{1em}
+% \textbf{Daily Design Value} (average)
+% \begin{itemize}
+%     \item[$\Rightarrow$] 3-year rolling average of 1-year average of daily averages
+%     \item[$\Rightarrow$] standard currently set at 15.0 $\mu$g/m$^3$
+% \end{itemize}
+% \vspace{1em}
 
-\textbf{24-Hour Design Value} (98$^{th}$ percentile)
-\begin{itemize}
-    \item[$\Rightarrow$] 3-year rolling average of 1-year 98$^{th}$ percentile of daily averages
-    \item[$\Rightarrow$] standard currently set at 35 $\mu$g/m$^3$
-\end{itemize}
-\end{frame}
-%---------------------------------------------------------
+% \textbf{24-Hour Design Value} (98$^{th}$ percentile)
+% \begin{itemize}
+%     \item[$\Rightarrow$] 3-year rolling average of 1-year 98$^{th}$ percentile of daily averages
+%     \item[$\Rightarrow$] standard currently set at 35 $\mu$g/m$^3$
+% \end{itemize}
+% \end{frame}
+% %---------------------------------------------------------
 
-%---------------------------------------------------------
-%Highlighting text
-\begin{frame}
-\frametitle{NAAQS Completeness Criteria for daily monitors}
+% %---------------------------------------------------------
+% %Highlighting text
+% \begin{frame}
+% \frametitle{NAAQS Completeness Criteria for daily monitors}
 
-\textbf{Valid day:}
-minimum of 75\% hours reported  (18 hours)
+% \textbf{Valid day:}
+% minimum of 75\% hours reported  (18 hours)
 
-\vspace{2em}
-\textbf{Valid quarter:}
-minimum of 75\% valid days  (22-23 days\footnote{fewer for monitors that report less frequent observations})
+% \vspace{2em}
+% \textbf{Valid quarter:}
+% minimum of 75\% valid days  (22-23 days\footnote{fewer for monitors that report less frequent observations})
 
-\vspace{2em}
-\textbf{Valid quarterly design value:}
-all 12 quarters in 3-year period (12 quarters)
+% \vspace{2em}
+% \textbf{Valid quarterly design value:}
+% every in 3-year period
+% (12 quarters)
 
-\end{frame}
-%---------------------------------------------------------
+% \end{frame}
+% %---------------------------------------------------------
 
@@
 \frametitle{Observed Completeness of NAAQS Monitors in sample}
-\includegraphics[width=0.8\textwidth]{example-image-a}
+\includegraphics[width=\textwidth]{output/figures/epa_vs_pa/completeness/completeness_epa_pa.png}
 \end{frame}

### 2022-04-04 paper.tex
Updates from Overleaf

@@
     \centering
-    \includegraphics[width=0.5\textwidth]{output/figures/final_results/DV_annual_plot_all_test_sites.png}
+    \includegraphics[width=0.5\textwidth]{output/figures/final_results/DV_annual_plot_all_test_sites_conservative.png}
     \smallskip\par
-    \includegraphics[width=\linewidth]{output/figures/final_results/DV_annual_plot_site_019-0500.png}
+    \includegraphics[width=\linewidth]{output/figures/final_results/DV_annual_plot_site_019-0500_conservative.png}
     \caption{Difference between the annual pseudo design value from reported data and the predicted design value from filling in missing observations with predicted PM2.5 measurements.
+Fill represents the 95\% confidence intervals after correcting for multiple hypothesis testing.
 Top: all 15 sites; Bottom: site 019-0500.}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{output/figures/final_results/DV_hour_plot_all_test_sites.png}
+    \includegraphics[width=0.5\textwidth]{output/figures/final_results/DV_hour_plot_all_test_sites_conservative.png}
     \smallskip\par
-    \includegraphics[width=\linewidth]{output/figures/final_results/DV_hour_plot_site_019-0500.png}
+    \includegraphics[width=\linewidth]{output/figures/final_results/DV_hour_plot_site_019-0500_conservative.png}
     \caption{Difference between the 24-hour (98th percentile) pseudo design value from reported data and the predicted design value from filling in missing observations with predicted PM2.5 measurements.
+Fill represents the 95\% confidence intervals after correcting for multiple hypothesis testing.
 Top: all 15 sites; Bottom: site 019-0500.}
@@
 
-\noindent\textbf{24-hour Design Value Confidence Intervals}
+\noindent\textbf{24-hour Design Value Confidence Intervals.}
 The design value confidence intervals for the annual design values are nearly perfectly symmetric, while the 24-hour design value confidence intervals are highly asymmetric.
@@
 
+As noted in Section \ref{confidence_intervals}, to 95\% construct confidence intervals that are robust to testing many hypotheses, I use an adjusted significance level to produce the OLS prediction intervals.
+In the specific case of this paper, the significance level ends up being about 0.002.
+
+\textbf{Magnitude of the Design Value Statistic.} To help get a sense of the size of gap between the raw design values and the design values with imputed missing values, I have estimated the maximum amount of manipulation left in the raw data.
+For a moment, assume all PM2.5 data must be uploaded for each site at the end of each quarter and that each site has a data manager that can choose which observations to remove from the data before uploading.
+Under the NAAQS rules, all sites can omit up to 25\% of days from each quarter, and up to 25\% of hours from each day that is reported.
+There are no rules about which observations to choose.
+
+Further assume that this fictitious data manager desires to reduce the annual and 24-hour design values as much as possible for their site.
+An easy way to reduce both the average and the 98$^{th}$ percentile of days is to remove the top PM2.5  hours from each day until 75\% of hours remain, create new daily averages from the hours left, and drop the top PM2.5 days from the quarter until the quarter only has 75\% of days left.
+I have assumed the role of this fictitious data manager -- figure \ref{fig:dv_room_for_manipulation}
+
+\begin{wrapfigure}{R}{0.5\textwidth}
+    \centering
+    \includegraphics[width=0.5\textwidth]{output/figures/final_results/DV_hour_plot_all_test_sites_conservative.png}
+    \smallskip\par
+    \includegraphics[width=\linewidth]{output/figures/final_results/DV_hour_plot_site_019-0500_conservative.png}
+    \caption{Difference between the 24-hour (98th percentile) pseudo design value from reported data and the predicted design value from filling in missing observations with predicted PM2.5 measurements.
+Fill represents the 95\% confidence intervals after correcting for multiple hypothesis testing.
+Top: all 15 sites; Bottom: site 019-0500.}
+    \label{fig:dv_room_for_manipulation}
+\end{wrapfigure}
+
 Using OLS to predict the missing EPA hourly observations using PurpleAir data results in prediction errors that are mean 0.
 Since we are taking 3-year averages, it is plausible that the errors from the imputation method would average out.
-These combined with the lower bounds being well above zero suggest there was indeed under-reporting of pollution in Fresno.
+These combined with the lower bounds being well above zero suggest there was indeed under-reporting of pollution in Fresno in 2020 and 2021.
 I cannot say whether or not that it happened intentionally or by mere chance of turning off the monitor at times that would likely have recorded higher pollution.
@@
 \bibliographystyle{chicago}
-\bibliography{references}
+\bibliography{references,references_old}
 % Citation command	Output

### 2022-04-04 paper.tex
Updates from Overleaf

@@
 Removing EE provides more realistic pseudo design value estimates.
+
+\textbf{}
 
@@
 
-\textbf{Magnitude of the Design Value Statistic.} To help get a sense of the size of gap between the raw design values and the design values with imputed missing values, I have estimated the maximum amount of manipulation left in the raw data.
-For a moment, assume all PM2.5 data must be uploaded for each site at the end of each quarter and that each site has a data manager that can choose which observations to remove from the data before uploading.
-Under the NAAQS rules, all sites can omit up to 25\% of days from each quarter, and up to 25\% of hours from each day that is reported.
-There are no rules about which observations to choose.
-
-Further assume that this fictitious data manager desires to reduce the annual and 24-hour design values as much as possible for their site.
-An easy way to reduce both the average and the 98$^{th}$ percentile of days is to remove the top PM2.5  hours from each day until 75\% of hours remain, create new daily averages from the hours left, and drop the top PM2.5 days from the quarter until the quarter only has 75\% of days left.
-I have assumed the role of this fictitious data manager -- figure \ref{fig:dv_room_for_manipulation}
+\FloatBarrier
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{output/figures/final_results/DV_hour_plot_all_test_sites_conservative.png}
+    \includegraphics[width=0.5\textwidth]{output/figures/final_results/room_for_manipulation_hour_DV_all_sites.png}
     \smallskip\par
-    \includegraphics[width=\linewidth]{output/figures/final_results/DV_hour_plot_site_019-0500_conservative.png}
-    \caption{Difference between the 24-hour (98th percentile) pseudo design value from reported data and the predicted design value from filling in missing observations with predicted PM2.5 measurements.
-Fill represents the 95\% confidence intervals after correcting for multiple hypothesis testing.
-Top: all 15 sites; Bottom: site 019-0500.}
+    \includegraphics[width=\linewidth]{output/figures/final_results/room_for_manipulation_hour_DV_site_019-0500_conservative.png}
+    \caption{Room left for manipulation in raw EPA data -- the number of pollution units ($\mu$g/m$^3$) that the 24-hour design value can be reduced by omitting more hourly observations, down the minimum completeness limit set by the NAAQS rules.
+This is the gap between the pseudo design value and a new design value caluclated from fewer observations.
+Top: room left for manipulation for all sites.
+Bottom: room left for manipulation (blue) compared to the estimated gap due to actual missing data (red).}
     \label{fig:dv_room_for_manipulation}
@@
 
+\textbf{Magnitude of the Design Value Statistic.} To help get a sense of the size of gap between the raw design values and the design values with imputed missing values, I have estimated the maximum amount of manipulation left in the raw, reported EPA data.
+For a moment, assume all PM2.5 data must be uploaded to the EPA database for each site at the end of each quarter.
+Also assume that each site has a data manager who can choose which observations to remove from the data before uploading.
+Under the NAAQS rules, all sites can omit up to 25\% of days from each quarter, and up to 25\% of hours from each day that is reported.
+To my knowledge, there are no rules about which observations to choose.
+
+Further assume that this fictitious data manager desires to reduce the annual and 24-hour design values as much as possible for their site.
+An easy way to reduce both the average and the 98$^{th}$ percentile of days is to remove the top PM2.5 hourly observations from each day until 75\% of hours remain, create new daily averages from the hours left, and drop the top PM2.5 days from the quarter until the quarter only has 75\% of days left.
+I have assumed the role of this fictitious data manager -- the top of figure \ref{fig:dv_room_for_manipulation} shows how much the 24-hour design value can be reduced by dropping hourly and daily observations this way.
+The numbers are quite large with most sites having values above 15 $\mu$g/m$^3$ in all available quarters.
+Keeping in mind that the 24-hour NAAQS threshold is currently 35 $\mu$g/m$^3$, this shows that the 75\% completeness rule leaves a meaningful amount of manipulation on the table, even for sites that have missing values already.
+
+The bottom of figure \ref{fig:dv_room_for_manipulation} shows the Fresno site, comparing how the room left for manipulation in the reported data compares to the estimated gap due to actual missing observations.
+
+For Fresno specifically, we can see that there is a lot of room left for intentional manipulation if a bad actor had
+
+\FloatBarrier
 Using OLS to predict the missing EPA hourly observations using PurpleAir data results in prediction errors that are mean 0.
